### 2014-01-23 article_main.tex
prepare to delete projection comparison property

\documentclass[10pt,a4paper]{article}
\author{Мосин Сергей, Зыкин Сергей}
\usepackage{amssymb,amsmath}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{xcolor}
\floatname{algorithm}{Алгоритм}
\newtheorem{theorem}{Теорема}
\newtheorem{mydef}{Определение}
\newtheorem{statement}{Утверждение}
\newtheorem{projectionproperty}{Свойство}
\def \eval #1#2{\left.#1\right\vert_{#2}}
\def \<#1> {\langle #1 \rangle}
\def \n #1{\mathit{#1}}
\begin{document}

\section{Введение}
Многие современные решения в области анализа больших объемов данных
используют технологии аналитической обрабоки данных (Online Analytical
Processing), сформировавшиеся еще в начале 90-х годов \cite{codd}. За последнее
время в этой области было получено большое количество фундаментальных
\cite{lecht,lehner,mazon} и практических результатов \cite{vassi, peder,
progressive, giorg}.

Целью данной работы является изучение проблемы динамического формирования
гиперкубического представления базы данных из реляционной базы данных (РБД) с
помощью зарезервированных представлений данных на компьютере, оснащенном
графическим процессором (GPU). Эта проблема связана с областью оптимизации
запросов, поскольку нацелена на сокращение объема передаваемых данных с сервера
БД. Зарезервированные данные активно используется в системах управления базами
данных (СУБД). Но в большинстве случаев это касается повторного использования
данных, предварительно записанных в буферный пул, без предварительного анализа
содержимого на предмет возможности частичного или комбинированного
использования. Работа СУБД ограничивается тем, что при выполнении очередного
запроса блоки данных не запрашиваются с внешних устройств, если они есть в
буферном пуле, т. е. анализируются номера блоков, а не их содержимое.

В области теоретических исследований ситуация складывается следующим образом.
Огромное число публикаций посвящено проблеме построения оптимального плана
запроса на основе формальных правил, в которых не используются области
определения предикатов в SQL-операторах (логическая оптимизация) либо эти
области учитываются при вычислении статистических оценок для оптимизации
физического доступа к БД. Близкими по методам решения являются задачи
выполнения запросов на потоках данных [9, 10], однако различные, по сравнению
с настоящей работой, цели приводят к различным результатам. Наиболее близка к
рассматриваемой проблеме работа [11]. В ней рассматриваются конъюнктивные
запросы над доменами данных с предикатами в виде арифметических сравнений и
представлены алгоритмы вычисления запросов с использованием существующих
представлений. В настоящей работе рассматривается специальный вид универсального
реляционного запроса над отношениями базы данных, а не над отдельными доменами.
Хотя цели в обеих работах совпадают, результаты различны по указанной причине.
В частности, в настоящей работе нет необходимости разрабатывать алгоритмы, так
как их замещает реляционная алгебра.

Существует ряд работ, в которых рассматривается подход к оптимизации запросов
именно в многомерных базах данных и хранилищах данных.
Есть два основных подхода для резервирования результатов запросов: статический
\cite{baralis, gupta, gupta-mumick} и динамический\cite{scheuermann, shim,
kalnis, chang}. Первый базируется на использовании набора фиксированных
запросов, во втором же предполагается динамический выбор результатов запросов
для резервирования на основе статистики появления а также вычислительной
стоимости выполнения запросов. В отличие от подходов, применяемых в реляционных
СУБД, здесь проводится анализ самих запросов. Однако подходы, описанные в
данных статьях, применимы лишь для технологий реляционных многомерных баз данных
(ROLAP) и используют специальные для многомерных баз данных типы запросов, в то
время как в нашем случае используются стандартные SQL запросы. Данная работа
базируется на результатах, полученных в статье \cite{zyk_pol}. Устранено
условие, ограничивающее атрибуты промежуточного представления, а также сделано
обобщение теорем на случай нескольких промежуточных представлений. В заключение
приводятся алгоритмы, позволяющие применять на практике полученные результаты.

\section{Определение и свойства логических ограничений}

Будем рассматривать логические формулы в дизъюнктивной нормальной форме. Тогда
формула $F$ имеет вид

\begin{equation}
F = K_1 \vee K_2 \vee \dots \vee K_m
\label{def_F_1}
\end{equation}
\begin{equation}
K_i = T_1 \& T_2 \dots \& T_n, i = 1, \dots, m
\label{def_F_2}
\end{equation}

$T_j, j = 1, \dots, n$ - предикаты, в которых явным образом
специфицированы расширенные имена атрибутов $R_i.A_j$ (атрибут $A_j$ в отношении
$R_i$):
\begin{itemize}
  \item	операция сравнения $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ – операция
  сравнения $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$ –
  согласованные по типам допустимые выражения, определенные на множестве
  расширенных имен атрибутов и констант;
  \item операция $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
  \n{Expr}_3$ (содержимое в прямоугольных скобках $[*]$ для предиката не
  является обязательным при написании);
  \item операция $\n{Expr} \n{[NOT]} \n{IN} S$, где $S$ – список значений либо
  подзапрос, результатом которого является столбец атрибута $A_j$ в отношении
  $R_i$;
  \item операция $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, где $\n{Str}_i$ –
  строки;
  \item операция $\n{Expr} \theta \n{ALL/ANY} S$.
\end{itemize}
Множество атрибутов, входящих в формулу, выражает размерность формулы и
обозначается $\<F> $.

\begin{equation}
\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\}
\label{def_F_3}
\end{equation}

Перечисленные варианты операций используют не все возможности языка SQL.
Например, предикат $\n{EXISTS}$ не используется, поскольку в нем явно не
специфицированы расширенные имена атрибутов, предикат $\n{NULL}$ используется
в данной работе для другой цели.

При вычислении логического выражения может быть получено значение $\n{UNKNOWN}$,
если на текущем кортеже $t$ атрибут принимает значение $\n{NULL}$, поскольку
результаты вычисления логических выражений в SQL-запросах соответствуют
трехзначной логике. Для избежания такой ситуации формулы преобразуются
определенным образом, изложенным в статье \cite{zyk_pol}.

В дальнейшем будем использовать множество $\mathcal{A} =$ $\{(a_1, \dots, a_l)
\mid a_i \in Dom(A_i), i=1,\dots,l\}$, $l$-мерное пространство значений всех
атрибутов базы данных.

\begin{mydef}
Областью истинности логической формулы $F$, заданной (\ref{def_F_1}),
(\ref{def_F_2}), (\ref{def_F_3}), является множество, определяемое по следующему правилу: $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
\end{mydef}

\textit{Замечание 1}: Размерность формулы $F$ может быть меньше
размерности $\mathcal{A}$, при этом атрибуты, не входящие в формулу, могут
принимать любые значения.
\\
\textit{Замечание 2}:  Сложность компонентов предикатов $\n{Expr}$, $S$ и
$\n{Str}$ определяется возможностями программного обеспечения по вычислению
областей определения $\n{Dom} (F_v)$, что необходимо для вычисления $TJ(C_s)$ из
промежуточных представлений $P_v$.

В соответствии с данными определениями легко понять, как будут устроены операции
над областями, соответствующими формулам. $\n{Dom} (F)$ для некоторой формулы
$F$, заданной своей ДНФ, является объединением областей, представленных
отдельными конъюнктами формулы. Область $M(F)$ является подобластью области
$\n{Dom} (F)$, то есть $M(F) \subseteq \n{Dom} (F)$. Включение $M(F_1) \subseteq
M(F_2)$ означает, что область истинности каждого конъюнкта $F_1$ является
подобластью некоторого объединения областей истинности конъюнктов $F_2$.

Теперь введем определения, касающиеся модификации вхождения атрибутов в
логические формулы.

\begin{mydef}
Проекцей логической формулы $F$, заданной (\ref{def_F_1}), (\ref{def_F_2}),
(\ref{def_F_3}), на множество атрибутов $X$ называется логическая
формула $F[X], \<F[X]> $
$= X$, в которой все термы, содержащие
атрибуты $R_i^F.A_i^F \notin X$, заменены на тривиальный терм TRUE.
\label{projection}
\end{mydef}

\begin{projectionproperty}[Свойство включения]
$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
\label{proj_property_of_inclusion}
\end{projectionproperty}
\begin{proof}
Рассмотрим некоторую точку $a \in M(F)$. 
Так как формула находится в ДНФ, то замена некоторого предиката $T$ на значение
$\n{TRUE}$, очевидно, может либо полностью сделать значение формулы $\n{TRUE}$,
если предикат $T$ единственный в конъюнкте $K_i = T$, либо сократить количество
условий в нём, если $K_i = T \& T'_1 \dots \& T'_n$. В любом из этих случаев
количество условий в формуле $F[X]$ лишь уменьшается, что гарантирует выполнение
$F[X](a) = \n{TRUE}$ и, следовательно, $a \in M(F[X])$
\end{proof}

\begin{projectionproperty}[Свойство сравнения]
$M(F_1) \subseteq M(F_2) \Rightarrow M(F_1[X]) \subseteq$\\ $\subseteq M(F_2[X])$, $X = \<F_2> \cap \<F_1> $
\label{proj_property_of_comparison}
\end{projectionproperty}
\begin{proof}
Это действительно так, ведь включение одного множества в другое означает
покоординатное включение точек, принадлежащих первому множеству. После взятия
проекции в точках обоих множеств остается один и тот же набор координат, по
которым точки принимают произвольные значения. Оставшиеся координаты сохраняют
отношение  включения, так как они не меняются при проекции. Таким образом, в
совокупности вновь достигается включение множеств.
\end{proof}

\begin{statement}
$M(F_1) \subseteq M(F_2) \Rightarrow \<F_2> \subseteq \<F_1> $
\label{attibute_inclusion_property}
\end{statement}
\begin{proof}
Пусть это условие не выполняется. Тогда $U' = \<F_2> \setminus \<F_1> \neq
\emptyset$. Выберем точку $a = (a_1, \dots, a_l) \in M(F_1)$. Для некоторого
атрибута $A_i \in U'$ выберем соответствующую координату $a_i \in Dom(A_i)$. Так
как по условию в формулах нет тривиальных условий на атрибуты, то $\exists a_i'
\in Dom(A_i) : a' = (a_1, \dots, a_{i - 1}, a_i', a_{i + 1}, \dots, a_l) \notin
M(F_2)$. Однако $a' \in M(F_1)$, так как атрибут $A_i$ отсутствует в формуле
$F_1$. Следовательно, $M(F_1) \nsubseteq M(F_2)$. Полученное противоречие
доказывает утверждение.
\end{proof}

\section{Исследование свойств промежуточных представлений данных}

\input{theorem_1}

\input{theorem_2}

Эти результаты можно обобщить, взяв в рассмотрение целый набор промежуточных
представлений $P_v$. Но сперва приведем один факт, связанный со свойствами
естественного соединения.

\begin{statement}
Пусть $\Re_1 = R_1 \bowtie \dots \bowtie R_k$ - результат естественного
соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \bowtie \dots
\bowtie R_k \bowtie R_{k+1} \bowtie \dots \bowtie R_{n}$. 
Тогда $\Re_2 [\langle \bowtie_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
\label{join_property}
\end{statement}

Это действительно так, ведь после проведения операции естественного соединения
с дополнительными отношениями исходные кортежи, содержавшиеся в $\Re_1$, могут
лишь ''отсеяться'' операцией естественного соединения. Тогда, взяв вырезку по
исходным атрибутам, получим как максимум то же самое отношение $\Re_1$.

\input{theorem_1_mult}

\input{theorem_2_mult}

\section{Удаление лишних кортежей из промежуточных представлений}

Результаты, полученные в теоремах \ref{th_base}, \ref{th_mult} нельзя сразу
использовать на практике. Отношение $P^{\ast}$ содержится в имеющемся
представлении, однако неизвестно, какие именно кортежи являются частью
$P^{\ast}$. Лишние кортежи могут возникнуть по двум причинам:
\begin{enumerate}
  \item В силу Утверждения \ref{proj_property}, так как правая часть содержит
  проекцию исходной формулы $F^{\ast}$ на некоторое множество атрибутов.
  \item В силу Утверждения \ref{join_property}, так как в промежуточном
  представлении отношений не больше, чем в искомом $P^{\ast}$.
\end{enumerate}
Поэтому необходимо сперва удалить лишние кортежи (если они есть) из
промежуточных представлений.

\input{alg_filter_tj}

\input{bibliography}
\end{document}

### 2014-01-23 bibliography.tex
prepare to delete projection comparison property

\begin{thebibliography}{9}
\bibitem{codd}
  \textit{Codd E.F., Codd S.B., Salley C.T.}
  Providing OLAP (On-line Analytical Processing) to User-Analysts: An IT
  Mandate. – Sunnyvale(CA): Codd \& Date Inc.,
  1993.
  – 31 p.
\bibitem{lecht}
  \textit{Lechtenborger J., Vossen G.}
  Multidimensional normal forms for data warehouse design
  // Inf. Syst.
  – 2003.
  – Vol. 28, N 5. – P. 415–434. 
\bibitem{lehner}
  \textit{Lehner W., Albrecht J., Wedekind H.}
  Normal forms for multidimensional databases
  // Proc. of the Tenth Intern. Conf. on Scientific and Statistical Database
  Management.
  – Capri, 1998.
  – P. 63–72.
\bibitem{mazon}
  \textit{Mazon J.-N., Trujillo J., Lechtenborger J.}
  Reconciling requirement-driven data warehouses with data sources via
  multidimensional normal forms
  // Data Knowledge Engineering.
  – 2007.
  – Vol. 63, N 3. – P. 725–751.
\bibitem{vassi}
  \textit{Vassiliadis P., Sellis T.}
  A survey of logical models for OLAP databases
  // SIGMOD Rec.
  – 1999.
  – Vol. 28, N 4. – P. 64–69.
\bibitem{peder}
  \textit{Pedersen T.B., Jensen C.S., Dyreson C.E.}
  A foundation for capturing and querying complex multidimensional data
  // Inf. Syst.
  – 2001.
  – Vol. 26, N 5. – P. 383–423.
\bibitem{progressive}
  \textit{Progressive ranking of range aggregates}
  / H.-G. Li, et al.
  // Data \& Knowledge Engineering.
  – 2007.
  – Vol. 63, N 1. – P. 4–25. 
\bibitem{giorg}
  \textit{Giorgini P., Rizzi S., Garzetti M.}
  Goal-oriented requirement analysis for data warehouse design
  // Proc. of the 8th ACM international Workshop on Data Warehousing and OLAP: DOLAP '05.
  – Bremen, 2005.
  – P. 47–56.
%static approach
\bibitem{baralis}
  \textit{Baralis, E., Paraboschi, S., Teniente, E.} Materialized view selection in a multidimensional database.
  // Proc. of the 23rd International Conference on Very Large Data Bases,
  Athens, Greece.
  - 1997
  - P. 318–329
\bibitem{gupta}
  \textit{Gupta, H.}
  Selection of views to materialize in a data warehouse.
  // Proceedings of the International Conference on Database Theory, Delphi.
  - Greece, 1997.
  - P. 98–112.
\bibitem{gupta-mumick}
  \textit{Gupta, H., Mumick, I.S.}
  Selection of views to materialize under a maintenance cost constraint.
  // Proceedings of the International Conference on Database Theory.
  - Israel, 1999.
  - P. 453–470.
%dynamic
\bibitem{scheuermann}
  \textit{Scheuermann, P., Shim, J., Vingralek R.}
  WATCHMAN: A data warehouse intelligent cache manager.
  // Proceedings of the 22nd International Conference on Very Large Data Bases.
  - Bombay, India, 1996.
  - P. 51–62.
\bibitem{shim}
  \textit{Shim, J., Scheuermann, P., Vingralek R.}
  Dynamic caching of query results for decision support systems.
  // Proceedings of the 11th International Conference on Scientific and
  Statistical Database Management.
  - Cleveland, OH, 1999.
  - P. 254–263.
\bibitem{kalnis}
  \textit{Kalnis, P., Papadias, D.}
  Proxy-server architectures for OLAP.
  // Proceedings of the 2001 ACM SIGMOD International Conference on Management
  of Data.
  - Santa Barbara, CA, 2001.
  - P. 367–378.
\bibitem{chang}
  \textit{Chang-Sup Park, Myoung Ho Kim b, Yoon-Joon Lee}
  Usability-based caching of query results in OLAP systems.
  // The Journal of Systems and Software
  - 2003
  - Vol. 68, - P. 103–119

%ours  
\bibitem{zyk_pol}
  \textit{С.В. Зыкин, А.Н. Полуянов}
  Формирование многомерных данных с использованием промежуточных представлений.
  "Проблемы управления".
  2013.
  № 5. Стр. 54-59.

\end{thebibliography}

### 2014-01-23 theorem_1.tex
prepare to delete projection comparison property

\author{Мосин Сергей, Зыкин Сергей}
\begin{theorem}
$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
, если:
\\а) $X^{\ast} \subseteq X_{v}$
\\б) 
$ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\}
\subseteq
\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
\\в) $\<F_v> \subseteq \<F^{\ast}> , M (F^{\ast}[\<F_v> ]) \subseteq
M (F_{v}) $.
\label{th_base}
\end{theorem} 
\begin{proof}
Пусть существует кортеж $t^{\ast} \in P^{\ast}$. По условию теоремы надо показать, 
что $t^{\ast} \in \pi_{X^{\ast}} (P_{v})$. Из условия $t^{\ast} \in P^{\ast}$ следует, 
что существует кортеж $t' \in R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots
\bowtie R^{\ast}_l$ и $t^{\ast} = t'[X^{\ast}]$. Заметим, что такой кортеж 
определяется не единственным образом. Среди возможных вариантов выберем такой, 
который удовлетворяет $F^{\ast}(t')=TRUE$. Такой кортеж существует, так как иначе
кортежа $t^{\ast}$ не было бы в $P^{\ast}$. Таким образом, существуют кортежи
$t^{\ast}_i \in R^{\ast}_i$:

\begin{equation}
t^{\ast}_i = t'[\langle R^{\ast}_i\rangle], i = 1,\dots,l,
\label{eval_eq_1}
\end{equation}
\def \intersecij {\langle R^{\ast}_i \rangle \cap \langle R^{\ast}_j \rangle}
и для любых пар $i$ и $j$, для которых $\intersecij \neq \emptyset$, выполнено:
\begin{equation}
t^{\ast}_i[\intersecij] = t'_i[\intersecij], i,j = 1,\dots,l,
\label{eval_eq_2}
\end{equation}

Поскольку условия (\ref{eval_eq_1}) и (\ref{eval_eq_2}) выполнены для всего множества
отношений $\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\}$, то они выполнены и для любого его
подмножества, в том числе и для $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\}$.
Следовательно, кортежи $t^{\ast}_i$ из отношений $R^{v}_j$ при выполнении операции
естественного соединения образуют кортеж $t''$, для которого выполнено:
$t'' = t'[\langle {\bowtie}_{i=1}^{s(v)} R^v_i \rangle]$. В силу условия а) и 
$X_v \subseteq \langle {\bowtie}_{i=1}^{s(v)} R^v_i \rangle$ получаем 
$t''[X^{\ast}] = t'[X^{\ast}] = t^{\ast}$.

Из условия  $F^{\ast}(t') = \n{TRUE}$, по утверждению
\ref{proj_property}, получаем, что проекции $F^{\ast}[\<F_v> ]$ и $F^{\ast} [X^{\ast}]$ верны на кортеже $t'$, а следовательно и на $t''$, так как формулы определены на общих атрибутах этих кортежей. Далее, по условию 3, $F_v (t'') = \n{TRUE}$.
Это значит, что $t^{\ast} = t''[X^{\ast}] \in \pi_{X^{\ast}} (P_{v})$. Теорема
доказана.
\end{proof}

### 2014-01-23 theorem_2.tex
prepare to delete projection comparison property

\author{Мосин Сергей, Зыкин Сергей}
\begin{theorem}
$P^{\ast} = \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
, если: 
\textcolor{red}{!!! Селекцию можно убрать $: P^{\ast} = \pi_{X^{\ast}} (P_{v})$}
\\а) $X^{\ast} \subseteq X_{v}$
\\б) 
$\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\}$
\\в) $\<F_v> = \<F^{\ast}> , M (F^{\ast}) = M (F_{v}) $.
\label{th_base_eq}
\end{theorem} 
\begin{proof}
Поскольку условия теоремы являются частным случаем теоремы \ref{th_base}, то
включение $P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
доказано. Необходимо показать, что
$\pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v})) \subseteq P^{\ast}$. Пусть
кортеж $t_v \in \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$. Покажем, что
$t_v \in P^{\ast}$. Из определения кортежа $t_v$ следует, что существует кортеж  $t' \in R^v_1 \bowtie R^v_2\bowtie\ldots \bowtie R^v_{s(v)}$ и $t_v = t'[X^{\ast}]$, $F_v(t')=\n{TRUE}$. По свойству операции естественного соединения $R^v_1 \bowtie R^v_2\bowtie\ldots \bowtie R^v_{s(v)} = R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots \bowtie R^{\ast}_l$, следовательно, $t' \in R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots \bowtie R^{\ast}_l$. В силу условия в) $F^{\ast} (t')=\n{TRUE}$. Отсюда следует, что $t_v \in P^{\ast}$. Теорема доказана.
\end{proof}

### 2014-01-23 article_main.tex
intermediate result point. All projections removed. 2 and 4th theorems lost their F[X] conditions.

@@
 \newtheorem{statement}{Утверждение}
-\newtheorem{projectionproperty}{Свойство}
+\newtheorem*{consequence}{Следствие}
+\newtheorem{note}{Замечание}
 \def \eval #1#2{\left.#1\right\vert_{#2}}
@@
 \end{itemize}
+
+\begin{note}
+Здесь и далее будем полагать, что в формулах нет тривиальных условий на
+атрибуты, то есть условий вида $Expr_1 = Expr_1$ и сводящихся к ним. Такие
+условия могут быть удалены из формулы, не меняя области истинности (будет
+определено далее).
+\label{trivial}
+\end{note}
+
+\begin{mydef}
 Множество атрибутов, входящих в формулу, выражает размерность формулы и
@@
 \end{equation}
+\end{mydef}
 
@@
 Областью истинности логической формулы $F$, заданной (\ref{def_F_1}),
-(\ref{def_F_2}), (\ref{def_F_3}), является множество, определяемое по следующему правилу: $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
+(\ref{def_F_2}), (\ref{def_F_3}), является множество, определяемое по следующему
+правилу: $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
 \end{mydef}
 
-\textit{Замечание 1}: Размерность формулы $F$ может быть меньше
+\begin{note}
+Размерность формулы $F$ может быть меньше
 размерности $\mathcal{A}$, при этом атрибуты, не входящие в формулу, могут
 принимать любые значения.
-\\
-\textit{Замечание 2}:  Сложность компонентов предикатов $\n{Expr}$, $S$ и
-$\n{Str}$ определяется возможностями программного обеспечения по вычислению
-областей определения $\n{Dom} (F_v)$, что необходимо для вычисления $TJ(C_s)$ из
-промежуточных представлений $P_v$.
+\end{note}
+
+\begin{note}
+Сложность компонентов предикатов $\n{Expr}$, $S$ и $\n{Str}$ определяется
+возможностями программного обеспечения по вычислению областей определения
+$\n{Dom} (F_v)$, что необходимо для вычисления $TJ(C_s)$ из промежуточных
+представлений $P_v$.
+\end{note}
 
@@
 
-\begin{projectionproperty}[Свойство включения]
+\begin{statement}[Свойство включения]
 $\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
 \label{proj_property_of_inclusion}
-\end{projectionproperty}
+\end{statement}
 \begin{proof}
@@
 
-\begin{projectionproperty}[Свойство сравнения]
-$M(F_1) \subseteq M(F_2) \Rightarrow M(F_1[X]) \subseteq$\\ $\subseteq M(F_2[X])$, $X = \<F_2> \cap \<F_1> $
-\label{proj_property_of_comparison}
-\end{projectionproperty}
-\begin{proof}
-Это действительно так, ведь включение одного множества в другое означает
-покоординатное включение точек, принадлежащих первому множеству. После взятия
-проекции в точках обоих множеств остается один и тот же набор координат, по
-которым точки принимают произвольные значения. Оставшиеся координаты сохраняют
-отношение  включения, так как они не меняются при проекции. Таким образом, в
-совокупности вновь достигается включение множеств.
-\end{proof}
-
 \begin{statement}
@@
 атрибута $A_i \in U'$ выберем соответствующую координату $a_i \in Dom(A_i)$. Так
-как по условию в формулах нет тривиальных условий на атрибуты, то $\exists a_i'
+как по замечанию \ref{trivial} в формулах нет тривиальных условий на атрибуты, то $\exists a_i'
 \in Dom(A_i) : a' = (a_1, \dots, a_{i - 1}, a_i', a_{i + 1}, \dots, a_l) \notin
@@
 доказывает утверждение.
+\end{proof}
+
+\begin{consequence}
+$M(F_1) = M(F_2) \Rightarrow \<F_2> = \<F_1> $
+\label{attibute_equality_property}
+\end{consequence}
+\begin{proof}
+Пусть это условие не выполняется. Тогда, в силу предыдущего утверждения,
+существует множество $U' = \<F_1> \setminus \<F_2> \neq \emptyset$. Аналогично
+предыдущему доказательству, выбираем точку из любого из множеств и меняем
+значение координаты, соответствующей атрибуту из множества $U'$ на такое,
+которое не принаджежит множеству $M(F_1)$. Полученная точка по-прежнему
+принадлежит $M(F_2)$. Мы снова пришли к противоречию. Утверждение доказано.
 \end{proof}
@@
 \begin{enumerate}
-  \item В силу Утверждения \ref{proj_property}, так как правая часть содержит
-  проекцию исходной формулы $F^{\ast}$ на некоторое множество атрибутов.
+  \item В силу Утверждения \ref{proj_property_of_inclusion}, так как правая
+  часть содержит проекцию исходной формулы $F^{\ast}$ на некоторое множество
+  атрибутов.
   \item В силу Утверждения \ref{join_property}, так как в промежуточном

### 2014-01-23 theorem_1.tex
intermediate result point. All projections removed. 2 and 4th theorems lost their F[X] conditions.

@@
 \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
-\\в) $\<F_v> \subseteq \<F^{\ast}> , M (F^{\ast}[\<F_v> ]) \subseteq
-M (F_{v}) $.
+\\в) $M (F^{\ast}) \subseteq M (F_{v}) $.
 \label{th_base}
@@
 Из условия  $F^{\ast}(t') = \n{TRUE}$, по утверждению
-\ref{proj_property}, получаем, что проекции $F^{\ast}[\<F_v> ]$ и $F^{\ast} [X^{\ast}]$ верны на кортеже $t'$, а следовательно и на $t''$, так как формулы определены на общих атрибутах этих кортежей.
-Далее, по условию 3, $F_v (t'') = \n{TRUE}$.
+\ref{proj_property_of_inclusion}, получаем, что проекция $F^{\ast} [X^{\ast}]$
+верна на кортеже $t'$, а следовательно и на $t''$, так как формула определена на
+общих атрибутах этих кортежей. Далее, по условию 3, $F_v (t') = \n{TRUE} \Rightarrow F_v (t'') = \n{TRUE}$.
 Это значит, что $t^{\ast} = t''[X^{\ast}] \in \pi_{X^{\ast}} (P_{v})$. Теорема

### 2014-01-23 theorem_2.tex
intermediate result point. All projections removed. 2 and 4th theorems lost their F[X] conditions.

@@
 \begin{theorem}
-$P^{\ast} = \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
+$P^{\ast} = \pi_{X^{\ast}} (P_{v})$
 , если: 
-\textcolor{red}{!!! Селекцию можно убрать $: P^{\ast} = \pi_{X^{\ast}} (P_{v})$}
+%\textcolor{red}{!!! Селекцию можно убрать $: P^{\ast} = \pi_{X^{\ast}} (P_{v})$}
 \\а) $X^{\ast} \subseteq X_{v}$
@@
 $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\}$
-\\в) $\<F_v> = \<F^{\ast}> , M (F^{\ast}) = M (F_{v}) $.
+\\в) $M (F^{\ast}) = M (F_{v}) $.
 \label{th_base_eq}
@@
 Поскольку условия теоремы являются частным случаем теоремы \ref{th_base}, то
-включение $P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
+включение $P^{\ast} \subseteq \pi_{X^{\ast}} (P_{v})$
 доказано. Необходимо показать, что
-$\pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v})) \subseteq P^{\ast}$. Пусть
-кортеж $t_v \in \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$. Покажем, что
+$\pi_{X^{\ast}} (P_{v}) \subseteq P^{\ast}$. Пусть
+кортеж $t_v \in \pi_{X^{\ast}} (P_{v})$. Покажем, что
 $t_v \in P^{\ast}$.

### 2014-02-28 article_main.tex
removed algorithm realization. No need for this paper
zykin changes and my corrections

@@
 \def \n #1{\mathit{#1}}
+\title{Формирование табличных приложений с использование зарезервированных представлений данных}
+\author{
+Мосин С.В., Зыкин С.В. \footnote{Работа выполнена при поддержке гранта РФФИ № 12-07-00066-а}
+}
+\date{}
+
 \begin{document}
+\maketitle
 
 \section{Введение}
+
 Многие современные решения в области анализа больших объемов данных
-используют технологии аналитической обрабоки данных (Online Analytical
+используют технологии аналитической обработки данных (Online Analytical
 Processing), сформировавшиеся еще в начале 90-х годов \cite{codd}. За последнее
@@
 
-Целью данной работы является изучение проблемы динамического формирования
-гиперкубического представления базы данных из реляционной базы данных (РБД) с
-помощью зарезервированных представлений данных на компьютере, оснащенном
-графическим процессором (GPU). Эта проблема связана с областью оптимизации
+Целью данной работы, включая результаты этой статьи, является изучение
+проблемы динамического формирования
+многомерного представления данных из реляционной базы данных с
+помощью зарезервированных представлений на компьютере пользователя, оснащенном
+графическим процессором (GPU). При этом не предполагается какое-либо преобразование
+исходной схемы базы данных, в том числе к иерархическому виду.
+
+Поставленная проблема связана с областью оптимизации
 запросов, поскольку нацелена на сокращение объема передаваемых данных с сервера
-БД. Зарезервированные данные активно используется в системах управления базами
+базы данных. Зарезервированные данные активно используется в системах управления базами
 данных (СУБД). Но в большинстве случаев это касается повторного использования
-данных, предварительно записанных в буферный пул, без предварительного анализа
+данных, записанных в кеш, без предварительного анализа
 содержимого на предмет возможности частичного или комбинированного
@@
 запроса блоки данных не запрашиваются с внешних устройств, если они есть в
-буферном пуле, т. е. анализируются номера блоков, а не их содержимое.
+кеше, т. е. анализируются номера блоков, а не их содержимое.
+
+Данная работа основывается на результатах, полученных в
+статье \cite{zyk_pol}. Устранено условие, ограничивающее атрибуты промежуточного
+представления, а также сделано обобщение теорем на случай нескольких
+промежуточных представлений.
+
+\section{Обзор публикаций и сравнение}
 
@@
 области учитываются при вычислении статистических оценок для оптимизации
-физического доступа к БД. Близкими по методам решения являются задачи
-выполнения запросов на потоках данных [9, 10], однако различные, по сравнению
+физического доступа к базе данных. Близкими по методам решения являются задачи выполнения
+запросов на потоках данных \cite{Olston03, Denny05}, однако различные, по сравнению
 с настоящей работой, цели приводят к различным результатам. Наиболее близка к
-рассматриваемой проблеме работа [11]. В ней рассматриваются конъюнктивные
+рассматриваемой проблеме работа \cite{Afrati06}. В ней рассматриваются конъюнктивные
 запросы над доменами данных с предикатами в виде арифметических сравнений и
@@
 Хотя цели в обеих работах совпадают, результаты различны по указанной причине.
-В частности, в настоящей работе нет необходимости разрабатывать алгоритмы, так
-как их замещает реляционная алгебра.
+В частности, в настоящей работе нет необходимости разрабатывать алгоритмы выборки
+данных из промежуточных представлений, так как их замещает реляционная алгебра.
 
-Существует ряд работ, в которых рассматривается подход к оптимизации запросов
-именно в многомерных базах данных и хранилищах данных.
-Есть два основных подхода для резервирования результатов запросов: статический
-\cite{baralis, gupta, gupta-mumick} и динамический\cite{scheuermann, shim,
-kalnis, chang}. Первый базируется на использовании набора фиксированных
-запросов, во втором же предполагается динамический выбор результатов запросов
-для резервирования на основе статистики появления а также вычислительной
-стоимости выполнения запросов. В отличие от подходов, применяемых в реляционных
-СУБД, здесь проводится анализ самих запросов. Однако подходы, описанные в
-данных статьях, применимы лишь для технологий реляционных многомерных баз данных
-(ROLAP) и используют специальные для многомерных баз данных типы запросов, в то
-время как в нашем случае используются стандартные SQL запросы. Данная работа
-базируется на результатах, полученных в статье \cite{zyk_pol}. Устранено
-условие, ограничивающее атрибуты промежуточного представления, а также сделано
-обобщение теорем на случай нескольких промежуточных представлений. В заключение
-приводятся алгоритмы, позволяющие применять на практике полученные результаты.
+Существует ряд работ, в которых рассматривается проблема оптимизации
+запросов к хранилищам данных, имеющих концептуальное многомерное представление,
+с использованием динамически формируемых представлений. При этом
+существуют два основных подхода для резервирования результатов запросов:
+статический \cite{baralis, gupta, gupta-mumick} и динамический \cite{scheuermann, shim,
+kalnis, chang}. Первый базируется на использовании набора фиксированных запросов,
+во втором же предполагается динамический выбор результатов запросов для резервирования на
+основе статистики появления а также вычислительной стоимости выполнения
+запросов. В качестве источника данных используются хранилища данных, либо
+реляционная база данных, преобразованная к иерархическому виду. При этом,
+рассматриваются специальные запросы манипулирования многомерными данными. 
+В этой статье рассматривается технология раздельного формирования размерностей
+представления данных, необходимого для текущего анализа. Поэтому интерес
+представляют стандартные SQL запросы к базе данных и промежуточным представлениям. 
+Результирующее представление может быть затем обработано специальными запросами
+манипулирования многомерными данными.
+
+В работе \cite{Keller96} решается задача анализа содержимого кеша,
+при этом промежуточным представлениям
+в кеше ставятся в соответствие предикаты. Проблема использования представлений
+решается за счет выводимости предикатов. В данной работе проблему использования
+промежуточных представлений предлагается решать за счет вычисления областей
+определения (доменов) результатов выполнения запросов, находящихся в кеше,
+соответствующих промежуточным представлениям. Это дает возможность
+определить кортежи, которые можно использовать в зарезервированных представлениях
+при формировании нового представления, и на основе вычисления областей определения
+сформировать SQL команды, которые позволят загрузить недостающие данные с сервера
+базы данных. Для демонстрации предлагаемых решений рассмотрим пример.
+
+{\bf Пример 1.} Допустим, что требуется получить представление данных, как показано в
+таблице \ref{def_T_1}. Жирным шрифтом в этой таблице представлены атрибуты размерностей,
+курсивом - атрибуты мер, значения атрибутов - обычным шрифтом.
+
+\begin{table}[h!]
+\caption{\label{def_T_1} Фрагмент учебного плана.}
+\begin{center}
+\begin{tabular}{|c|c||c|c|c|c|}
+\cline{2-6}
+\multicolumn{1}{c|}{} & \multicolumn{1}{|r||}{\bf Семестр} & \multicolumn{2}{|c|}{2} & \multicolumn{2}{|c|}{3}  \\
+\cline{2-6}
+\multicolumn{1}{c|}{} & {\bf Предмет} & Физика & Химия & Физика & Химия \\
+\hline \cline{2-6}
+{\bf Факультет} & {\bf Группа} & {\it Часы} & {\it Часы} & {\it Часы} & {\it Часы} \\
+\hline
+ & Х-321 & & 36 & 18 & 9 \\
+\cline{2-6}
+Химический & Х-322 & & 48 & 18 & 9 \\
+\cline{2-6}
+ & Х-331 & & 24 & 18 & 9 \\
+\hline
+ & Ф-321 & 48 & & 24 & \\
+\cline{2-6}
+Физический & Ф-322 & 48 & & 24 & \\
+\cline{2-6}
+ & Ф-331 & 64 & & 24 & \\
+\hline
+\end{tabular}
+\end{center}
+\end{table}
+Пусть на компьютере пользователя имеются следующие таблицы: $P_1$,
+$P_2$, $P_3$, сформированные заранее. Схема таблицы $P_1$=(Номер факультета, Факультет), без ограничений на данные;
+схема таблицы $P_2$=(Номер предмета, Предмет), без ограничений на данные;
+схема таблицы $P_3$=(Номер факультета, Номер предмета, Семестр, Группа, Часы), со следующим ограничением:
+$F_3 =(\mbox{Семестр} \leq 2) \wedge (\mbox{Факультет} = \mbox{Химический})$.
+Пусть при формировании схемы представления данных в таблице \ref{def_T_1} пользователь задал следующее
+ограничение на нее: $F^{\ast} = (\mbox{Семестр} \leq 3)$.
+Сравнение областей истинности формул для ограничений позволит сделать вывод, что $P_1$ и $P_2$ имеют
+достаточно данных для формирования таблицы \ref{def_T_1}, а $P_3$ требует дополнения данных.
+Вычисление областей истинности формул $F^{\ast}$ и $F_3$ позволит сформировать новую формулу:
+$F=((\mbox{Семестр} = 3) \wedge (\mbox{Факультет=Химический}))
+\vee ((\mbox{Семестр} \leq 3) \wedge (\mbox{Факультет} \neq \mbox{Химический}))$. Использую схему $P_3$ и
+формулу $F$ формируется команда SQL, которая передается на сервер для исполнения. В конечном счете,
+по сети передается меньше данных.
 
@@
 
-Будем рассматривать логические формулы в дизъюнктивной нормальной форме. Тогда
-формула $F$ имеет вид
+Для упрощения вычисления областей истинности будем рассматривать логические формулы в дизъюнктивной
+нормальной форме. В общем случае формула $F$ имеет вид
 
 \begin{equation}
-F = K_1 \vee K_2 \vee \dots \vee K_m
+F = K_1 \vee K_2 \vee \dots \vee K_m ,
 \label{def_F_1}
@@
 \begin{equation}
-K_i = T_1 \& T_2 \dots \& T_n, i = 1, \dots, m
+K_i = T_1 \& T_2 \dots \& T_n, i = 1, \dots, m ,
 \label{def_F_2}
 \end{equation}
-
-$T_j, j = 1, \dots, n$ - предикаты, в которых явным образом
+где $T_j, j = 1, \dots, n$ - предикаты, в которых явным образом
 специфицированы расширенные имена атрибутов $R_i.A_j$ (атрибут $A_j$ в отношении
@@
 \begin{note}
-Здесь и далее будем полагать, что в формулах нет тривиальных условий на
-атрибуты, то есть условий вида $Expr_1 = Expr_1$ и сводящихся к ним. Такие
+Здесь и далее будем полагать, что в формулах отсутствуют тривиальные условия на
+атрибуты, например, условия вида $Expr_1 = Expr_1$ и сводящихся к ним.
+В общем случае будем полагать, что множество возможных значений атрибута $A_l$
+в отношении $R_i$ не включается целиком в область истинности предиката $T_j (\dots , R_i . A_l ,\dots )$. Такие
 условия могут быть удалены из формулы, не меняя области истинности (будет
@@
 результаты вычисления логических выражений в SQL-запросах соответствуют
-трехзначной логике. Для избежания такой ситуации формулы преобразуются
-определенным образом, изложенным в статье \cite{zyk_pol}.
+трехзначной логике. Это приводит к неоднозначной интерпретации результата не только
+обычными пользователями, но и опытными программистами. Для решения этой проблемы
+предлагается ограничение: каждому атрибуту, входящему в $F^{\ast}$, явно присваивается признак
+<<Использование неопределенного значения>> с двумя взаимоисключающими значениями <<Да>> или <<Нет>>.
+Семантика этого признака такова, что если ему присвоено значение <<Да>>, то появление значения $\n{NULL}$
+для указанного атрибута в текущем кортеже $t$ не служит основанием удаления этого кортежа из
+дальнейшего рассмотрения. В противном случае значение признака <<Нет>> гарантирует, что появление
+значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$ приведет к удалению этого кортежа из
+дальнейшего рассмотрения.
 
-В дальнейшем будем использовать множество $\mathcal{A} =$ $\{(a_1, \dots, a_l)
-\mid a_i \in Dom(A_i), i=1,\dots,l\}$, $l$-мерное пространство значений всех
-атрибутов базы данных.
+Запишем формулу (\ref{def_F_1}) для $F$ в следующем виде: $ F ( \dots, T_j , \dots )$, где $T_j$ - предикаты
+формулы (\ref{def_F_2}). Тогда после предложенного преобразования она примет следующий вид:
+$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, где  $\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - конъюнкция
+по всем атрибутам формулы $F$, для которых не допустимо значение $\n{NULL}$.
+Операция $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, где  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - дизъюнкция
+по всем атрибутам предиката $T_j$, для которых допустимо значение $\n{NULL}$. Внешние скобки для предиката
+$T'_j$ определяют приоритет выполнения операций. Несложно убедиться, что в рамках трехзначной логики
+преобразованная формула принимает только значения $\n{TRUE}$ и $\n{FALSE}$. Кроме того, несложно убедиться,
+что в рамках двузначной логики, когда в кортежах отсутствуют неопределенные значения, исходная формула $F$
+будет эквивалентна преобразованной формуле, поэтому семантика представления $P$ практически не искажается.
+Для раскрытия термина "практически" рассмотрим наиболее неудобный пример: пусть $F = R_1 .A_2 > 3 \vee R_3 .A_4 < 4$
+и на кортеже $t$ атрибут $R_1 .A_2$ принимает допустимое значение $\n{NULL}$, а предикат $R_3 .A_4 < 4$ принимает
+значение $\n{FALSE}$. Тогда преобразование формулы $F$ на кортеже $t$ будет иметь значение $\n{TRUE}$,
+что не совсем очевидно. Далее будем предполагать, что все формулы $F$ являются преобразованными.
+
+Введем в рассмотрение множество $\mathcal{A} =$ $\{(a_1, \dots, a_n)
+\mid a_i \in Dom(A_i), i=1,\dots,n\}$, где $Dom(A_i)$ - множество всех допустимых значений атрибута $A_i$.
+Декартово прозведение $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ - $n$-мерное пространство
+значений всех атрибутов базы данных. Текущее состояние базы данных соответствует точке в этом пространстве.
+Допустимые состояния образуют некоторую область в этом пространстве, соответствующую ограничениям целостности на данные.
+В замечании \ref{trivial} речь идет о всем пространстве состояний, а не только допустимых состояниях.
 
@@
 Сложность компонентов предикатов $\n{Expr}$, $S$ и $\n{Str}$ определяется
-возможностями программного обеспечения по вычислению областей определения
-$\n{Dom} (F_v)$, что необходимо для вычисления $TJ(C_s)$ из промежуточных
-представлений $P_v$.
+возможностями программного обеспечения по вычислению областей истинности
+$F$, как это показано в примере 1.
 \end{note}
@@
 В соответствии с данными определениями легко понять, как будут устроены операции
-над областями, соответствующими формулам. $\n{Dom} (F)$ для некоторой формулы
-$F$, заданной своей ДНФ, является объединением областей, представленных
-отдельными конъюнктами формулы. Область $M(F)$ является подобластью области
-$\n{Dom} (F)$, то есть $M(F) \subseteq \n{Dom} (F)$. Включение $M(F_1) \subseteq
-M(F_2)$ означает, что область истинности каждого конъюнкта $F_1$ является
-подобластью некоторого объединения областей истинности конъюнктов $F_2$.
+над областями истинности, соответствующими формулам. $M (F)$ для некоторой формулы
+$F$, заданной своей ДНФ, является объединением областей истинности, представленных
+отдельными конъюнктами формулы. Область истинности каждого конъюнкта определяется
+определяется как пересечение областей истинности предикатов, входящих в него.
 
-Теперь введем определения, касающиеся модификации вхождения атрибутов в
+Далее введем определения, касающиеся модификации вхождения атрибутов в
 логические формулы.
@@
 $= X$, в которой все термы, содержащие
-атрибуты $R_i^F.A_i^F \notin X$, заменены на тривиальный терм TRUE.
+атрибуты $R_i^F.A_i^F \notin X$, заменены на тривиальный терм $\n{TRUE}$.
 \label{projection}
@@
 \begin{proof}
-Рассмотрим некоторую точку $a \in M(F)$. 
-Так как формула находится в ДНФ, то замена некоторого предиката $T$ на значение
-$\n{TRUE}$, очевидно, может либо полностью сделать значение формулы $\n{TRUE}$,
-если предикат $T$ единственный в конъюнкте $K_i = T$, либо сократить количество
-условий в нём, если $K_i = T \& T'_1 \dots \& T'_n$. В любом из этих случаев
-количество условий в формуле $F[X]$ лишь уменьшается, что гарантирует выполнение
-$F[X](a) = \n{TRUE}$ и, следовательно, $a \in M(F[X])$
+Если $X = \<F> $, то $F = F[X]$ и $M(F) = M(F[X])$. Пусть $X \subset \<F> $. Рассмотрим
+произвольную точку $a \in M(F)$, то есть $F(a)=\n{TRUE}$. Преобразование формулы $F$
+к виду $F[X]$ выполняется заменой предикатов $T_j$ значением $\n{TRUE}$, если они содержат
+атрибут $A_j \in \<F> $ и $A_j \notin X$. Учитывая соотношения  (\ref{def_F_1}) и (\ref{def_F_2}),
+получим $F([X])=\n{TRUE}$. Следовательно, $a \in M(F[X])$, что и требовалось показать.
 \end{proof}
@@
 Пусть это условие не выполняется. Тогда $U' = \<F_2> \setminus \<F_1> \neq
-\emptyset$. Выберем точку $a = (a_1, \dots, a_l) \in M(F_1)$. Для некоторого
+\emptyset$. Выберем точку $a = (a_1, \dots, a_n) \in M(F_1)$. Для некоторого
 атрибута $A_i \in U'$ выберем соответствующую координату $a_i \in Dom(A_i)$. Так
@@
 
+Рассмотренные свойства логических ограничений, заданных на состояниях базы данных,
+в дальнейшем используются при формировании нового представления данных из
+существующих сохраненных представлений.
+
 \section{Исследование свойств промежуточных представлений данных}
+
+Сохраненные промежуточные представления данных обозначим $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$,
+где $P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
+$s(v)$ – количество отношений в базе данных, использованных при формировании представления $P_v$,
+$\pi_{X_v}$ - операция проекции по множеству атрибутов $X_v$, $\sigma_{F_v}$ - операция селекции
+с логическим ограничением на кортежи $F_v$.
+Целевое выражение, которое надо будет получить из представлений $P$, запишем в виде:
+$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l )$$
+
+Рассмотрим проблему формирования представления данных $P^{\ast}$ из существующих промежуточных
+представлений $P_v$.
 
@@
 
+Предложенные в теореме условия гарантируют, что данные, необходимые для формирования
+представления $P^{\ast}$ содержатся в промежуточном представлении $P_v$. Однако, в нем
+могут быть лишние кортежи, которые дают значение $\n{TRUE}$ при подстановке в формулу
+$F^{\ast}$. Дело в том, что эти кортежи будут удалены при выполнении операции естественного
+соединения с отношениями, которых не хватает в множестве $R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} $
+для совпадения с множеством $R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l $.
+Используя области области истинности логических выражений у СУБД можно запросить
+минимально необходимый набор данных для определения лишних кортежей.
+
+Следующая теорема соответствует частному случаю, однако проблема лишних кортежей не возникает.
+
 \input{theorem_2}
 
-Эти результаты можно обобщить, взяв в рассмотрение целый набор промежуточных
-представлений $P_v$. Но сперва приведем один факт, связанный со свойствами
-естественного соединения.
+Для формирования результирующего представления возможно использование
+не одного, а нескольких промежуточных представлений $P_v$. Для этого
+рассмотрим достаточно очевидное свойство операции естественного соединения.
 
 \begin{statement}
-Пусть $\Re_1 = R_1 \bowtie \dots \bowtie R_k$ - результат естественного
-соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \bowtie \dots
-\bowtie R_k \bowtie R_{k+1} \bowtie \dots \bowtie R_{n}$. 
-Тогда $\Re_2 [\langle \bowtie_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
+Пусть $\Re_1 = R_1 \Join \dots \Join R_k$ - результат естественного
+соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \Join \dots
+\Join R_k \Join R_{k+1} \Join \dots \Join R_{n}$. 
+Тогда $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
 \label{join_property}
@@
 
-Это действительно так, ведь после проведения операции естественного соединения
+Это действительно так, поскольку после проведения операции естественного соединения
 с дополнительными отношениями исходные кортежи, содержавшиеся в $\Re_1$, могут
 лишь ''отсеяться'' операцией естественного соединения. Тогда, взяв вырезку по
-исходным атрибутам, получим как максимум то же самое отношение $\Re_1$.
+исходным атрибутам, получим как максимум то же самое отношение $\Re_1$. Поскольку
+операция естественного соединения коммутативна, то порядок присоединения отношений
+не имеет значения.
 
@@
 
+Как и в предыдущем случае существуют условия, при которых целевое представление
+данных можно вычислить точно.
+
 \input{theorem_2_mult}
 
-\section{Удаление лишних кортежей из промежуточных представлений}
+Выполнение условий рассмотренных теорем не представляет собой  исключительный случай,
+поскольку набор размерностей при формировании гиперкубов изменяется редко, что позволит
+формировать эти размерности из промежуточных представлений без обращения к СУБД.
 
-Результаты, полученные в теоремах \ref{th_base}, \ref{th_mult} нельзя сразу
-использовать на практике. Отношение $P^{\ast}$ содержится в имеющемся
-представлении, однако неизвестно, какие именно кортежи являются частью
-$P^{\ast}$. Лишние кортежи могут возникнуть по двум причинам:
-\begin{enumerate}
-  \item В силу Утверждения \ref{proj_property_of_inclusion}, так как правая
-  часть содержит проекцию исходной формулы $F^{\ast}$ на некоторое множество
-  атрибутов.
-  \item В силу Утверждения \ref{join_property}, так как в промежуточном
-  представлении отношений не больше, чем в искомом $P^{\ast}$.
-\end{enumerate}
-Поэтому необходимо сперва удалить лишние кортежи (если они есть) из
-промежуточных представлений.
+\section{Заключение}
 
-\input{alg_filter_tj}
+Основной результат данной работы состоит в обосновании условий использования
+зарезервированных представлений данных для формирования новых гиперкубов.
+Эти представления данных могут храниться на компьютере пользователя-аналитика
+и существенно сократить время на формирование данных, необходимых для принятия решений.
+
+Традиционная проблема в системах, работающих не под управлением СУБД, заключается
+в актуализации данных, полученных из БД. В технологии MOLAP она решается путем периодической
+актуализации содержимого гиперкуба службой сопровождения. Аналогичный способ может быть
+применен для актуализации представлений $P_v$. Для сокращения времени актуализации возможно
+получить доступ к журналу изменений, который сопровождает СУБД, и актуализировать только те
+представления, для которых изменились исходные данные в БД. Однако еще раз отметим, что данные,
+используемые в размерностях гиперкуба, обычно редко модифицируются в БД.
+
+Предлагаемый подход подразумевает в дальнейшем применение графических процессоров
+для выполнения промежуточных операций фильтрации и соединения различных представлений данных.
+Исследования в этой области показали, что операции на графическом процессоре выполняются
+существенно быстрее, чем на центральном процессоре, для небольших объемов данных. При увеличении
+объемов производительность падает. Причина очевидна: пока данные помещаются в быструю память
+графического процессора, операции с ними выполняются быстро. При увеличении объемов данных появляется
+необходимость их многократного перемещения между различными видами памяти, что снижает
+производительность. С одной стороны, производители графических процессоров постоянно наращивают
+быструю память, с другой – повторное использование данных наиболее актуально для размерностей
+многомерных таблиц, для которых требуется сравнительно небольшой объем памяти, что в совокупности
+делает целесообразным применение графических процессоров при решении указанной проблемы.
 

### 2014-02-28 bibliography.tex
zykin changes and my corrections

@@
   – P. 47–56.
+%ours  
+\bibitem{zyk_pol}
+  \textit{С.В. Зыкин, А.Н. Полуянов}
+  Формирование многомерных данных с использованием промежуточных представлений.
+  "Проблемы управления".
+  2013.
+  № 5. Стр. 54-59.
+%stream
+\bibitem{Olston03}
+  \textit{Olston C., Jiang J., Widom J.}
+  Adaptive filters for continuous queries over distributed data streams
+  // Proc. of the 2003 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '03).
+  – San Diego, 2003.
+  – P. 563–574.
+\bibitem{Denny05}
+  \textit{Denny M., Franklin M.J.}
+  Predicate result range caching for continuous queries
+  // Proc. of the 2005 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '05).
+  – N.-Y., 2005.
+  – P. 646–657.
+%logic
+\bibitem{Afrati06}
+  \textit{Afrati F., Li C., Mitra P.}
+  Rewriting queries using views in the presence of arithmetic comparisons
+  // Theoretical Computer Science.
+  – 2006.
+ – Vol. 368, N 1–2. – P. 88–123.
 %static approach
@@
   - Vol. 68, - P. 103–119
-
-%ours  
-\bibitem{zyk_pol}
-  \textit{С.В. Зыкин, А.Н. Полуянов}
-  Формирование многомерных данных с использованием промежуточных представлений.
-  "Проблемы управления".
-  2013.
-  № 5. Стр. 54-59.
+ %domain
+\bibitem{Keller96}
+  \textit{Keller M., Basu J.}
+  A predicate-based caching scheme for client-server database architectures
+  // VLDB Journal. – 1996. N 5. – P. 35–47.
 

### 2014-02-28 theorem_1.tex
zykin changes and my corrections

@@
 что $t^{\ast} \in \pi_{X^{\ast}} (P_{v})$. Из условия $t^{\ast} \in P^{\ast}$ следует, 
-что существует кортеж $t' \in R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots
-\bowtie R^{\ast}_l$ и $t^{\ast} = t'[X^{\ast}]$. Заметим, что такой кортеж 
+что существует кортеж $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots
+\Join R^{\ast}_l$ и $t^{\ast} = t'[X^{\ast}]$. Заметим, что такой кортеж 
 определяется не единственным образом. Среди возможных вариантов выберем такой, 
@@
 естественного соединения образуют кортеж $t''$, для которого выполнено:
-$t'' = t'[\langle {\bowtie}_{i=1}^{s(v)} R^v_i \rangle]$. В силу условия а) и 
-$X_v \subseteq \langle {\bowtie}_{i=1}^{s(v)} R^v_i \rangle$ получаем 
+$t'' = t'[\langle {\Join}_{i=1}^{s(v)} R^v_i \rangle]$. В силу условия а) и 
+$X_v \subseteq \langle {\Join}_{i=1}^{s(v)} R^v_i \rangle$ получаем 
 $t''[X^{\ast}] = t'[X^{\ast}] = t^{\ast}$.

### 2014-03-01 article_main.tex
some minor corrections

@@
 \def \n #1{\mathit{#1}}
-\title{Формирование табличных приложений с использование зарезервированных представлений данных}
+\title{Формирование табличных приложений с использование зарезервированных
+представлений данных}
 \author{
-Мосин С.В., Зыкин С.В. \footnote{Работа выполнена при поддержке гранта РФФИ № 12-07-00066-а}
+Мосин С.В., Зыкин С.В. \footnote{Работа выполнена при поддержке гранта РФФИ №
+12-07-00066-а}
 }
@@
 помощью зарезервированных представлений на компьютере пользователя, оснащенном
-графическим процессором (GPU). При этом не предполагается какое-либо преобразование
-исходной схемы базы данных, в том числе к иерархическому виду.
+графическим процессором (GPU). При этом не предполагается какое-либо
+преобразование исходной схемы базы данных, в том числе к иерархическому виду.
 
@@
 запросов, поскольку нацелена на сокращение объема передаваемых данных с сервера
-базы данных. Зарезервированные данные активно используется в системах управления базами
-данных (СУБД). Но в большинстве случаев это касается повторного использования
-данных, записанных в кеш, без предварительного анализа
+базы данных. Зарезервированные данные активно используется в системах управления
+базами данных (СУБД). Но в большинстве случаев это касается повторного
+использования данных, записанных в кеш, без предварительного анализа
 содержимого на предмет возможности частичного или комбинированного
@@
 области учитываются при вычислении статистических оценок для оптимизации
-физического доступа к базе данных. Близкими по методам решения являются задачи выполнения
-запросов на потоках данных \cite{Olston03, Denny05}, однако различные, по сравнению
-с настоящей работой, цели приводят к различным результатам. Наиболее близка к
-рассматриваемой проблеме работа \cite{Afrati06}. В ней рассматриваются конъюнктивные
-запросы над доменами данных с предикатами в виде арифметических сравнений и
-представлены алгоритмы вычисления запросов с использованием существующих
-представлений. В настоящей работе рассматривается специальный вид универсального
-реляционного запроса над отношениями базы данных, а не над отдельными доменами.
-Хотя цели в обеих работах совпадают, результаты различны по указанной причине.
-В частности, в настоящей работе нет необходимости разрабатывать алгоритмы выборки
-данных из промежуточных представлений, так как их замещает реляционная алгебра.
+физического доступа к базе данных. Близкими по методам решения являются задачи
+выполнения запросов на потоках данных \cite{Olston03, Denny05}, однако
+различные, по сравнению с настоящей работой, цели приводят к различным
+результатам. Наиболее близка к рассматриваемой проблеме работа \cite{Afrati06}.
+В ней рассматриваются конъюнктивные запросы над доменами данных с предикатами
+в виде арифметических сравнений и представлены алгоритмы вычисления запросов с
+использованием существующих представлений. В настоящей работе рассматривается
+специальный вид универсального реляционного запроса над отношениями базы данных,
+а не над отдельными доменами. Хотя цели в обеих работах совпадают, результаты
+различны по указанной причине. В частности, в настоящей работе нет необходимости
+разрабатывать алгоритмы выборки данных из промежуточных представлений, так как
+их замещает реляционная алгебра.
 
@@
 Предложенные в теореме условия гарантируют, что данные, необходимые для формирования
-представления $P^{\ast}$ содержатся в промежуточном представлении $P_v$. Однако, в нем
+представления $P^{\ast}$, содержатся в промежуточном представлении $P_v$. Однако в нем
 могут быть лишние кортежи, которые дают значение $\n{TRUE}$ при подстановке в формулу

### 2014-03-01 bibliography.tex
some minor corrections

--- 2014-01-23 bibliography.tex
@@
 \bibitem{lecht}
-  \textit{Lechtenborger J., Vossen G.}
+  \textit{Lechtenborger J.}, \textit{Vossen G.}
   Multidimensional normal forms for data warehouse design
@@
   – P. 47–56.
+
+%ours  
+\bibitem{zyk_pol}
+  \textit{С.В. Зыкин, А.Н. Полуянов}
+  Формирование многомерных данных с использованием промежуточных представлений.
+  "Проблемы управления".
+  2013.
+  № 5. Стр. 54-59.
+
+%stream
+\bibitem{Olston03}
+  \textit{Olston C., Jiang J., Widom J.}
+  Adaptive filters for continuous queries over distributed data streams
+  // Proc. of the 2003 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '03).
+  – San Diego, 2003.
+  – P. 563–574.
+\bibitem{Denny05}
+  \textit{Denny M., Franklin M.J.}
+  Predicate result range caching for continuous queries
+  // Proc. of the 2005 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '05).
+  – N.-Y., 2005.
+  – P. 646–657.
+
+%logic
+\bibitem{Afrati06}
+  \textit{Afrati F., Li C., Mitra P.}
+  Rewriting queries using views in the presence of arithmetic comparisons
+  // Theoretical Computer Science.
+  – 2006.
+ – Vol. 368, N 1–2. – P. 88–123.
+
 %static approach
@@
   - P. 453–470.
+
 %dynamic
@@
 
-%ours  
-\bibitem{zyk_pol}
-  \textit{С.В. Зыкин, А.Н. Полуянов}
-  Формирование многомерных данных с использованием промежуточных представлений.
-  "Проблемы управления".
-  2013.
-  № 5. Стр. 54-59.
+%domain
+\bibitem{Keller96}
+  \textit{Keller M., Basu J.}
+  A predicate-based caching scheme for client-server database architectures
+  // VLDB Journal. – 1996. N 5. – P. 35–47.
 

### 2014-03-13 english_version/article_main.tex
starting translation

--- 2014-03-01 article_main.tex
@@
 
-\begin{statement}
-$M(F_1) \subseteq M(F_2) \Rightarrow \<F_2> \subseteq \<F_1> $
-\label{attibute_inclusion_property}
-\end{statement}
-\begin{proof}
-Пусть это условие не выполняется. Тогда $U' = \<F_2> \setminus \<F_1> \neq
-\emptyset$. Выберем точку $a = (a_1, \dots, a_n) \in M(F_1)$. Для некоторого
-атрибута $A_i \in U'$ выберем соответствующую координату $a_i \in Dom(A_i)$. Так
-как по замечанию \ref{trivial} в формулах нет тривиальных условий на атрибуты, то $\exists a_i'
-\in Dom(A_i) : a' = (a_1, \dots, a_{i - 1}, a_i', a_{i + 1}, \dots, a_l) \notin
-M(F_2)$. Однако $a' \in M(F_1)$, так как атрибут $A_i$ отсутствует в формуле
-$F_1$. Следовательно, $M(F_1) \nsubseteq M(F_2)$. Полученное противоречие
-доказывает утверждение.
-\end{proof}
-
-\begin{consequence}
-$M(F_1) = M(F_2) \Rightarrow \<F_2> = \<F_1> $
-\label{attibute_equality_property}
-\end{consequence}
-\begin{proof}
-Пусть это условие не выполняется. Тогда, в силу предыдущего утверждения,
-существует множество $U' = \<F_1> \setminus \<F_2> \neq \emptyset$. Аналогично
-предыдущему доказательству, выбираем точку из любого из множеств и меняем
-значение координаты, соответствующей атрибуту из множества $U'$ на такое,
-которое не принаджежит множеству $M(F_1)$. Полученная точка по-прежнему
-принадлежит $M(F_2)$. Мы снова пришли к противоречию. Утверждение доказано.
-\end{proof}
-
-Рассмотренные свойства логических ограничений, заданных на состояниях базы данных,
-в дальнейшем используются при формировании нового представления данных из
-существующих сохраненных представлений.
+Данное свойство логических ограничений, заданных на состояниях базы
+данных, в дальнейшем используется при формировании нового представления данных
+из существующих сохраненных представлений.
 

### 2014-03-13 english_version/theorem_2.tex
starting translation

--- 2014-01-23 theorem_2.tex
@@
 $t_v \in P^{\ast}$.
-Из определения кортежа $t_v$ следует, что существует кортеж  $t' \in R^v_1 \bowtie R^v_2\bowtie\ldots \bowtie R^v_{s(v)}$ и $t_v = t'[X^{\ast}]$, $F_v(t')=\n{TRUE}$.
-По свойству операции естественного соединения $R^v_1 \bowtie R^v_2\bowtie\ldots \bowtie R^v_{s(v)} = R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots \bowtie R^{\ast}_l$, следовательно, $t' \in R^{\ast}_1 \bowtie R^{\ast}_2\bowtie\ldots \bowtie R^{\ast}_l$.
+Из определения кортежа $t_v$ следует, что существует кортеж  $t' \in R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)}$ и $t_v = t'[X^{\ast}]$, $F_v(t')=\n{TRUE}$.
+По свойству операции естественного соединения $R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)} = R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join R^{\ast}_l$, следовательно, $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join R^{\ast}_l$.
 В силу условия в) $F^{\ast} (t')=\n{TRUE}$.

### 2014-03-13 article_main.tex
started very first translation work and corrected russian title

@@
 \def \n #1{\mathit{#1}}
-\title{Формирование табличных приложений с использование зарезервированных
+\title{Формирование табличных приложений с использованием зарезервированных
 представлений данных}

### 2014-03-25 english_version/article_main.tex
started very first translation work and corrected russian title
continuing translation

@@
 \documentclass[10pt,a4paper]{article}
-\author{Мосин Сергей, Зыкин Сергей}
+\author{Mosin Sergey, Zykin Sergey}
 \usepackage{amssymb,amsmath}
 \usepackage[utf8]{inputenc}
-\usepackage[russian]{babel}
 \usepackage{amsthm}
@@
 \usepackage{xcolor}
-\floatname{algorithm}{Алгоритм}
-\newtheorem{theorem}{Теорема}
-\newtheorem{mydef}{Определение}
-\newtheorem{statement}{Утверждение}
-\newtheorem*{consequence}{Следствие}
-\newtheorem{note}{Замечание}
+\floatname{algorithm}{Algorithm}
+\newtheorem{theorem}{Theorem}
+\newtheorem{mydef}{Definition}
+\newtheorem{statement}{Statement}
+\newtheorem*{consequence}{Consequence}
+\newtheorem{note}{Note}
 \def \eval #1#2{\left.#1\right\vert_{#2}}
@@
 \def \n #1{\mathit{#1}}
-\title{Формирование табличных приложений с использование зарезервированных
-представлений данных}
+\title{Creating table-like applications using cached data}
 \author{
-Мосин С.В., Зыкин С.В. \footnote{Работа выполнена при поддержке гранта РФФИ №
-12-07-00066-а}
+Mosin Sergey, Zykin Sergey \footnote{The paper is supported by RFFI grant
+#12-07-00066-a}
+\date{}
 }
-\date{}
-
 \begin{document}
 \maketitle
+\section{Abstract}
 
-\section{Введение}
+\section{Introduction}
+A lot of nowadays Big Data analysis solutions use Online Analytical Processing
+technologies, formed in early 90s already \cite{codd}. A lot of fundamental
+\cite{lecht,lehner,mazon} and applied \cite{vassi, peder, progressive, giorg}
+results were obtained in this field lately.
 
-Многие современные решения в области анализа больших объемов данных
-используют технологии аналитической обработки данных (Online Analytical
-Processing), сформировавшиеся еще в начале 90-х годов \cite{codd}. За последнее
-время в этой области было получено большое количество фундаментальных
-\cite{lecht,lehner,mazon} и практических результатов \cite{vassi, peder,
-progressive, giorg}.
+The purpose of this paper is the research of the problem of creating Dynamic
+Multidimensional Data Representation from the Relational Database with a help of
+cached data representations on a user computer, supplied with Graphical
+Processing Unit (GPU). We don't require any transformation of the initial
+Database schema, even to the hierarchical model.
 
-Целью данной работы, включая результаты этой статьи, является изучение
-проблемы динамического формирования
-многомерного представления данных из реляционной базы данных с
-помощью зарезервированных представлений на компьютере пользователя, оснащенном
-графическим процессором (GPU). При этом не предполагается какое-либо
-преобразование исходной схемы базы данных, в том числе к иерархическому виду.
+This problem is connected with a field of query optimization, because it aims
+the decrease in data transfer from a Database server. Cached data is actively
+used in Database Management Systems (DBMS). But mostly it is only repeating use
+of data, written in cache, without any prior data analysis to define any partial
+or combined use. The DBMS can only avoid requesting blocks of data from external
+devices while serving the query, if they are present(?ed) in cache. It means
+block number analysis, not their information.
 
-Поставленная проблема связана с областью оптимизации
-запросов, поскольку нацелена на сокращение объема передаваемых данных с сервера
-базы данных. Зарезервированные данные активно используется в системах управления
-базами данных (СУБД). Но в большинстве случаев это касается повторного
-использования данных, записанных в кеш, без предварительного анализа
-содержимого на предмет возможности частичного или комбинированного
-использования. Работа СУБД ограничивается тем, что при выполнении очередного
-запроса блоки данных не запрашиваются с внешних устройств, если они есть в
-кеше, т. е. анализируются номера блоков, а не их содержимое.
+This paper is based on the results, obtained in \cite{zyk_pol}. We removed
+condition that restricted(?) Intermediate Representation (IR) attributes and
+also made some generalizations on multiple IRs case.
 
-Данная работа основывается на результатах, полученных в
-статье \cite{zyk_pol}. Устранено условие, ограничивающее атрибуты промежуточного
-представления, а также сделано обобщение теорем на случай нескольких
-промежуточных представлений.
+\section{Existing solutions overview and comparison}
 
-\section{Обзор публикаций и сравнение}
+In a field of fundamental researches huge amount of publications aim optimal
+query plan generation based on a formal set of rules, that don't use SQL operator
+predicate domains (logical optimization). In other cases those domains are
+taken into account for statistical evaluations to improve physical access to the
+data. Stream data query executing problems, described in
+\cite{Olston03, Denny05}, are related to our problem closer, but different aims
+lead us to different results. The most similar to our work is paper
+\cite{Afrati06}. They analyze conjunctive queries on data domains with
+predicates in a form of arithmetical comparisons and present query computation
+algorithms using IRs. In this paper the special case of universal relational
+query is considered. It is a query on the Database relations, not particular
+domains. Although we had similar aims, the results obtained differ because of
+stated factors. In particular, in our work there is no need to create any
+algorithms of data selection, as they are replaced by Relational Algebra.
 
-В области теоретических исследований ситуация складывается следующим образом.
-Огромное число публикаций посвящено проблеме построения оптимального плана
-запроса на основе формальных правил, в которых не используются области
-определения предикатов в SQL-операторах (логическая оптимизация) либо эти
-области учитываются при вычислении статистических оценок для оптимизации
-физического доступа к базе данных. Близкими по методам решения являются задачи
-выполнения запросов на потоках данных \cite{Olston03, Denny05}, однако
-различные, по сравнению с настоящей работой, цели приводят к различным
-результатам. Наиболее близка к рассматриваемой проблеме работа \cite{Afrati06}.
-В ней рассматриваются конъюнктивные запросы над доменами данных с предикатами
-в виде арифметических сравнений и представлены алгоритмы вычисления запросов с
-использованием существующих представлений. В настоящей работе рассматривается
-специальный вид универсального реляционного запроса над отношениями базы данных,
-а не над отдельными доменами. Хотя цели в обеих работах совпадают, результаты
-различны по указанной причине. В частности, в настоящей работе нет необходимости
-разрабатывать алгоритмы выборки данных из промежуточных представлений, так как
-их замещает реляционная алгебра.
+There are some works, which examine query optimization problem in the case of
+multidimensional warehouse representation with a use of dynamically formed
+representations. Two basic methods are used there: statical \cite{baralis,
+gupta, gupta-mumick} and dynamical one \cite{scheuermann, shim, kalnis, chang}.
+The first one is based on a fixed amount of queries. In the latter one queries
+for reservation are chosen based on appearance statistic and computational cost
+of query execution. As a data source Data Warehouses or Relational Databases in
+the hierarchical form are used. Special queries for data manipulation are
+considered. In this paper we use split dimension generation for the data
+required for current analysis. Hence, we are interested in standard SQL queries
+to the Database and IRs. The resulting representation can then be processed by
+special multidimensional data queries.
 
-Существует ряд работ, в которых рассматривается проблема оптимизации
-запросов к хранилищам данных, имеющих концептуальное многомерное представление,
-с использованием динамически формируемых представлений. При этом
-существуют два основных подхода для резервирования результатов запросов:
-статический \cite{baralis, gupta, gupta-mumick} и динамический \cite{scheuermann, shim,
-kalnis, chang}. Первый базируется на использовании набора фиксированных запросов,
-во втором же предполагается динамический выбор результатов запросов для резервирования на
-основе статистики появления а также вычислительной стоимости выполнения
-запросов. В качестве источника данных используются хранилища данных, либо
-реляционная база данных, преобразованная к иерархическому виду. При этом,
-рассматриваются специальные запросы манипулирования многомерными данными. 
-В этой статье рассматривается технология раздельного формирования размерностей
-представления данных, необходимого для текущего анализа. Поэтому интерес
-представляют стандартные SQL запросы к базе данных и промежуточным представлениям. 
-Результирующее представление может быть затем обработано специальными запросами
-манипулирования многомерными данными.
+In the \cite{Keller96} paper they claim for cache content analysis, and
+predicates correspond to IRs in our approach. Cached data usage problem is solved
+in terms of predicate deduction. In our paper we solve this problem by means of
+computation of cached query results domains that correspond to the IRs. It allows
+us to define IR records that can be used to form new representation and new SQL
+queries that will let us load missing data from Database server. The following
+example demonstrates suggested approach.
 
-В работе \cite{Keller96} решается задача анализа содержимого кеша,
-при этом промежуточным представлениям
-в кеше ставятся в соответствие предикаты. Проблема использования представлений
-решается за счет выводимости предикатов. В данной работе проблему использования
-промежуточных представлений предлагается решать за счет вычисления областей
-определения (доменов) результатов выполнения запросов, находящихся в кеше,
-соответствующих промежуточным представлениям. Это дает возможность
-определить кортежи, которые можно использовать в зарезервированных представлениях
-при формировании нового представления, и на основе вычисления областей определения
-сформировать SQL команды, которые позволят загрузить недостающие данные с сервера
-базы данных. Для демонстрации предлагаемых решений рассмотрим пример.
-
-{\bf Пример 1.} Допустим, что требуется получить представление данных, как показано в
-таблице \ref{def_T_1}. Жирным шрифтом в этой таблице представлены атрибуты размерностей,
-курсивом - атрибуты мер, значения атрибутов - обычным шрифтом.
+{\bf Example 1.} The task is to receive data representation as in the table
+\ref{def_T_1}. Dimension attributes are in bold, measure attributes are italic,
+else are attribute values.
 
 \begin{table}[h!]
-\caption{\label{def_T_1} Фрагмент учебного плана.}
+\caption{\label{def_T_1} Curriculum.}
 \begin{center}
@@
 \cline{2-6}
-\multicolumn{1}{c|}{} & \multicolumn{1}{|r||}{\bf Семестр} & \multicolumn{2}{|c|}{2} & \multicolumn{2}{|c|}{3}  \\
+\multicolumn{1}{c|}{} & \multicolumn{1}{|r||}{\bf Semester} & \multicolumn{2}{|c|}{2} & \multicolumn{2}{|c|}{3}  \\
 \cline{2-6}
-\multicolumn{1}{c|}{} & {\bf Предмет} & Физика & Химия & Физика & Химия \\
+\multicolumn{1}{c|}{} & {\bf Discipline} & Physics & Chemistry & Physics & Chemistry \\
 \hline \cline{2-6}
-{\bf Факультет} & {\bf Группа} & {\it Часы} & {\it Часы} & {\it Часы} & {\it Часы} \\
+{\bf Department} & {\bf Class} & {\it Hours} & {\it Hours} & {\it Hours} & {\it Hours} \\
 \hline
- & Х-321 & & 36 & 18 & 9 \\
+ & C-321 & & 36 & 18 & 9 \\
 \cline{2-6}
-Химический & Х-322 & & 48 & 18 & 9 \\
+D. of Chemistry & Х-322 & & 48 & 18 & 9 \\
 \cline{2-6}
- & Х-331 & & 24 & 18 & 9 \\
+ & C-331 & & 24 & 18 & 9 \\
 \hline
- & Ф-321 & 48 & & 24 & \\
+ & P-321 & 48 & & 24 & \\
 \cline{2-6}
-Физический & Ф-322 & 48 & & 24 & \\
+D. of Physics & P-322 & 48 & & 24 & \\
 \cline{2-6}
- & Ф-331 & 64 & & 24 & \\
+ & P-331 & 64 & & 24 & \\
 \hline
@@
 \end{table}
-Пусть на компьютере пользователя имеются следующие таблицы: $P_1$,
-$P_2$, $P_3$, сформированные заранее. Схема таблицы $P_1$=(Номер факультета, Факультет), без ограничений на данные;
-схема таблицы $P_2$=(Номер предмета, Предмет), без ограничений на данные;
-схема таблицы $P_3$=(Номер факультета, Номер предмета, Семестр, Группа, Часы), со следующим ограничением:
-$F_3 =(\mbox{Семестр} \leq 2) \wedge (\mbox{Факультет} = \mbox{Химический})$.
-Пусть при формировании схемы представления данных в таблице \ref{def_T_1} пользователь задал следующее
-ограничение на нее: $F^{\ast} = (\mbox{Семестр} \leq 3)$.
-Сравнение областей истинности формул для ограничений позволит сделать вывод, что $P_1$ и $P_2$ имеют
-достаточно данных для формирования таблицы \ref{def_T_1}, а $P_3$ требует дополнения данных.
-Вычисление областей истинности формул $F^{\ast}$ и $F_3$ позволит сформировать новую формулу:
-$F=((\mbox{Семестр} = 3) \wedge (\mbox{Факультет=Химический}))
-\vee ((\mbox{Семестр} \leq 3) \wedge (\mbox{Факультет} \neq \mbox{Химический}))$. Использую схему $P_3$ и
-формулу $F$ формируется команда SQL, которая передается на сервер для исполнения. В конечном счете,
-по сети передается меньше данных.
+Assume the following precomputed relations on the user PC:  $P_1$,
+$P_2$, $P_3$.
+Table schemes are the following then: $P_1$=(Department number, Department), without any data constraints; $P_2$=(Discipline number, Discipline), without data constraints; $P_3$=(Department number, Discipline number, Semester, Class, Hours), with the following constraint:
+$F_3 =(\mbox{Semester} \leq 2) \wedge (\mbox{Department} = \mbox{D. of Chemistry})$.
+Assume the user has made the following logical constraint on the data to be represented in the table: $F^{\ast} = (\mbox{Semester} \leq 3)$.
+Comparing these formulas domains, it's easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1}, whereas $P_3$ needs to be expanded with data.
+After defining $F^{\ast}$ and $F_3$ domains we can form a new formula: $F=((\mbox{Semester} = 3) \wedge (\mbox{Deparmtent=D.
+of Chemistry}))
+\vee ((\mbox{Semester} \leq 3) \wedge (\mbox{Deparmtent} \neq \mbox{D.
+of Chemistry}))$.
+Using $P_3$ schema and $F$ formula, we can now make an SQL command to be executed on server.
+As a result, we transfer less data over the network.
 

### 2014-04-12 english_version/article_main.tex
translation of main file and bibliography complete

@@
 \def \n #1{\mathit{#1}}
-\title{Creating table-like applications using cached data}
+\title{Building table-like applications using cached data}
 \author{
-Mosin Sergey, Zykin Sergey \footnote{The paper is supported by RFFI grant
+Mosin Sergey, Zykin Sergey \footnote{The paper is supported by RFBR grant
 #12-07-00066-a}
@@
 \maketitle
+
 \section{Abstract}
+We primarily address the usage of cached Database query results in our paper.
+This problem was already discussed and has some software implementations. In our
+work we propose logical formulas Truth spaces usage while querying Database.
+This approach is supposed to help an analyst, who works with dynamically built
+data representations logically described as multidimensional tables. Those
+representations don't take much disk space, so one can use GPU to speed up the
+computations. We consider queries in a form of relational algebra expressions,
+similar to "universal relational query". Logical formulas in the queries are
+presented in Disjunctive Normal Form. They consist of atomic expressions
+consisting of SQL predicates. In the paper we consider single and multiple table
+cases to build new data representations.
 
@@
 
-The purpose of this paper is the research of the problem of creating Dynamic
-Multidimensional Data Representation from the Relational Database with a help of
-cached data representations on a user computer, supplied with Graphical
+The purpose of this paper is to study problem of building Dynamic
+Multidimensional Data Representation from the Relational Database using
+cached data representations on a user computer, equipped with a Graphical
 Processing Unit (GPU). We don't require any transformation of the initial
@@
 This paper is based on the results, obtained in \cite{zyk_pol}. We removed
-condition that restricted(?) Intermediate Representation (IR) attributes and
-also made some generalizations on multiple IRs case.
+condition that limited Intermediate Data Representation (IDR) attributes and
+also made some generalizations on multiple IDRs case.
 
@@
 In a field of fundamental researches huge amount of publications aim optimal
-query plan generation based on a formal set of rules, that don't use SQL operator
-predicate domains (logical optimization). In other cases those domains are
-taken into account for statistical evaluations to improve physical access to the
-data. Stream data query executing problems, described in
-\cite{Olston03, Denny05}, are related to our problem closer, but different aims
-lead us to different results. The most similar to our work is paper
-\cite{Afrati06}. They analyze conjunctive queries on data domains with
-predicates in a form of arithmetical comparisons and present query computation
-algorithms using IRs. In this paper the special case of universal relational
-query is considered. It is a query on the Database relations, not particular
-domains. Although we had similar aims, the results obtained differ because of
-stated factors. In particular, in our work there is no need to create any
-algorithms of data selection, as they are replaced by Relational Algebra.
+query plan generation based on a formal set of rules, that don't use SQL
+operator predicate domains (logical optimization). In other cases those domains
+are taken into account for statistical evaluations to improve physical access to
+the data. Stream data query executing problems, described in \cite{Olston03,
+Denny05}, are related to our problem closer, but different aims lead us to
+different results. The most similar to our work is paper \cite{Afrati06}. They
+analyze conjunctive queries on data domains with predicates in a form of
+arithmetical comparisons and present query computation algorithms using IDRs. In
+this paper the special case of universal relational query is considered. It is a
+query on the Database relations, not particular domains. Although we had similar
+aims, the results obtained differ because of stated factors. In particular, in
+our work there is no need to create any algorithms of data selection, as they
+are replaced by Relational Algebra.
 
@@
 required for current analysis. Hence, we are interested in standard SQL queries
-to the Database and IRs. The resulting representation can then be processed by
+to the Database and IDRs. The resulting representation can then be processed by
 special multidimensional data queries.
@@
 In the \cite{Keller96} paper they claim for cache content analysis, and
-predicates correspond to IRs in our approach. Cached data usage problem is solved
+predicates correspond to IDRs in our approach. Cached data usage problem is solved
 in terms of predicate deduction. In our paper we solve this problem by means of
-computation of cached query results domains that correspond to the IRs. It allows
+computation of cached query results domains that correspond to the IDRs. It allows
 us to define IR records that can be used to form new representation and new SQL
@@
 \cline{2-6}
-D. of Chemistry & Х-322 & & 48 & 18 & 9 \\
+D. of Chemistry & C-322 & & 48 & 18 & 9 \\
 \cline{2-6}
@@
 
-\section{Определение и свойства логических ограничений}
+\section{Logical constraint definition and properties}
 
-Для упрощения вычисления областей истинности будем рассматривать логические формулы в дизъюнктивной
-нормальной форме. В общем случае формула $F$ имеет вид
+To simplify domain computations we will consider logical formulas in Disjunctive Normal Form (DNF).
+In general case formula $F$ has the following form:
 
@@
 \end{equation}
-где $T_j, j = 1, \dots, n$ - предикаты, в которых явным образом
-специфицированы расширенные имена атрибутов $R_i.A_j$ (атрибут $A_j$ в отношении
-$R_i$):
+here $T_j, j = 1, \dots, n$ - predicates, where expanded attribute names are specified. $R_i.A_j$ means attribute $A_j$ in relation $R_i$.
+Those predicates can be:
 \begin{itemize}
-  \item	операция сравнения $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ – операция
-  сравнения $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$ –
-  согласованные по типам допустимые выражения, определенные на множестве
-  расширенных имен атрибутов и констант;
-  \item операция $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
-  \n{Expr}_3$ (содержимое в прямоугольных скобках $[*]$ для предиката не
-  является обязательным при написании);
-  \item операция $\n{Expr} \n{[NOT]} \n{IN} S$, где $S$ – список значений либо
-  подзапрос, результатом которого является столбец атрибута $A_j$ в отношении
-  $R_i$;
-  \item операция $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, где $\n{Str}_i$ –
-  строки;
-  \item операция $\n{Expr} \theta \n{ALL/ANY} S$.
+  \item	comparison operation $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ –
+  comparison operator $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$
+  – type conformant expressions, defined in a space of expanded attribute names and constants;
+  \item operation $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
+  \n{Expr}_3$ (symbols inside square brackets $[*]$ aren't mandatory for the predicate);
+  \item operation $\n{Expr} \n{[NOT]} \n{IN} S$, here $S$ – values list or subquery, having a set of attribute $R_i.A_j$ values as a result;
+  \item operation $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, here $\n{Str}_i$ –
+  strings;
+  \item operation $\n{Expr} \theta \n{ALL/ANY} S$.
 \end{itemize}
@@
 \begin{note}
-Здесь и далее будем полагать, что в формулах отсутствуют тривиальные условия на
-атрибуты, например, условия вида $Expr_1 = Expr_1$ и сводящихся к ним.
-В общем случае будем полагать, что множество возможных значений атрибута $A_l$
-в отношении $R_i$ не включается целиком в область истинности предиката $T_j (\dots , R_i . A_l ,\dots )$. Такие
-условия могут быть удалены из формулы, не меняя области истинности (будет
-определено далее).
+We assume logical formulas have no trivial conditions on attributes, for example, $Expr_1 = Expr_1$ and those reduced to such form.
+In general, we assume that $A_l$ in $R_i$ domain is not fully contained in $T_j (\dots , R_i .
+A_l ,\dots )$ predicate truth space.
+Such conditions can be removed from the formula without a change in the truth space (we will define it later).
 \label{trivial}
@@
 \begin{mydef}
-Множество атрибутов, входящих в формулу, выражает размерность формулы и
-обозначается $\<F> $.
+Space of attributes contained in a formula show the dimension of formula and is denoted as $\<F> $.
 
@@
 
-Перечисленные варианты операций используют не все возможности языка SQL.
-Например, предикат $\n{EXISTS}$ не используется, поскольку в нем явно не
-специфицированы расширенные имена атрибутов, предикат $\n{NULL}$ используется
-в данной работе для другой цели.
+Listed variants of operations don't use every single SQL capability.
+For example, we don't use $\n{EXISTS}$ predicate, because it has no expanded attribute names in it.
+$\n{NULL}$ predicate is used for another purpose in our paper.
 
-При вычислении логического выражения может быть получено значение $\n{UNKNOWN}$,
-если на текущем кортеже $t$ атрибут принимает значение $\n{NULL}$, поскольку
-результаты вычисления логических выражений в SQL-запросах соответствуют
-трехзначной логике. Это приводит к неоднозначной интерпретации результата не только
-обычными пользователями, но и опытными программистами. Для решения этой проблемы
-предлагается ограничение: каждому атрибуту, входящему в $F^{\ast}$, явно присваивается признак
-<<Использование неопределенного значения>> с двумя взаимоисключающими значениями <<Да>> или <<Нет>>.
-Семантика этого признака такова, что если ему присвоено значение <<Да>>, то появление значения $\n{NULL}$
-для указанного атрибута в текущем кортеже $t$ не служит основанием удаления этого кортежа из
-дальнейшего рассмотрения. В противном случае значение признака <<Нет>> гарантирует, что появление
-значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$ приведет к удалению этого кортежа из
-дальнейшего рассмотрения.
+During logical formula domain calculation if we have some attribute having $NULL$ value on a row $t$, we then get the $\n{UNKNOWN}$ value for the whole formula, because SQL-query results follow Three-valued logic.
+It leads us to ambiguous interpretations of result of both usual users and experienced programmers.
+To solve this problem we suggest the following constraint: every attribute in $F^{\ast}$ is supplied with a property <<Use of undefined value>> with two mutually exclusive values: <<Yes>> or <<No>>.
+The motivation behind this property is the following: if it is assigned <<Yes>>, then we leave rows with $\n{NULL}$ value for further consideration.
+Otherwise, having <<No>> in that property guarantees us removing all such rows.
 
-Запишем формулу (\ref{def_F_1}) для $F$ в следующем виде: $ F ( \dots, T_j , \dots )$, где $T_j$ - предикаты
-формулы (\ref{def_F_2}). Тогда после предложенного преобразования она примет следующий вид:
-$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, где  $\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - конъюнкция
-по всем атрибутам формулы $F$, для которых не допустимо значение $\n{NULL}$.
-Операция $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, где  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - дизъюнкция
-по всем атрибутам предиката $T_j$, для которых допустимо значение $\n{NULL}$. Внешние скобки для предиката
-$T'_j$ определяют приоритет выполнения операций. Несложно убедиться, что в рамках трехзначной логики
-преобразованная формула принимает только значения $\n{TRUE}$ и $\n{FALSE}$. Кроме того, несложно убедиться,
-что в рамках двузначной логики, когда в кортежах отсутствуют неопределенные значения, исходная формула $F$
-будет эквивалентна преобразованной формуле, поэтому семантика представления $P$ практически не искажается.
-Для раскрытия термина "практически" рассмотрим наиболее неудобный пример: пусть $F = R_1 .A_2 > 3 \vee R_3 .A_4 < 4$
-и на кортеже $t$ атрибут $R_1 .A_2$ принимает допустимое значение $\n{NULL}$, а предикат $R_3 .A_4 < 4$ принимает
-значение $\n{FALSE}$. Тогда преобразование формулы $F$ на кортеже $t$ будет иметь значение $\n{TRUE}$,
-что не совсем очевидно. Далее будем предполагать, что все формулы $F$ являются преобразованными.
+Let's write expression (\ref{def_F_1}) for $F$ in the following form: $ F ( \dots, T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
+Then after the modification it will be the following: 
+$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here  $\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - conjunction of all $F$ attributes,for which $\n{NULL}$ is not allowed, and $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, here  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - disjunction of all $F$ attributes,for which $\n{NULL}$ is allowed.
+Outer brackets for $T'_j$ predicate define operation priority.
+We can see that this logical formula can only have $\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic.
+We also can note that if rows don't have undefined values, the initial formula $F$ will be equivalent to the transformed one.
+There can be some cases when the meaning of $P$ IR changes with such a transformation.
+For example, assume $F = R_1 .A_2 > 3 \vee R_3 .A_4 < 4$.
+Let row $t$ has $R_1 .A_2$ value set to $\n{NULL}$, with a property <<Use of undefined value>> set to <<Yes>>, and $R_3 .A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$.
+Then $F$ transformation on $t$ row will be $\n{TRUE}$, what is not so obvious indeed.
 
-Введем в рассмотрение множество $\mathcal{A} =$ $\{(a_1, \dots, a_n)
-\mid a_i \in Dom(A_i), i=1,\dots,n\}$, где $Dom(A_i)$ - множество всех допустимых значений атрибута $A_i$.
-Декартово прозведение $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ - $n$-мерное пространство
-значений всех атрибутов базы данных. Текущее состояние базы данных соответствует точке в этом пространстве.
-Допустимые состояния образуют некоторую область в этом пространстве, соответствующую ограничениям целостности на данные.
-В замечании \ref{trivial} речь идет о всем пространстве состояний, а не только допустимых состояниях.
+Hereafter we will suppose all the $F$ formulas to be transformed.
+
+We consider $\mathcal{A} =$ $\{(a_1, \dots, a_n)
+\mid a_i \in Dom(A_i), i=1,\dots,n\}$ set, here $Dom(A_i)$ - is a domain of $A_i$.
+Cartesian product $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ is an $n$-dimensional space of all values for all Database attributes.
+Data constraints bound this space to some set of points that represents available Database states.
+In the Note \ref{trivial} we consider the whole space, not only the available states.
 
 \begin{mydef}
-Областью истинности логической формулы $F$, заданной (\ref{def_F_1}),
-(\ref{def_F_2}), (\ref{def_F_3}), является множество, определяемое по следующему
-правилу: $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
+Truth space of logical formula $F$, defined by (\ref{def_F_1}),
+(\ref{def_F_2}), (\ref{def_F_3}), is a set $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
 \end{mydef}
@@
 \begin{note}
-Размерность формулы $F$ может быть меньше
-размерности $\mathcal{A}$, при этом атрибуты, не входящие в формулу, могут
-принимать любые значения.
+$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case attributes that aren't presented in formula, can take any values in their domains.
 \end{note}
@@
 \begin{note}
-Сложность компонентов предикатов $\n{Expr}$, $S$ и $\n{Str}$ определяется
-возможностями программного обеспечения по вычислению областей истинности
-$F$, как это показано в примере 1.
+The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by software ability to compute Truth spaces of formulas, as it is shown in example 1.
 \end{note}
 
-В соответствии с данными определениями легко понять, как будут устроены операции
-над областями истинности, соответствующими формулам. $M (F)$ для некоторой формулы
-$F$, заданной своей ДНФ, является объединением областей истинности, представленных
-отдельными конъюнктами формулы. Область истинности каждого конъюнкта определяется
-определяется как пересечение областей истинности предикатов, входящих в него.
-
-Далее введем определения, касающиеся модификации вхождения атрибутов в
-логические формулы.
+We now see how can Truth space operations be handled with respect to the corresponding formulas.
+$M (F)$ for a given formula $F$, written in DNF, is nothing but a union of Truth spaces of distinct conjunctive clauses.
+Truth space of a single conjunctive clause is an intersection of Truth spaces of it's predicates.
 
 \begin{mydef}
-Проекцей логической формулы $F$, заданной (\ref{def_F_1}), (\ref{def_F_2}),
-(\ref{def_F_3}), на множество атрибутов $X$ называется логическая
-формула $F[X], \<F[X]> $
-$= X$, в которой все термы, содержащие
-атрибуты $R_i^F.A_i^F \notin X$, заменены на тривиальный терм $\n{TRUE}$.
+Given a logical formula $F$, defined by (\ref{def_F_1}), (\ref{def_F_2}),
+(\ref{def_F_3}), the projection of $F$ on the $X$ attribute set is a logical formula $F[X], \<F[X]> = X$, that has all its predicates with $R_i^F.A_i^F \notin X$ replaced with trivial predicate $\n{TRUE}$.
 \label{projection}
@@
 
-\begin{statement}[Свойство включения]
+\begin{statement}[Inclusion property]
 $\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
@@
 \begin{proof}
-Если $X = \<F> $, то $F = F[X]$ и $M(F) = M(F[X])$. Пусть $X \subset \<F> $. Рассмотрим
-произвольную точку $a \in M(F)$, то есть $F(a)=\n{TRUE}$. Преобразование формулы $F$
-к виду $F[X]$ выполняется заменой предикатов $T_j$ значением $\n{TRUE}$, если они содержат
-атрибут $A_j \in \<F> $ и $A_j \notin X$. Учитывая соотношения  (\ref{def_F_1}) и (\ref{def_F_2}),
-получим $F([X])=\n{TRUE}$. Следовательно, $a \in M(F[X])$, что и требовалось показать.
+If $X = \<F> $, then $F = F[X]$ and $M(F) = M(F[X])$.
+Assume $X \subset \<F> $.
+Consider arbitrary point $a \in M(F)$, i.e.
+$F(a)=\n{TRUE}$.
+$F$ to $F[X]$ transformation is made by replacing $T_j$ predicates that contain $A_j \in \<F> $ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value.
+We get $F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}).
+Therefore, $a \in M(F[X])$.
 \end{proof}
 
-Данное свойство логических ограничений, заданных на состояниях базы
-данных, в дальнейшем используется при формировании нового представления данных
-из существующих сохраненных представлений.
+This property of logical formulas is used while building new data representation from given IDRs.
 
-\section{Исследование свойств промежуточных представлений данных}
+\section{Intermediate Data Representation properties research}
 
-Сохраненные промежуточные представления данных обозначим $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$,
-где $P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
-$s(v)$ – количество отношений в базе данных, использованных при формировании представления $P_v$,
-$\pi_{X_v}$ - операция проекции по множеству атрибутов $X_v$, $\sigma_{F_v}$ - операция селекции
-с логическим ограничением на кортежи $F_v$.
-Целевое выражение, которое надо будет получить из представлений $P$, запишем в виде:
+We denote saved IDRs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
+$s(v)$ – amount of relations in Database that were used while building $P_v$ IDR,
+$\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection with $F_v$ logical formula. We need to get the following representation out of $P$ IDRs:
 $$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l )$$
 
-Рассмотрим проблему формирования представления данных $P^{\ast}$ из существующих промежуточных
-представлений $P_v$.
+Lets study the problem of building data representation $P^{\ast}$ using existing $P_v$ IDRs.
 
@@
 
-Предложенные в теореме условия гарантируют, что данные, необходимые для формирования
-представления $P^{\ast}$, содержатся в промежуточном представлении $P_v$. Однако в нем
-могут быть лишние кортежи, которые дают значение $\n{TRUE}$ при подстановке в формулу
-$F^{\ast}$. Дело в том, что эти кортежи будут удалены при выполнении операции естественного
-соединения с отношениями, которых не хватает в множестве $R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} $
-для совпадения с множеством $R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l $.
-Используя области области истинности логических выражений у СУБД можно запросить
-минимально необходимый набор данных для определения лишних кортежей.
+The conditions given in theorem guarantee the data to build $P^{\ast}$ to be contained in $P_v$ IDR.
+However, there can be excess rows that make $F^{\ast}$ take $\n{TRUE}$ value.
+They appear because there are some relations in $R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l $ that aren't presented in $R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} $ and will be deleted if we join those missing relations.
+Using Truth spaces of logical formulas we can query DBMS to get a minimal required data set to define excess rows.
 
-Следующая теорема соответствует частному случаю, однако проблема лишних кортежей не возникает.
+In the next theorem we describe the case of coinciding relation sets.
 
@@
 
-Для формирования результирующего представления возможно использование
-не одного, а нескольких промежуточных представлений $P_v$. Для этого
-рассмотрим достаточно очевидное свойство операции естественного соединения.
+It is possible to use several IRs to build resulting representation. But at first let's consider the following simple property of natural join.
 
 \begin{statement}
-Пусть $\Re_1 = R_1 \Join \dots \Join R_k$ - результат естественного
-соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \Join \dots
-\Join R_k \Join R_{k+1} \Join \dots \Join R_{n}$. 
-Тогда $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
+Assume $\Re_1 = R_1 \Join \dots \Join R_k$ - natural join of some $k$ relations.
+Also assume $\Re_2 = R_1 \Join \dots \Join R_k \Join R_{k+1} \Join \dots \Join R_{n}$.
+Then $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
 \label{join_property}
 \end{statement}
-
-Это действительно так, поскольку после проведения операции естественного соединения
-с дополнительными отношениями исходные кортежи, содержавшиеся в $\Re_1$, могут
-лишь ''отсеяться'' операцией естественного соединения. Тогда, взяв вырезку по
-исходным атрибутам, получим как максимум то же самое отношение $\Re_1$. Поскольку
-операция естественного соединения коммутативна, то порядок присоединения отношений
-не имеет значения.
+\begin{proof}
+It is so indeed, because joining additional relations to $\Re_1$ can only remove some rows that were presented there.
+After the projection we will get the same relation $\Re_1$ if no rows were deleted or some subset otherwise.
+The order of joins doesn't matter thanks to commutativity property of this operation.
+\end{proof}
 
@@
 
-Как и в предыдущем случае существуют условия, при которых целевое представление
-данных можно вычислить точно.
+Just like in previous case, there are some additional conditions that allow us to get precise data presentation from IRs.
 
@@
 
-Выполнение условий рассмотренных теорем не представляет собой  исключительный случай,
-поскольку набор размерностей при формировании гиперкубов изменяется редко, что позволит
-формировать эти размерности из промежуточных представлений без обращения к СУБД.
+The conditions listed above aren't exceptional in any case, because the set of dimensions that form hypercube changes rarely, so we can create new dimensions (data representations) from IRs with much less DB interaction.
 
-\section{Заключение}
+\section{Conclusion}
+The prime result of our paper is the justification of IR usage conditions for new hypercube creation.
+Those IRs can be stored on an analyst PC and decrease the time of getting the required data drastically.
 
-Основной результат данной работы состоит в обосновании условий использования
-зарезервированных представлений данных для формирования новых гиперкубов.
-Эти представления данных могут храниться на компьютере пользователя-аналитика
-и существенно сократить время на формирование данных, необходимых для принятия решений.
+The common problem of systems that don't use any DBMS is the data refreshing.
+In MOLAP approach technical support stuff is responsible for periodical refreshing of data.
+The same method can be used to refresh $P_v$ representations.
+To decrease the refresh time period it is possible to get access to the DBMS's DB modification log and refresh only modified data.
+But let's notice again that those dimensions don't change often.
 
-Традиционная проблема в системах, работающих не под управлением СУБД, заключается
-в актуализации данных, полученных из БД. В технологии MOLAP она решается путем периодической
-актуализации содержимого гиперкуба службой сопровождения. Аналогичный способ может быть
-применен для актуализации представлений $P_v$. Для сокращения времени актуализации возможно
-получить доступ к журналу изменений, который сопровождает СУБД, и актуализировать только те
-представления, для которых изменились исходные данные в БД. Однако еще раз отметим, что данные,
-используемые в размерностях гиперкуба, обычно редко модифицируются в БД.
-
-Предлагаемый подход подразумевает в дальнейшем применение графических процессоров
-для выполнения промежуточных операций фильтрации и соединения различных представлений данных.
-Исследования в этой области показали, что операции на графическом процессоре выполняются
-существенно быстрее, чем на центральном процессоре, для небольших объемов данных. При увеличении
-объемов производительность падает. Причина очевидна: пока данные помещаются в быструю память
-графического процессора, операции с ними выполняются быстро. При увеличении объемов данных появляется
-необходимость их многократного перемещения между различными видами памяти, что снижает
-производительность. С одной стороны, производители графических процессоров постоянно наращивают
-быструю память, с другой – повторное использование данных наиболее актуально для размерностей
-многомерных таблиц, для которых требуется сравнительно небольшой объем памяти, что в совокупности
-делает целесообразным применение графических процессоров при решении указанной проблемы.
-
+We assume the usage of GPU to implement intermediate filtration and union of different data representations.
+The researches in this field show that GPU operations are much faster than those on CPU for small data.
+There is a decrease in perfomance with data scale.
+The reason is obvious: big amount of data leads to a much more frequent data transfer between different memory types and increases the overall time spent.
+But the constant GPU memory growth and the fact that data amount required for multidimensional data dimensions building is comparatively small make the usage of graphical processors quite a good choice for these purposes.
 \input{bibliography}

### 2014-04-22 english_version/article_main.tex
translation corrected

@@
 \usepackage{xcolor}
+\usepackage{textcomp}
 \floatname{algorithm}{Algorithm}
@@
 \author{
-Mosin Sergey, Zykin Sergey \footnote{The paper is supported by RFBR grant
-#12-07-00066-a}
-\date{}
+Mosin Sergey, Zykin Sergey
+\footnote{The paper is supported by RFBR grant \#12-07-00066-a}
 }
@@
 This problem was already discussed and has some software implementations. In our
-work we propose logical formulas Truth spaces usage while querying Database.
+work, we propose Truth spaces usage in logical formulas while querying a Database.
 This approach is supposed to help an analyst, who works with dynamically built
@@
 \section{Introduction}
-A lot of nowadays Big Data analysis solutions use Online Analytical Processing
-technologies, formed in early 90s already \cite{codd}. A lot of fundamental
+A lot of current Big Data analysis solutions use Online Analytical Processing
+technologies, formed in early 90s \cite{codd}. A lot of fundamental
 \cite{lecht,lehner,mazon} and applied \cite{vassi, peder, progressive, giorg}
@@
 
-The purpose of this paper is to study problem of building Dynamic
+The purpose of this paper is to study the problem of building Dynamic
 Multidimensional Data Representation from the Relational Database using
@@
 
-This problem is connected with a field of query optimization, because it aims
-the decrease in data transfer from a Database server. Cached data is actively
-used in Database Management Systems (DBMS). But mostly it is only repeating use
+This problem is connected with a field of query optimization, because it aims to
+decrease the data transfer from a Database server. Cached data is actively
+used in Database Management Systems (DBMS), but mostly, it is only repeating use
 of data, written in cache, without any prior data analysis to define any partial
 or combined use. The DBMS can only avoid requesting blocks of data from external
-devices while serving the query, if they are present(?ed) in cache. It means
-block number analysis, not their information.
+devices while serving the query, if they are present in the cache. So, the
+block numbers are analyzed, not their information.
 
@@
 
-In a field of fundamental researches huge amount of publications aim optimal
-query plan generation based on a formal set of rules, that don't use SQL
-operator predicate domains (logical optimization). In other cases those domains
+In a field of fundamental researches, huge amounts of publications aim at optimal
+query plan generation based on a formal set of rules that don't use SQL
+operator predicate domains (logical optimization). In other cases, those domains
 are taken into account for statistical evaluations to improve physical access to
@@
 different results. The most similar to our work is paper \cite{Afrati06}. They
-analyze conjunctive queries on data domains with predicates in a form of
-arithmetical comparisons and present query computation algorithms using IDRs. In
-this paper the special case of universal relational query is considered. It is a
+analyze conjunctive queries on data domains with predicates in the form of
+arithmetical comparisons, and present query computation algorithms using IDRs. In
+our paper, the special case of universal relational query is considered. It is a
 query on the Database relations, not particular domains. Although we had similar
-aims, the results obtained differ because of stated factors. In particular, in
-our work there is no need to create any algorithms of data selection, as they
+aims, the results obtained differ because of stated factors. In particular,
+there is no need to create any algorithms of data selection in our work, as they
 are replaced by Relational Algebra.
 
-There are some works, which examine query optimization problem in the case of
-multidimensional warehouse representation with a use of dynamically formed
+There are some works, which examine the query optimization problem in the case of
+multidimensional warehouse representation with the use of dynamically formed
 representations. Two basic methods are used there: statical \cite{baralis,
 gupta, gupta-mumick} and dynamical one \cite{scheuermann, shim, kalnis, chang}.
-The first one is based on a fixed amount of queries. In the latter one queries
+The former is based on a fixed amount of queries, whereas in the latter queries
 for reservation are chosen based on appearance statistic and computational cost
@@
 
-In the \cite{Keller96} paper they claim for cache content analysis, and
-predicates correspond to IDRs in our approach. Cached data usage problem is solved
-in terms of predicate deduction. In our paper we solve this problem by means of
+In the \cite{Keller96} paper, they analyze the cache content,
+and predicates in their work correspond to IDRs in our approach. Cached data
+usage problem is resolved in terms of predicate deduction. In our paper, we address this problem by means of
 computation of cached query results domains that correspond to the IDRs. It allows
-us to define IR records that can be used to form new representation and new SQL
+us to define IR records that can be used to form a new representation and new SQL
 queries that will let us load missing data from Database server. The following
@@
 \end{table}
-Assume the following precomputed relations on the user PC:  $P_1$,
-$P_2$, $P_3$.
-Table schemes are the following then: $P_1$=(Department number, Department), without any data constraints; $P_2$=(Discipline number, Discipline), without data constraints; $P_3$=(Department number, Discipline number, Semester, Class, Hours), with the following constraint:
+Assume the precomputed relations on the user PC:  $P_1$,
+$P_2$, $P_3$ with the following schemes: $P_1$=(Department number, Department), without any data constraints; $P_2$=(Discipline number, Discipline), without data constraints; $P_3$=(Department number, Discipline number, Semester, Class, Hours), with the following constraint:
 $F_3 =(\mbox{Semester} \leq 2) \wedge (\mbox{Department} = \mbox{D. of Chemistry})$.
@@
 of Chemistry}))$.
-Using $P_3$ schema and $F$ formula, we can now make an SQL command to be executed on server.
+Using $P_3$ schema and $F$ formula, we can now make an SQL command to be executed on the server.
 As a result, we transfer less data over the network.
@@
 \begin{note}
-We assume logical formulas have no trivial conditions on attributes, for example, $Expr_1 = Expr_1$ and those reduced to such form.
-In general, we assume that $A_l$ in $R_i$ domain is not fully contained in $T_j (\dots , R_i .
-A_l ,\dots )$ predicate truth space.
-Such conditions can be removed from the formula without a change in the truth space (we will define it later).
+We assume logical formulas have no trivial conditions on attributes, for example
+, $Expr_1 = Expr_1$ and those reduced to such form. In general, we assume that
+$R_i.A_l$ domain is not fully contained in $T_j (\dots , R_i . A_l ,\dots )$
+predicate truth space. Such conditions can be removed from the formula without a
+change in the truth space (we will define it later).
 \label{trivial}
@@
 \begin{mydef}
-Space of attributes contained in a formula show the dimension of formula and is denoted as $\<F> $.
+Space of attributes contained in a formula show the dimension of the formula and
+is denoted as $\<F> $.
 
@@
 
-Listed variants of operations don't use every single SQL capability.
-For example, we don't use $\n{EXISTS}$ predicate, because it has no expanded attribute names in it.
-$\n{NULL}$ predicate is used for another purpose in our paper.
+Listed variants of operations don't use every single SQL capability. For example
+, we don't use $\n{EXISTS}$ predicate, because it has no expanded attribute
+names in it. $\n{NULL}$ predicate is used for another purpose in our paper.
 
-During logical formula domain calculation if we have some attribute having $NULL$ value on a row $t$, we then get the $\n{UNKNOWN}$ value for the whole formula, because SQL-query results follow Three-valued logic.
-It leads us to ambiguous interpretations of result of both usual users and experienced programmers.
-To solve this problem we suggest the following constraint: every attribute in $F^{\ast}$ is supplied with a property <<Use of undefined value>> with two mutually exclusive values: <<Yes>> or <<No>>.
-The motivation behind this property is the following: if it is assigned <<Yes>>, then we leave rows with $\n{NULL}$ value for further consideration.
-Otherwise, having <<No>> in that property guarantees us removing all such rows.
+During logical formula domain calculation if we have some attribute having
+$NULL$ value on a row $t$, we then get the $\n{UNKNOWN}$ value for the whole
+formula, because SQL-query results follow the Three-valued logic. It leads us to
+ambiguous interpretations of result of both usual users and experienced
+programmers. To solve this problem we suggest the following constraint: every
+attribute in $F^{\ast}$ is supplied with a property ``Use of undefined value''
+with two mutually exclusive values: ``Yes'' or ``No''. The reasoning behind
+this property is the following: if it is assigned ``Yes'', then we leave rows
+with $\n{NULL}$ value for further consideration. Otherwise, having ``No'' in
+that property guarantees us removing all such rows.
 
-Let's write expression (\ref{def_F_1}) for $F$ in the following form: $ F ( \dots, T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
+Let's write expression (\ref{def_F_1}) for $F$ in the following form: $ F (\dots
+, T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
 Then after the modification it will be the following: 
-$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here  $\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - conjunction of all $F$ attributes,for which $\n{NULL}$ is not allowed, and $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, here  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - disjunction of all $F$ attributes,for which $\n{NULL}$ is allowed.
-Outer brackets for $T'_j$ predicate define operation priority.
-We can see that this logical formula can only have $\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic.
-We also can note that if rows don't have undefined values, the initial formula $F$ will be equivalent to the transformed one.
-There can be some cases when the meaning of $P$ IR changes with such a transformation.
-For example, assume $F = R_1 .A_2 > 3 \vee R_3 .A_4 < 4$.
-Let row $t$ has $R_1 .A_2$ value set to $\n{NULL}$, with a property <<Use of undefined value>> set to <<Yes>>, and $R_3 .A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$.
-Then $F$ transformation on $t$ row will be $\n{TRUE}$, what is not so obvious indeed.
+$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here
+$\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - conjunction of all $F$ attributes,for
+which $\n{NULL}$ is not allowed, and $T'_j = (T_j \vee_{i,j}(R_i .A_j =
+\n{NULL}))$, here  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - disjunction of all $F$
+attributes, for which $\n{NULL}$ is allowed. Outer brackets for $T'_j$ predicate
+define operation priority. We can see that this logical formula can only have
+$\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic. We
+also can note that if rows don't have undefined values, the initial formula $F$
+will be equivalent to the transformed one. There can be some cases when the
+meaning of $P$ IR changes with such a transformation. For example, assume $F =
+R_1 .A_2 > 3 \vee R_3 .A_4 < 4$. Let row $t$ has $R_1 .A_2$ value set to
+$\n{NULL}$, with a property ``Use of undefined value'' set to ``Yes'', and $R_3
+.A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$. Then $F$
+transformation on $t$ row will be $\n{TRUE}$, what is not so obvious indeed.
 
@@
 We consider $\mathcal{A} =$ $\{(a_1, \dots, a_n)
-\mid a_i \in Dom(A_i), i=1,\dots,n\}$ set, here $Dom(A_i)$ - is a domain of $A_i$.
-Cartesian product $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ is an $n$-dimensional space of all values for all Database attributes.
-Data constraints bound this space to some set of points that represents available Database states.
-In the Note \ref{trivial} we consider the whole space, not only the available states.
+\mid a_i \in Dom(A_i), i=1,\dots,n\}$ set, here $Dom(A_i)$ - is a domain of
+$A_i$. Cartesian product $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$
+is an $n$-dimensional space of all values for all Database attributes. Data
+constraints bound this space to some set of points that represents available
+Database states. In the Note \ref{trivial} we consider the whole space, not only
+the available states.
 
@@
 Truth space of logical formula $F$, defined by (\ref{def_F_1}),
-(\ref{def_F_2}), (\ref{def_F_3}), is a set $M (F) = \{a \in \mathcal{A} \mid F(a) = \n {TRUE}\}$.
+(\ref{def_F_2}), (\ref{def_F_3}), is a set $M (F) = \{a \in \mathcal{A} \mid
+F(a) = \n {TRUE}\}$.
 \end{mydef}
@@
 \begin{note}
-$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case attributes that aren't presented in formula, can take any values in their domains.
+$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case
+attributes that aren't presented in formula can take any values in their
+domains.
 \end{note}
@@
 
-We now see how can Truth space operations be handled with respect to the corresponding formulas.
+We now see how Truth space operations can be handled with respect to the corresponding formulas.
 $M (F)$ for a given formula $F$, written in DNF, is nothing but a union of Truth spaces of distinct conjunctive clauses.
@@
 \begin{proof}
-If $X = \<F> $, then $F = F[X]$ and $M(F) = M(F[X])$.
-Assume $X \subset \<F> $.
-Consider arbitrary point $a \in M(F)$, i.e.
-$F(a)=\n{TRUE}$.
-$F$ to $F[X]$ transformation is made by replacing $T_j$ predicates that contain $A_j \in \<F> $ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value.
-We get $F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}).
-Therefore, $a \in M(F[X])$.
+If $X = \<F> $, then $F = F[X]$ and $M(F) = M(F[X])$. Assume $X \subset \<F> $.
+Consider arbitrary point $a \in M(F)$, i.e. $F(a)=\n{TRUE}$. $F$ to $F[X]$
+transformation is made by replacing $T_j$ predicates that contain $A_j \in \<F> 
+$ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value. We get
+$F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}). Therefore,
+$a \in M(F[X])$.
 \end{proof}
 
-This property of logical formulas is used while building new data representation from given IDRs.
+This property of logical formulas is used while building new data representation
+from given IDRs.
 
@@
 
-We denote saved IDRs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
+We denote saved IDRs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v =
+\pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
 $s(v)$ – amount of relations in Database that were used while building $P_v$ IDR,
-$\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection with $F_v$ logical formula. We need to get the following representation out of $P$ IDRs:
-$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l )$$
+$\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection with
+$F_v$ logical formula. We need to get the following representation out of $P$
+IDRs:
+$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join R^{\ast}_2\Join
+\dots \Join R^{\ast}_l )$$
 
-Lets study the problem of building data representation $P^{\ast}$ using existing $P_v$ IDRs.
+Lets study the problem of building data representation $P^{\ast}$ using existing
+$P_v$ IDRs.
 
@@
 
-It is possible to use several IRs to build resulting representation. But at first let's consider the following simple property of natural join.
+It is possible to use several IRs to build resulting representation. At first,
+let's consider the following simple property of natural join.
 
@@
 It is so indeed, because joining additional relations to $\Re_1$ can only remove some rows that were presented there.
-After the projection we will get the same relation $\Re_1$ if no rows were deleted or some subset otherwise.
+After the projection, we will get the same relation $\Re_1$ if no rows were deleted or some subset otherwise.
 The order of joins doesn't matter thanks to commutativity property of this operation.
@@
 \section{Conclusion}
-The prime result of our paper is the justification of IR usage conditions for new hypercube creation.
-Those IRs can be stored on an analyst PC and decrease the time of getting the required data drastically.
+The prime result of our paper is the justification of IR usage conditions for
+new hypercube creation. Those IRs can be stored on an analyst's PC and
+drastically shorten the time of getting the required data.
 
-The common problem of systems that don't use any DBMS is the data refreshing.
-In MOLAP approach technical support stuff is responsible for periodical refreshing of data.
-The same method can be used to refresh $P_v$ representations.
-To decrease the refresh time period it is possible to get access to the DBMS's DB modification log and refresh only modified data.
-But let's notice again that those dimensions don't change often.
+The common problem of systems that don't use any DBMS is the data refreshing. In
+MOLAP approach, the technical support staff is responsible for periodical
+refreshing of data. The same method can be used to refresh $P_v$
+representations. The access to the DMBS's DB modification log and refreshing
+only modified data decreases the refresh time period. However, let's notice
+again that those dimensions don't change often.
 
-We assume the usage of GPU to implement intermediate filtration and union of different data representations.
-The researches in this field show that GPU operations are much faster than those on CPU for small data.
-There is a decrease in perfomance with data scale.
-The reason is obvious: big amount of data leads to a much more frequent data transfer between different memory types and increases the overall time spent.
-But the constant GPU memory growth and the fact that data amount required for multidimensional data dimensions building is comparatively small make the usage of graphical processors quite a good choice for these purposes.
+We assume the usage of GPU to implement intermediate filtration and union of
+different data representations. The research in this field shows that GPU
+operations are much faster than those on CPU for small data. There is a decrease
+in performance with data scale. The reason is obvious: big amount of data leads
+to a much more frequent data transfer between different memory types and
+increases the overall time elapsed. But the constant GPU memory growth and the
+fact that data amount required for multidimensional data dimensions building is
+comparatively small make the usage of graphical processors quite a good choice
+for these purposes.
 \input{bibliography}

### 2014-04-22 english_version/theorem_2.tex
translation corrected

@@
-\author{Мосин Сергей, Зыкин Сергей}
 \begin{theorem}
 $P^{\ast} = \pi_{X^{\ast}} (P_{v})$
-, если: 
-%\textcolor{red}{!!! Селекцию можно убрать $: P^{\ast} = \pi_{X^{\ast}} (P_{v})$}
-\\а) $X^{\ast} \subseteq X_{v}$
-\\б) 
+, if: 
+\\a) $X^{\ast} \subseteq X_{v}$
+\\b) 
 $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\}$
-\\в) $M (F^{\ast}) = M (F_{v}) $.
+\\c) $M (F^{\ast}) = M (F_{v}) $.
 \label{th_base_eq}
@@
 \begin{proof}
-Поскольку условия теоремы являются частным случаем теоремы \ref{th_base}, то
-включение $P^{\ast} \subseteq \pi_{X^{\ast}} (P_{v})$
-доказано. Необходимо показать, что
-$\pi_{X^{\ast}} (P_{v}) \subseteq P^{\ast}$. Пусть
-кортеж $t_v \in \pi_{X^{\ast}} (P_{v})$. Покажем, что
-$t_v \in P^{\ast}$.
-Из определения кортежа $t_v$ следует, что существует кортеж  $t' \in R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)}$ и $t_v = t'[X^{\ast}]$, $F_v(t')=\n{TRUE}$.
-По свойству операции естественного соединения $R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)} = R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join R^{\ast}_l$, следовательно, $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join R^{\ast}_l$.
-В силу условия в) $F^{\ast} (t')=\n{TRUE}$.
-Отсюда следует, что $t_v \in P^{\ast}$.
-Теорема доказана.
+Conditions of the theorem are a special case of theorem \ref{th_base}, so
+inclusion $P^{\ast} \subseteq \pi_{X^{\ast}} (P_{v})$ is considered  to be
+proved. It is necessary to show that $\pi_{X^{\ast}} (P_{v}) \subseteq
+P^{\ast}$. Assume tuple $t_v \in \pi_{X^{\ast}} (P_{v})$. Let's show that $t_v
+\in P^{\ast}$. From definition of tuple $t_v$ it follows that there is tuple
+$t' \in R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)}$ and $t_v = t'[X^{\ast}]$,
+$F_v(t')=\n{TRUE}$. According to commutativity of natural join, $R^v_1 \Join
+R^v_2\Join\ldots \Join R^v_{s(v)} = R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join
+R^{\ast}_l$. Therefore, $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join
+R^{\ast}_l$. From condition c) we have $F^{\ast} (t')=\n{TRUE}$. Hence, $t_v \in
+P^{\ast}$. The theorem is proved.
 \end{proof}

### 2014-04-23 english_version/IEEE_TKDE/bibliography.tex
IEEEtran version added

--- 2014-03-01 bibliography.tex
@@
 \bibitem{codd}
-  \textit{Codd E.F., Codd S.B., Salley C.T.}
-  Providing OLAP (On-line Analytical Processing) to User-Analysts: An IT
-  Mandate. – Sunnyvale(CA): Codd \& Date Inc.,
+  E.F. Codd, S.B. Codd, C.T. Salley
+  \textit{Providing OLAP (On-line Analytical Processing) to User-Analysts: An IT
+  Mandate} – Sunnyvale(CA): Codd \& Date Inc.,
   1993.
@@
 \bibitem{zyk_pol}
-  \textit{С.В. Зыкин, А.Н. Полуянов}
-  Формирование многомерных данных с использованием промежуточных представлений.
-  "Проблемы управления".
+  \textit{Sergey Zykin, Andrey Poluyanov}
+  Multidimensional data building using Intermediate Representations.
+  "Administration problems".
   2013.
-  № 5. Стр. 54-59.
+  NO. 5, P. 54-59.
 

### 2014-04-23 english_version/IEEE_TKDE/paper_main.tex
IEEEtran version added

%% bare_jrnl_compsoc.tex
%% V1.3
%% 2007/01/11
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
%% Society journal paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. IEEE's font choices can trigger bugs that do  ***
% *** not appear when using other class files.                            ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

% Note that the a4paper option is mainly intended so that authors in
% countries using A4 can easily print to A4 and see how their papers will
% look in print - the typesetting of the document will not typically be
% affected with changes in paper size (but the bottom and side margins will).
% Use the testflow package mentioned above to verify correct handling of
% both paper sizes by the user's LaTeX system.
%
% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
% should be used if it is desired that the figures are to be displayed in
% draft mode.
%
% The Computer Society usually requires 10pt for submissions.
%
\documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
\ifCLASSOPTIONcompsoc
  % IEEE Computer Society needs nocompress option
  % requires cite.sty v4.0 or later (November 2003)
  % \usepackage[nocompress]{cite}
\else
  % normal IEEE
  % \usepackage{cite}
\fi
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
% not currently provide for hyperlinked citations.
% The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
% The documentation is contained in the cite.sty file itself.
%
% Note that some packages require special options to format as the Computer
% Society requires. In particular, Computer Society  papers do not use
% compressed citation ranges as is done in typical IEEE papers
% (e.g., [1]-[4]). Instead, they list every citation separately in order
% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
% package with the nocompress option which is supported by cite.sty v4.0
% and later. Note also the use of a CLASSOPTION conditional provided by
% IEEEtran.cls V1.7 and later.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at: 
% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage[cmex10]{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics. If using
% it, be sure to load this package with the cmex10 option to ensure that
% only type 1 fonts will utilized at all point sizes. Without this option,
% it is possible that some math symbols, particularly those within
% footnotes, will be rendered in bitmap form which will result in a
% document that can not be IEEE Xplore compliant!
%
% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
% There is also a support site at:
% http://algorithms.berlios.de/index.html
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/required/tools/

%\usepackage{mdwmath}
%\usepackage{mdwtab}
% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
% especially mdwmath.sty and mdwtab.sty which are used to format equations
% and tables, respectively. The MDWtools set is already installed on most
% LaTeX systems. The lastest version and documentation is available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

%\usepackage{eqparbox}
% Also of notable interest is Scott Pakin's eqparbox package for creating
% (automatically sized) equal width boxes - aka "natural width parboxes".
% Available at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%\usepackage[tight,normalsize,sf,SF]{subfigure}
%\else
%\usepackage[tight,footnotesize]{subfigure}
%\fi
% subfigure.sty was written by Steven Douglas Cochran. This package makes it
% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
% work, it is a good idea to load it with the tight package option to reduce
% the amount of white space around the subfigures. Computer Society papers
% use a larger font and \sffamily font for their captions, hence the
% additional options needed under compsoc mode. subfigure.sty is already
% installed on most LaTeX systems. The latest version and documentation can
% be obtained at:
% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
% subfigure.sty has been superceeded by subfig.sty.

%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false]{caption}
%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false]{caption}
%  \usepackage[font=footnotesize]{subfig}
%\fi
% subfig.sty, also written by Steven Douglas Cochran, is the modern
% replacement for subfigure.sty. However, subfig.sty requires and
% automatically loads Axel Sommerfeldt's caption.sty which will override
% IEEEtran.cls handling of captions and this will result in nonIEEE style
% figure/table captions. To prevent this problem, be sure and preload
% caption.sty with its "caption=false" package option. This is will preserve
% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
% (recommended due to many improvements over 1.2) of subfig.sty supports
% the caption=false option directly:
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
%
% The latest version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
% The latest version and documentation of caption.sty can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/

% *** FLOAT PACKAGES ***
%
\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure. The latest version and documentation can be found at:
% http://www.ctan.org/tex-archive/macros/latex/base/

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
% Documentation is contained in the stfloats.sty comments as well as in the
% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
% does not allow \baselineskip to stretch. Authors submitting work to the
% IEEE should note that IEEE rarely uses double column equations and
% that authors should try to avoid such use. Do not be tempted to use the
% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
% not format its papers in such ways.

%\ifCLASSOPTIONcaptionsoff
%  \usepackage[nomarkers]{endfloat}
% \let\MYoriglatexcaption\caption
% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
%\fi
% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
% This package may be useful when used in conjunction with IEEEtran.cls'
% captionsoff option. Some IEEE journals/societies require that submissions
% have lists of figures/tables at the end of the paper and that
% figures/tables without any captions are placed on a page by themselves at
% the end of the document. If needed, the draftcls IEEEtran class option or
% \CLASSINPUTbaselinestretch interface can be used to increase the line
% spacing as well. Be sure and use the nomarkers option of endfloat to
% prevent endfloat from "marking" where the figures would have been placed
% in the text. The two hack lines of code above are a slight modification of
% that suggested by in the endfloat docs (section 8.3.1) to ensure that
% the full captions always appear in the list of figures/tables - even if
% the user used the short optional argument of \caption[]{}.
% IEEE papers do not typically make use of \caption[]'s optional argument,
% so this should not be an issue. A similar trick can be used to disable
% captions of packages such as subfig.sty that lack options to turn off
% the subcaptions:
% For subfig.sty:
% \let\MYorigsubfloat\subfloat
% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
% For subfigure.sty:
% \let\MYorigsubfigure\subfigure
% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
% However, the above trick will not work if both optional arguments of
% the \subfloat/subfig command are used. Furthermore, there needs to be a
% description of each subfigure *somewhere* and endfloat does not add
% subfigure captions to its list of figures. Thus, the best approach is to
% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
% and instead reference/explain all the subfigures within the main caption.
% The latest version of endfloat.sty and its documentation can obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
%
% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
% later in the document, say, to conditionally put the References on a 
% page by themselves.

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version can be obtained at:
% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
% Read the url.sty source comments for usage information. Basically,
% \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\author{Mosin Sergey, Zykin Sergey}
\usepackage{amssymb,amsmath}
\usepackage[utf8]{inputenc}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{xcolor}
\usepackage{textcomp}
\newtheorem{theorem}{Theorem}
\newtheorem{mydef}{Definition}
\newtheorem{statement}{Statement}
\newtheorem*{consequence}{Consequence}
\newtheorem{note}{Note}
\def \eval #1#2{\left.#1\right\vert_{#2}}
\def \<#1> {\langle #1 \rangle}
\def \n #1{\mathit{#1}}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Building table-like applications using cached data}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%
%
%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
% lists the Computer Society journals use for "first footnote" author
% affiliations. Use \IEEEcompsocthanksitem which works much like \item
% for each affiliation group. When not in compsoc mode,
% \IEEEcompsocitemizethanks becomes like \thanks and
% \IEEEcompsocthanksitem becomes a line break with idention. This
% facilitates dual compilation, although admittedly the differences in the
% desired content of \author between the different types of papers makes a
% one-size-fits-all approach a daunting prospect. For instance, compsoc 
% journal papers have the author affiliations above the "Manuscript
% received ..."  text while in non-compsoc journals this is reversed. Sigh.

\author{Sergey~Mosin,
        Sergey~Zykin % <-this % stops a space
\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. Mosin, S. Zykin are with the
Omsk Department of Sobolev Institute of Mathematics, Omsk, Russia. The paper is
supported by RFBR grant \#12-07-00066-a.\protect\\
% note need leading \protect in front of \\ to get a newline within \thanks as
% \\ is fragile and will error, could use \hfil\break instead.
E-mail: svmosin@gmail.com, szykin@mail.ru}% <-this % stops a space
\thanks{}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Shell \MakeLowercase{\textit{et al.}}: Building table-like applications using
cached data}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% The publisher's ID mark at the bottom of the page is less important with
% Computer Society journal papers as those publications place the marks
% outside of the main text columns and, therefore, unlike regular IEEE
% journals, the available text space is not reduced by their presence.
% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% or like this to get the Computer Society new two part style.
%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark (Computer Society jorunal
% papers don't need this extra clearance.)

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
We primarily address the usage of cached Database query results in our paper.
This problem was already discussed and has some software implementations. In our
work, we propose Truth spaces usage in logical formulas while querying a Database.
This approach is supposed to help an analyst, who works with dynamically built
data representations logically described as multidimensional tables. Those
representations don't take much disk space, so one can use GPU to speed up the
computations. We consider queries in a form of relational algebra expressions,
similar to "universal relational query". Logical formulas in the queries are
presented in Disjunctive Normal Form. They consist of atomic expressions
consisting of SQL predicates. In the paper we consider single and multiple table
cases to build new data representations.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway. In particular, the Computer Society does
% not want either math or citations to appear in the abstract.

% Note that keywords are not normally used for peer review papers.
\begin{keywords}
Database, OLAP, cache, Data Warehouse, predicate
\end{keywords}}

% make the title area
\maketitle

% To allow for easy dual compilation without having to reenter the
% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
% not be used in maketitle, but will appear (i.e., to be "transported")
% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
% is not selected <OR> if conference mode is selected - because compsoc
% conference papers position the abstract like regular (non-compsoc)
% papers do!
\IEEEdisplaynotcompsoctitleabstractindextext
% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
% compsoc under a non-conference mode.

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
% Computer Society journal papers do something a tad strange with the very
% first section heading (almost always called "Introduction"). They place it
% ABOVE the main text! IEEEtran.cls currently does not do this for you.
% However, You can achieve this effect by making LaTeX jump through some
% hoops via something like:
%
%\ifCLASSOPTIONcompsoc
%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
%  \global\everypar=\everypar}}%
%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
%\else
%  \section{Introduction}\label{sec:introduction}\par
%\fi
%
% Admittedly, this is a hack and may well be fragile, but seems to do the
% trick for me. Note the need to keep any \label that may be used right
% after \section in the above as the hack puts \section within a raised box.

% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps (small caps for compsoc).
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A}{lot} of current Big Data analysis solutions use Online
Analytical Processing technologies, formed in early 90s \cite{codd}. A lot of
fundamental \cite{lecht,lehner,mazon} and applied \cite{vassi, peder,
progressive, giorg} results were obtained in this field lately.

The purpose of this paper is to study the problem of building Dynamic
Multidimensional Data Representation from the Relational Database using
cached data representations on a user computer, equipped with a Graphical
Processing Unit (GPU). We don't require any transformation of the initial
Database schema, even to the hierarchical model.

This problem is connected with a field of query optimization, because it aims to
decrease the data transfer from a Database server. Cached data is actively
used in Database Management Systems (DBMS), but mostly, it is only repeating use
of data, written in cache, without any prior data analysis to define any partial
or combined use. The DBMS can only avoid requesting blocks of data from external
devices while serving the query, if they are present in the cache. So, the
block numbers are analyzed, not their information.

This paper is based on the results, obtained in \cite{zyk_pol}. We removed
condition that limited Intermediate Data Representation (IDR) attributes and
also made some generalizations on multiple IDRs case.

\hfill mds

\hfill April 22, 2014

\section{Existing solutions overview and comparison}
In a field of fundamental researches, huge amounts of publications aim at optimal
query plan generation based on a formal set of rules that don't use SQL
operator predicate domains (logical optimization). In other cases, those domains
are taken into account for statistical evaluations to improve physical access to
the data. Stream data query executing problems, described in \cite{Olston03,
Denny05}, are related to our problem closer, but different aims lead us to
different results. The most similar to our work is paper \cite{Afrati06}. They
analyze conjunctive queries on data domains with predicates in the form of
arithmetical comparisons, and present query computation algorithms using IDRs. In
our paper, the special case of universal relational query is considered. It is a
query on the Database relations, not particular domains. Although we had similar
aims, the results obtained differ because of stated factors. In particular,
there is no need to create any algorithms of data selection in our work, as they
are replaced by Relational Algebra.

There are some works, which examine the query optimization problem in the case of
multidimensional warehouse representation with the use of dynamically formed
representations. Two basic methods are used there: statical \cite{baralis,
gupta, gupta-mumick} and dynamical one \cite{scheuermann, shim, kalnis, chang}.
The former is based on a fixed amount of queries, whereas in the latter queries
for reservation are chosen based on appearance statistic and computational cost
of query execution. As a data source Data Warehouses or Relational Databases in
the hierarchical form are used. Special queries for data manipulation are
considered. In this paper we use split dimension generation for the data
required for current analysis. Hence, we are interested in standard SQL queries
to the Database and IDRs. The resulting representation can then be processed by
special multidimensional data queries.

In the \cite{Keller96} paper, they analyze the cache content,
and predicates in their work correspond to IDRs in our approach. Cached data
usage problem is resolved in terms of predicate deduction. In our paper, we address this problem by means of
computation of cached query results domains that correspond to the IDRs. It allows
us to define IR records that can be used to form a new representation and new SQL
queries that will let us load missing data from Database server. The following
example demonstrates suggested approach.

{\bf Example 1.} The task is to receive data representation as in the table
\ref{def_T_1}. Dimension attributes are in bold, measure attributes are italic,
else are attribute values.

\begin{table*}[h!]
\renewcommand{\arraystretch}{1.3}
\caption{\label{def_T_1} Curriculum.}
\centering
\begin{tabular}{|c|c||c|c|c|c|}
\cline{2-6}
\multicolumn{1}{c|}{} & \multicolumn{1}{|r||}{\bf Semester} & \multicolumn{2}{|c|}{2} & \multicolumn{2}{|c|}{3}  \\
\cline{2-6}
\multicolumn{1}{c|}{} & {\bf Discipline} & Physics & Chemistry & Physics & Chemistry \\
\hline \cline{2-6}
{\bf Department} & {\bf Class} & {\it Hours} & {\it Hours} & {\it Hours} & {\it Hours} \\
\hline
 & C-321 & & 36 & 18 & 9 \\
\cline{2-6}
D. of Chemistry & C-322 & & 48 & 18 & 9 \\
\cline{2-6}
 & C-331 & & 24 & 18 & 9 \\
\hline
 & P-321 & 48 & & 24 & \\
\cline{2-6}
D. of Physics & P-322 & 48 & & 24 & \\
\cline{2-6}
 & P-331 & 64 & & 24 & \\
\hline
\end{tabular}
\end{table*}
Assume the precomputed relations on the user PC:  $P_1$,
$P_2$, $P_3$ with the following schemes: $P_1$=(Department number, Department), without any data constraints; $P_2$=(Discipline number, Discipline), without data constraints; $P_3$=(Department number, Discipline number, Semester, Class, Hours), with the following constraint:
$F_3 =(\mbox{Semester} \leq 2) \wedge (\mbox{Department} = \mbox{D. of Chemistry})$.
Assume the user has made the following logical constraint on the data to be represented in the table: $F^{\ast} = (\mbox{Semester} \leq 3)$. Comparing these formulas domains, it's easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1}, whereas $P_3$ needs to be expanded with data. After defining $F^{\ast}$ and $F_3$ domains we can form a new formula: $F=((\mbox{Semester} = 3) \wedge (\mbox{Deparmtent=D. of Chemistry}))
\vee ((\mbox{Semester} \leq 3) \wedge (\mbox{Deparmtent} \neq \mbox{D. of Chemistry}))$. Using $P_3$ schema and $F$ formula, we can now make an SQL command to be executed on the server. As a result, we transfer less data over the network.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\section{Logical constraint definition and properties}

To simplify domain computations we will consider logical formulas in Disjunctive Normal Form (DNF).
In general case formula $F$ has the following form:

\begin{equation}
F = K_1 \vee K_2 \vee \dots \vee K_m ,
\label{def_F_1}
\end{equation}
\begin{equation}
K_i = T_1 \& T_2 \dots \& T_n, i = 1, \dots, m ,
\label{def_F_2}
\end{equation}
here $T_j, j = 1, \dots, n$ - predicates, where expanded attribute names are specified. $R_i.A_j$ means attribute $A_j$ in relation $R_i$.
Those predicates can be:
\begin{itemize}
  \item	comparison operation $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ ?
  comparison operator $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$
  ? type conformant expressions, defined in a space of expanded attribute names and constants;
  \item operation $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
  \n{Expr}_3$ (symbols inside square brackets $[*]$ aren't mandatory for the predicate);
  \item operation $\n{Expr} \n{[NOT]} \n{IN} S$, here $S$ ? values list or subquery, having a set of attribute $R_i.A_j$ values as a result;
  \item operation $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, here $\n{Str}_i$ ?
  strings;
  \item operation $\n{Expr} \theta \n{ALL/ANY} S$.
\end{itemize}

\begin{note}
We assume logical formulas have no trivial conditions on attributes, for example
, $Expr_1 = Expr_1$ and those reduced to such form. In general, we assume that
$R_i.A_l$ domain is not fully contained in $T_j (\dots , R_i . A_l ,\dots )$
predicate truth space. Such conditions can be removed from the formula without a
change in the truth space (we will define it later).
\label{trivial}
\end{note}

\begin{mydef}
Space of attributes contained in a formula show the dimension of the formula and
is denoted as $\<F> $.

\begin{equation}
\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\}
\label{def_F_3}
\end{equation}
\end{mydef}

Listed variants of operations don't use every single SQL capability. For example
, we don't use $\n{EXISTS}$ predicate, because it has no expanded attribute
names in it. $\n{NULL}$ predicate is used for another purpose in our paper.

During logical formula domain calculation if we have some attribute having
$NULL$ value on a row $t$, we then get the $\n{UNKNOWN}$ value for the whole
formula, because SQL-query results follow the Three-valued logic. It leads us to
ambiguous interpretations of result of both usual users and experienced
programmers. To solve this problem we suggest the following constraint: every
attribute in $F^{\ast}$ is supplied with a property ``Use of undefined value''
with two mutually exclusive values: ``Yes'' or ``No''. The reasoning behind
this property is the following: if it is assigned ``Yes'', then we leave rows
with $\n{NULL}$ value for further consideration. Otherwise, having ``No'' in
that property guarantees us removing all such rows.

Let's write expression (\ref{def_F_1}) for $F$ in the following form: $ F (\dots
, T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
Then after the modification it will be the following: 
$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here
$\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - conjunction of all $F$ attributes,for
which $\n{NULL}$ is not allowed, and $T'_j = (T_j \vee_{i,j}(R_i .A_j =
\n{NULL}))$, here  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - disjunction of all $F$
attributes, for which $\n{NULL}$ is allowed. Outer brackets for $T'_j$ predicate
define operation priority. We can see that this logical formula can only have
$\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic. We
also can note that if rows don't have undefined values, the initial formula $F$
will be equivalent to the transformed one. There can be some cases when the
meaning of $P$ IR changes with such a transformation. For example, assume $F =
R_1 .A_2 > 3 \vee R_3 .A_4 < 4$. Let row $t$ has $R_1 .A_2$ value set to
$\n{NULL}$, with a property ``Use of undefined value'' set to ``Yes'', and $R_3
.A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$. Then $F$
transformation on $t$ row will be $\n{TRUE}$, what is not so obvious indeed.

Hereafter we will suppose all the $F$ formulas to be transformed.

We consider $\mathcal{A} =$ $\{(a_1, \dots, a_n)
\mid a_i \in Dom(A_i), i=1,\dots,n\}$ set, here $Dom(A_i)$ - is a domain of
$A_i$. Cartesian product $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$
is an $n$-dimensional space of all values for all Database attributes. Data
constraints bound this space to some set of points that represents available
Database states. In the Note \ref{trivial} we consider the whole space, not only
the available states.

\begin{mydef}
Truth space of logical formula $F$, defined by (\ref{def_F_1}),
(\ref{def_F_2}), (\ref{def_F_3}), is a set $M (F) = \{a \in \mathcal{A} \mid
F(a) = \n {TRUE}\}$.
\end{mydef}

\begin{note}
$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case
attributes that aren't presented in formula can take any values in their
domains.
\end{note}

\begin{note}
The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by software ability to compute Truth spaces of formulas, as it is shown in example 1.
\end{note}

We now see how Truth space operations can be handled with respect to the corresponding formulas. $M (F)$ for a given formula $F$, written in DNF, is nothing but a union of Truth spaces of distinct conjunctive clauses. Truth space of a single conjunctive clause is an intersection of Truth spaces of it's predicates.

\begin{mydef}
Given a logical formula $F$, defined by (\ref{def_F_1}), (\ref{def_F_2}),
(\ref{def_F_3}), the projection of $F$ on the $X$ attribute set is a logical formula $F[X], \<F[X]> = X$, that has all its predicates with $R_i^F.A_i^F \notin X$ replaced with trivial predicate $\n{TRUE}$.
\label{projection}
\end{mydef}

\begin{statement}[Inclusion property]
$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
\label{proj_property_of_inclusion}
\end{statement}
\begin{proof}
If $X = \<F> $, then $F = F[X]$ and $M(F) = M(F[X])$. Assume $X \subset \<F> $.
Consider arbitrary point $a \in M(F)$, i.e. $F(a)=\n{TRUE}$. $F$ to $F[X]$
transformation is made by replacing $T_j$ predicates that contain $A_j \in \<F> 
$ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value. We get
$F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}). Therefore,
$a \in M(F[X])$.
\end{proof}

This property of logical formulas is used while building new data representation
from given IDRs.

\section{Intermediate Data Representation properties research}

We denote saved IDRs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v =
\pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
$s(v)$ ? amount of relations in Database that were used while building $P_v$
IDR, $\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection
with $F_v$ logical formula. We need to get the following representation out of
$P$ IDRs: $$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join
R^{\ast}_2\Join\dots \Join R^{\ast}_l )$$

Lets study the problem of building data representation $P^{\ast}$ using existing
$P_v$ IDRs.

\input{theorem_1}

The conditions given in theorem guarantee the data to build $P^{\ast}$ to be
contained in $P_v$ IDR. However, there can be excess rows that make $F^{\ast}$
take $\n{TRUE}$ value. They appear because there are some relations in
$R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join R^{\ast}_l $ that aren't
presented in $R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} $ and will be
deleted if we join those missing relations. Using Truth spaces of logical
formulas we can query DBMS to get a minimal required data set to define excess
rows.

In the next theorem we describe the case of coinciding relation sets.

\input{theorem_2}

It is possible to use several IRs to build resulting representation. At first,
let's consider the following simple property of natural join.

\begin{statement}
Assume $\Re_1 = R_1 \Join \dots \Join R_k$ - natural join of some $k$ relations.
Also assume $\Re_2 = R_1 \Join \dots \Join R_k \Join R_{k+1} \Join \dots \Join
R_{n}$. Then $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
\label{join_property}
\end{statement}
\begin{proof}
It is so indeed, because joining additional relations to $\Re_1$ can only
remove some rows that were presented there. After the projection, we will get
the same relation $\Re_1$ if no rows were deleted or some subset otherwise. The
order of joins doesn't matter thanks to commutativity property of this
operation.
\end{proof}

\input{theorem_1_mult}

Just like in previous case, there are some additional conditions that allow us
to get precise data presentation from IRs.

\input{theorem_2_mult}

The conditions listed above aren't exceptional in any case, because the set of
dimensions that form hypercube changes rarely, so we can create new dimensions
(data representations) from IRs with much less DB interaction.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE CS typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.
% However, the Computer Society has been known to put floats at the bottom.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE CS papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% However, Computer Society journals sometimes do use bottom floats - bear
% this in mind when choosing appropriate optional arguments for the
% figure/table environments.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

\section{Conclusion}
The prime result of our paper is the justification of IR usage conditions for
new hypercube creation. Those IRs can be stored on an analyst's PC and
drastically shorten the time of getting the required data.

The common problem of systems that don't use any DBMS is the data refreshing. In
MOLAP approach, the technical support staff is responsible for periodical
refreshing of data. The same method can be used to refresh $P_v$
representations. The access to the DMBS's DB modification log and refreshing
only modified data decreases the refresh time period. However, let's notice
again that those dimensions don't change often.

We assume the usage of GPU to implement intermediate filtration and union of
different data representations. The research in this field shows that GPU
operations are much faster than those on CPU for small data. There is a decrease
in performance with data scale. The reason is obvious: big amount of data leads
to a much more frequent data transfer between different memory types and
increases the overall time elapsed. But the constant GPU memory growth and the
fact that data amount required for multidimensional data dimensions building is
comparatively small make the usage of graphical processors quite a good choice
for these purposes.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{2}
%\input{bibliography}
\bibitem{IEEEhowto:kopka}
%This is an example of a book reference
H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.: Addison-Wesley, 1999.
%
%\bibitem{IEEEhowto:coming}
%%This is an example of a Transactions article reference
%D.S. Coming and O.G. Staadt, "Velocity-Aligned Discrete Oriented Polytopes for
%Dynamic Collision Detection," IEEE Trans. Visualization and Computer Graphics,
%vol. 14,  no. 1,  pp. 1-12,  Jan/Feb  2008, doi:10.1109/TVCG.2007.70405.

%This is an example of a article from a conference proceeding
%H. Goto, Y. Hasegawa, and M. Tanaka, "Efficient Scheduling Focusing on the Duality of MPL Representation," Proc. IEEE Symp. Computational Intelligence in Scheduling (SCIS '07), pp. 57-64, Apr. 2007, doi:10.1109/SCIS.2007.367670.

%This is an example of a PrePrint reference
%J.M.P. Martinez, R.B. Llavori, M.J.A. Cabo, and T.B. Pedersen, "Integrating Data Warehouses with Web Data: A Survey," IEEE Trans. Knowledge and Data Eng., preprint, 21 Dec. 2007, doi:10.1109/TKDE.2007.190746.

%Again, see the IEEEtrans_HOWTO.pdf for several more bibliographical examples. Also, more style examples
%can be seen at http://www.computer.org/author/style/transref.htm
\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

% if you will not have a photo at all:
\begin{IEEEbiographynophoto}{John Doe}
Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.
\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.
\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2014-04-23 english_version/IEEE_TKDE/theorem_1.tex
IEEEtran version added

\begin{theorem}
$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
, if:
\\a) $X^{\ast} \subseteq X_{v}$
\\b) 
$ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\}
\subseteq
\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
\\c) $M (F^{\ast}) \subseteq M (F_{v}) $.
\label{th_base}
\end{theorem} 
\begin{proof}
Let $t^{\ast}$ be any tuple in $P^{\ast}$. We need to show that
$t^{\ast} \in \pi_{X^{\ast}} (P_{v})$. From condition  $t^{\ast} \in P^{\ast}$,
there is the tuple $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots
\Join R^{\ast}_l$ and $t^{\ast} = t'[X^{\ast}]$. There may be several such tuples. We will choose the one that satisfies $F^{\ast}(t')=TRUE$.
This tuple exists indeed, because otherwise $t^{\ast}$ would not be in $P^{\ast}$. Thus, there are tuples
$t^{\ast}_i \in R^{\ast}_i$:

\begin{equation}
t^{\ast}_i = t'[\langle R^{\ast}_i\rangle], i = 1,\dots,l,
\label{eval_eq_1}
\end{equation}
\def \intersecij {\langle R^{\ast}_i \rangle \cap \langle R^{\ast}_j \rangle}
and for any pair $i$ and $j$, such that  $\intersecij \neq \emptyset$, equality is held:
\begin{equation}
t^{\ast}_i[\intersecij] = t'_i[\intersecij], i,j = 1,\dots,l,
\label{eval_eq_2}
\end{equation}

Whereas conditions (\ref{eval_eq_1}) and (\ref{eval_eq_2}) are held for the
whole set of relations $\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\}$, they are
true for any its subset also, including $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\}$.
Hence, after joining tuples $t^{\ast}_i$ from relations $R^{v}_j$ we get a tuple
$t''$, such that $t'' = t'[\langle {\Join}_{i=1}^{s(v)} R^v_i \rangle]$.
Equality $t''[X^{\ast}] = t'[X^{\ast}] = t^{\ast}$ follows from a) and $X_v
\subseteq \langle {\Join}_{i=1}^{s(v)} R^v_i \rangle$.

Condition  $F^{\ast}(t') = \n{TRUE}$ and statement 
\ref{proj_property_of_inclusion} imply the truth of the projection $F^{\ast}
[X^{\ast}]$ on a tuple $t'$, and consequently on $t''$, as the formula is defined at common attributes of these tuples. Furthermore, according to condition 3, we have $F_v (t') = \n{TRUE} \Rightarrow F_v (t'') = \n{TRUE}$.
It means $t^{\ast} = t''[X^{\ast}] \in \pi_{X^{\ast}} (P_{v})$. The theorem is
proved.
\end{proof}

### 2014-04-24 english_version/IEEE_TKDE/paper_main.tex
bibliography references created, biography added

@@
 \usepackage[utf8]{inputenc}
+%\usepackage[russian]{babel}
 \usepackage{amsthm}
@@
 \IEEEPARstart{A}{lot} of current Big Data analysis solutions use Online
-Analytical Processing technologies, formed in early 90s \cite{codd}. A lot of
+Analytical Processing technologies, formed in early 90s \cite{codd}. Many
 fundamental \cite{lecht,lehner,mazon} and applied \cite{vassi, peder,
@@
 also made some generalizations on multiple IDRs case.
-
-\hfill mds
-
-\hfill April 22, 2014
 
@@
 Assume the precomputed relations on the user PC:  $P_1$,
-$P_2$, $P_3$ with the following schemes: $P_1$=(Department number, Department), without any data constraints; $P_2$=(Discipline number, Discipline), without data constraints; $P_3$=(Department number, Discipline number, Semester, Class, Hours), with the following constraint:
-$F_3 =(\mbox{Semester} \leq 2) \wedge (\mbox{Department} = \mbox{D. of Chemistry})$.
-Assume the user has made the following logical constraint on the data to be represented in the table: $F^{\ast} = (\mbox{Semester} \leq 3)$.
-Comparing these formulas domains, it's easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1}, whereas $P_3$ needs to be expanded with data.
-After defining $F^{\ast}$ and $F_3$ domains we can form a new formula: $F=((\mbox{Semester} = 3) \wedge (\mbox{Deparmtent=D.
-of Chemistry}))
-\vee ((\mbox{Semester} \leq 3) \wedge (\mbox{Deparmtent} \neq \mbox{D.
-of Chemistry}))$.
-Using $P_3$ schema and $F$ formula, we can now make an SQL command to be executed on the server.
-As a result, we transfer less data over the network.
+$P_2$, $P_3$ with the following schemes: $P_1$=(Department number, Department),
+without any data constraints; $P_2$=(Discipline number, Discipline), without
+data constraints; $P_3$=(Department number, Discipline number, Semester, Class,
+Hours), with the following constraint: $F_3 =(\mbox{Semester} \leq 2) \wedge
+(\mbox{Department} = \mbox{D. of Chemistry})$. Assume the user has made the
+following logical constraint on the data to be represented in the table:
+$F^{\ast} = (\mbox{Semester} \leq 3)$. Comparing these formulas domains, it's
+easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1},
+whereas $P_3$ needs to be expanded with data. After defining $F^{\ast}$ and
+$F_3$ domains we can form a new formula: $F=((\mbox{Semester} = 3) \wedge
+(\mbox{Deparmtent=D. of Chemistry})) \vee ((\mbox{Semester} \leq 3) \wedge
+(\mbox{Deparmtent} \neq \mbox{D. of Chemistry}))$. Using $P_3$ schema and $F$
+formula, we can now make an SQL command to be executed on the server. As a
+result, we transfer less data over the network.
 
@@
 \begin{note}
-The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by software ability to compute Truth spaces of formulas, as it is shown in example 1.
+The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by
+software ability to compute Truth spaces of formulas, as it is shown in example
+1.
 \end{note}
 
-We now see how Truth space operations can be handled with respect to the corresponding formulas.
-$M (F)$ for a given formula $F$, written in DNF, is nothing but a union of Truth spaces of distinct conjunctive clauses.
-Truth space of a single conjunctive clause is an intersection of Truth spaces of it's predicates.
+We now see how Truth space operations can be handled with respect to the
+corresponding formulas. $M (F)$ for a given formula $F$, written in DNF, is
+nothing but a union of Truth spaces of distinct conjunctive clauses. Truth space
+of a single conjunctive clause is an intersection of Truth spaces of it's
+predicates.
 
@@
 Given a logical formula $F$, defined by (\ref{def_F_1}), (\ref{def_F_2}),
-(\ref{def_F_3}), the projection of $F$ on the $X$ attribute set is a logical formula $F[X], \<F[X]> = X$, that has all its predicates with $R_i^F.A_i^F \notin X$ replaced with trivial predicate $\n{TRUE}$.
+(\ref{def_F_3}), the projection of $F$ on the $X$ attribute set is a logical
+formula $F[X], \<F[X]> = X$, that has all its predicates with $R_i^F.A_i^F
+\notin X$ replaced with trivial predicate $\n{TRUE}$.
 \label{projection}
@@
 \bibliographystyle{IEEEtran}
+\bibliography{references}
 % argument is your BibTeX string definitions and bibliography database(s)
@@
 % (used to reserve space for the reference number labels box)
-\begin{thebibliography}{2}
+
+%\begin{thebibliography}{2}
 %\input{bibliography}
-\bibitem{IEEEhowto:kopka}
-%This is an example of a book reference
-H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.: Addison-Wesley, 1999.
-%
+%\bibitem{IEEEhowto:kopka}
+%%This is an example of a book reference
+%H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.:
+%Addison-Wesley, 1999.
+
 %\bibitem{IEEEhowto:coming}
@@
 %can be seen at http://www.computer.org/author/style/transref.htm
-\end{thebibliography}
+%\end{thebibliography}
 
@@
 
-\begin{IEEEbiography}{Michael Shell}
-Biography text here.
-\end{IEEEbiography}
+%\begin{IEEEbiography}{Michael Shell}
+%Biography text here.
+%\end{IEEEbiography}
+
+\newpage
 
 % if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{John Doe}
-Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.
+\begin{IEEEbiographynophoto}{Sergey Mosin}
+is a PhD student in an Omsk Department of Sobolev Institute of Mathematics. His
+research interests include Databases, Data analysis, Algorithm Theory and
+Internet technologies. Mosin received a Specialist degree in Applied Mathematics
+and Computer Science from Omsk F. M. Dostoevsky State University. Contact him at
+svmosin@gmail.com
 \end{IEEEbiographynophoto}
@@
 
-\begin{IEEEbiographynophoto}{Jane Doe}
+\begin{IEEEbiographynophoto}{Sergey Zykin}
 Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.
@@
 
-%\vfill
+\vfill
 

### 2014-04-27 english_version/IEEE_TKDE/paper_main.tex
Chicago corrections. Second reissue

@@
 multidimensional warehouse representation with the use of dynamically formed
-representations. Two basic methods are used there: statical \cite{baralis,
-gupta, gupta-mumick} and dynamical one \cite{scheuermann, shim, kalnis, chang}.
-The former is based on a fixed amount of queries, whereas in the latter queries
+representations. Two basic methods are used: statical \cite{baralis,
+gupta, gupta-mumick} and dynamical \cite{scheuermann, shim, kalnis, chang}.
+The former is based on a fixed amount of queries, whereas in the latter, queries
 for reservation are chosen based on appearance statistic and computational cost
-of query execution. As a data source Data Warehouses or Relational Databases in
-the hierarchical form are used. Special queries for data manipulation are
-considered. In this paper we use split dimension generation for the data
+of query execution. Data Warehouses or Relational Databases in
+the hierarchical form are used as a data source. Special queries for data manipulation are
+considered. In our paper, we use split dimension generation for the data
 required for current analysis. Hence, we are interested in standard SQL queries
@@
 
-In the \cite{Keller96} paper, they analyze the cache content,
+In the \cite{Keller96} paper, the authors analyze the cache content,
 and predicates in their work correspond to IDRs in our approach. Cached data
@@
 \ref{def_T_1}. Dimension attributes are in bold, measure attributes are italic,
-else are attribute values.
+the rest are attribute values.
 
@@
 following logical constraint on the data to be represented in the table:
-$F^{\ast} = (\mbox{Semester} \leq 3)$. Comparing these formulas domains, it's
+$F^{\ast} = (\mbox{Semester} \leq 3)$. Comparing these formulas' domains, it's
 easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1},
@@
 
-Listed variants of operations don't use every single SQL capability. For example
-, we don't use $\n{EXISTS}$ predicate, because it has no expanded attribute
-names in it. $\n{NULL}$ predicate is used for another purpose in our paper.
-
-During logical formula domain calculation if we have some attribute having
-$NULL$ value on a row $t$, we then get the $\n{UNKNOWN}$ value for the whole
+Listed variants of operations don't use every single SQL capability. For
+example, we don't use the $\n{EXISTS}$ predicate, because it has no expanded attribute
+names in it. The $\n{NULL}$ predicate is used for another purpose in our paper.
+
+During logical formula domain calculation, if we have some attribute having
+$NULL$ value on a row $t$, we then get $\n{UNKNOWN}$ value for the whole
 formula, because SQL-query results follow the Three-valued logic. It leads us to
-ambiguous interpretations of result of both usual users and experienced
+ambiguous interpretations of results of both usual users and experienced
 programmers. To solve this problem we suggest the following constraint: every
@@
 , T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
-Then after the modification it will be the following: 
+After the modification it will be the following: 
 $F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here
@@
 $\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic. We
-also can note that if rows don't have undefined values, the initial formula $F$
+can also note that if rows don't have undefined values, the initial formula $F$
 will be equivalent to the transformed one. There can be some cases when the
@@
 .A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$. Then $F$
-transformation on $t$ row will be $\n{TRUE}$, what is not so obvious indeed.
-
-Hereafter we will suppose all the $F$ formulas to be transformed.
+transformation on $t$ row will be $\n{TRUE}$, which is, indeed, not so obvious.
+
+Hereafter we will assume all the $F$ formulas to be transformed.
 
@@
 \begin{note}
-$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case
+$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case,
 attributes that aren't presented in formula can take any values in their
@@
 \begin{note}
-The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by
-software ability to compute Truth spaces of formulas, as it is shown in example
+The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by the
+software's ability to compute Truth spaces of formulas, as it is shown in example
 1.
@@
 
-The conditions given in theorem guarantee the data to build $P^{\ast}$ to be
+The conditions given in the aforementioned theorem guarantee the data to build $P^{\ast}$ to be
 contained in $P_v$ IDR. However, there can be excess rows that make $F^{\ast}$
@@
 
-It is possible to use several IRs to build resulting representation. At first,
+It is possible to use several IRs to build a resulting representation. First,
 let's consider the following simple property of natural join.
@@
 
-Just like in previous case, there are some additional conditions that allow us
+Just like in the previous case, there are some additional conditions that allow us
 to get precise data presentation from IRs.
@@
 The conditions listed above aren't exceptional in any case, because the set of
-dimensions that form hypercube changes rarely, so we can create new dimensions
+dimensions that form a hypercube changes rarely, so we can create new dimensions
 (data representations) from IRs with much less DB interaction.
@@
 The common problem of systems that don't use any DBMS is the data refreshing. In
-MOLAP approach, the technical support staff is responsible for periodical
+the MOLAP approach, the technical support staff is responsible for periodical
 refreshing of data. The same method can be used to refresh $P_v$
 representations. The access to the DMBS's DB modification log and refreshing
-only modified data decreases the refresh time period. However, let's notice
-again that those dimensions don't change often.
+only modified data decreases the refresh time period. However, it should be
+noted again that those dimensions don't change often.
 
@@
 to a much more frequent data transfer between different memory types and
-increases the overall time elapsed. But the constant GPU memory growth and the
-fact that data amount required for multidimensional data dimensions building is
-comparatively small make the usage of graphical processors quite a good choice
-for these purposes.
+increases the overall time elapsed. But the constant GPU memory growth and
+comparatively small data amount required for dimensions building make the usage
+of graphical processors quite a good choice for these purposes.
 

### 2014-04-27 english_version/IEEE_TKDE/theorem_2.tex
Chicago corrections. Second reissue

--- 2014-04-22 english_version/theorem_2.tex
@@
 inclusion $P^{\ast} \subseteq \pi_{X^{\ast}} (P_{v})$ is considered  to be
-proved. It is necessary to show that $\pi_{X^{\ast}} (P_{v}) \subseteq
+proven. It is necessary to show that $\pi_{X^{\ast}} (P_{v}) \subseteq
 P^{\ast}$. Assume tuple $t_v \in \pi_{X^{\ast}} (P_{v})$. Let's show that $t_v
-\in P^{\ast}$. From definition of tuple $t_v$ it follows that there is tuple
+\in P^{\ast}$. From definition of tuple $t_v$, it follows that there is tuple
 $t' \in R^v_1 \Join R^v_2\Join\ldots \Join R^v_{s(v)}$ and $t_v = t'[X^{\ast}]$,
-$F_v(t')=\n{TRUE}$. According to commutativity of natural join, $R^v_1 \Join
+$F_v(t')=\n{TRUE}$. According to the commutativity of natural join, $R^v_1 \Join
 R^v_2\Join\ldots \Join R^v_{s(v)} = R^{\ast}_1 \Join R^{\ast}_2\Join\ldots \Join
@@
 R^{\ast}_l$. From condition c) we have $F^{\ast} (t')=\n{TRUE}$. Hence, $t_v \in
-P^{\ast}$. The theorem is proved.
+P^{\ast}$. The theorem has been proven.
 \end{proof}

### 2014-05-04 english_version/IEEE_TKDE/paper_main.tex
added biography, acknowledgements

@@
 \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. Mosin, S. Zykin are with the
-Omsk Department of Sobolev Institute of Mathematics, Omsk, Russia. The paper is
-supported by RFBR grant \#12-07-00066-a.\protect\\
+the Siberian Branch of the Russian Academy of Sciences, Omsk, Russia.\protect\\
 % note need leading \protect in front of \\ to get a newline within \thanks as
@@
 This approach is supposed to help an analyst, who works with dynamically built
-data representations logically described as multidimensional tables. Those
-representations don't take much disk space, so one can use GPU to speed up the
-computations. We consider queries in a form of relational algebra expressions,
-similar to "universal relational query". Logical formulas in the queries are
+data representations logically described as multidimensional tables. We consider
+queries in a form of relational algebra expressions, similar to "universal
+relational query". Logical formulas in the queries are
 presented in Disjunctive Normal Form. They consist of atomic expressions
@@
 
+% use section* for acknowledgement
+\ifCLASSOPTIONcompsoc
+  % The Computer Society usually uses the plural form
+  \section*{Acknowledgments}
+\else
+  % regular IEEE prefers the singular form
+  \section*{Acknowledgment}
+\fi
+
+The authors would like to thank the Russian Foundation for Basic Research for the
+grant \#12-07-00066-a.
+
 % Can use something like this to put references on a page
@@
 % if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Sergey Mosin}
+\begin{IEEEbiographynophoto}{Sergey Vladimirovich Mosin}
 is a PhD student in an Omsk Department of Sobolev Institute of Mathematics. His
@@
 
-\begin{IEEEbiographynophoto}{Sergey Zykin}
-Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.Biography text here.
+\begin{IEEEbiographynophoto}{Sergey Vladimirovich Zykin}
+is a professor, doctor of science (informatics) with a focus on theory of
+databases designing and head of the laboratory at the Sobolev Institute of
+Mathematics of the Siberian Branch of the Russian Academy of Sciences. He has
+published some 70 scientific papers as conference-, journal-, and book
+contributions. Contact him at szykin@mail.ru
 \end{IEEEbiographynophoto}

### 2014-05-04 english_version/IEEE_TKDE/paper_main.tex
removed GPU sentences from Abstract and Conclusion

@@
 
-We assume the usage of GPU to implement intermediate filtration and union of
-different data representations. The research in this field shows that GPU
-operations are much faster than those on CPU for small data. There is a decrease
-in performance with data scale. The reason is obvious: big amount of data leads
-to a much more frequent data transfer between different memory types and
-increases the overall time elapsed. But the constant GPU memory growth and
-comparatively small data amount required for dimensions building make the usage
-of graphical processors quite a good choice for these purposes.
-
 % if have a single appendix:

### 2014-05-04 english_version/IEEE_TKDE/paper_main.tex
domain word added and middle names provided

@@
 
-\author{Sergey~Mosin,
-        Sergey~Zykin % <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. Mosin, S. Zykin are with the
-the Siberian Branch of the Russian Academy of Sciences, Omsk, Russia.\protect\\
+\author{Sergey~Vladimirovich~Mosin,
+        Sergey~Vladimirovich~Zykin % <-this % stops a space
+\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. V. Mosin, S. V. Zykin are
+with the the Siberian Branch of the Russian Academy of Sciences, Omsk, Russia.
+\protect\\
 % note need leading \protect in front of \\ to get a newline within \thanks as
@@
 This problem was already discussed and has some software implementations. In our
-work, we propose Truth spaces usage in logical formulas while querying a Database.
+work, we propose the usage of logical formulas' domain subsets, called Truth
+spaces, while querying a Database.
 This approach is supposed to help an analyst, who works with dynamically built

### 2014-07-18 english_version/IEEE_TKDE/paper_main.tex
symbols fixed

@@
 \begin{itemize}
-  \item	comparison operation $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ ?
+  \item	comparison operation $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ -
   comparison operator $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$
-  ? type conformant expressions, defined in a space of expanded attribute names and constants;
+  - type conformant expressions, defined in a space of expanded attribute names and constants;
   \item operation $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
   \n{Expr}_3$ (symbols inside square brackets $[*]$ aren't mandatory for the predicate);
-  \item operation $\n{Expr} \n{[NOT]} \n{IN} S$, here $S$ ? values list or subquery, having a set of attribute $R_i.A_j$ values as a result;
-  \item operation $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, here $\n{Str}_i$ ?
+  \item operation $\n{Expr} \n{[NOT]} \n{IN} S$, here $S$ - values list or subquery, having a set of attribute $R_i.A_j$ values as a result;
+  \item operation $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, here $\n{Str}_i$ -
   strings;
@@
 \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
-$s(v)$ ? amount of relations in Database that were used while building $P_v$
+$s(v)$ - amount of relations in Database that were used while building $P_v$
 IDR, $\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection

### 2014-11-11 english_version/IEEE_TKDE/bibliography.tex
Paper version 2.0. Examples updated. Introduction and Conclusion massively refactored.

@@
 \begin{thebibliography}{9}
-\bibitem{codd}
-  E.F. Codd, S.B. Codd, C.T. Salley
-  \textit{Providing OLAP (On-line Analytical Processing) to User-Analysts: An IT
-  Mandate} – Sunnyvale(CA): Codd \& Date Inc.,
-  1993.
-  – 31 p.
-\bibitem{lecht}
-  \textit{Lechtenborger J.}, \textit{Vossen G.}
-  Multidimensional normal forms for data warehouse design
-  // Inf. Syst.
-  – 2003.
-  – Vol. 28, N 5. – P. 415–434. 
-\bibitem{lehner}
-  \textit{Lehner W., Albrecht J., Wedekind H.}
-  Normal forms for multidimensional databases
-  // Proc. of the Tenth Intern. Conf. on Scientific and Statistical Database
-  Management.
-  – Capri, 1998.
-  – P. 63–72.
-\bibitem{mazon}
-  \textit{Mazon J.-N., Trujillo J., Lechtenborger J.}
-  Reconciling requirement-driven data warehouses with data sources via
-  multidimensional normal forms
-  // Data Knowledge Engineering.
-  – 2007.
-  – Vol. 63, N 3. – P. 725–751.
-\bibitem{vassi}
-  \textit{Vassiliadis P., Sellis T.}
-  A survey of logical models for OLAP databases
-  // SIGMOD Rec.
-  – 1999.
-  – Vol. 28, N 4. – P. 64–69.
-\bibitem{peder}
-  \textit{Pedersen T.B., Jensen C.S., Dyreson C.E.}
-  A foundation for capturing and querying complex multidimensional data
-  // Inf. Syst.
-  – 2001.
-  – Vol. 26, N 5. – P. 383–423.
-\bibitem{progressive}
-  \textit{Progressive ranking of range aggregates}
-  / H.-G. Li, et al.
-  // Data \& Knowledge Engineering.
-  – 2007.
-  – Vol. 63, N 1. – P. 4–25. 
-\bibitem{giorg}
-  \textit{Giorgini P., Rizzi S., Garzetti M.}
-  Goal-oriented requirement analysis for data warehouse design
-  // Proc. of the 8th ACM international Workshop on Data Warehousing and OLAP: DOLAP '05.
-  – Bremen, 2005.
-  – P. 47–56.
-
-%ours  
 \bibitem{zyk_pol}
@@
 2013.
-  NO. 5, P. 54-59.
+NO. 5, P. 54--59.
 
-%stream
-\bibitem{Olston03}
-  \textit{Olston C., Jiang J., Widom J.}
-  Adaptive filters for continuous queries over distributed data streams
-  // Proc. of the 2003 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '03).
-  – San Diego, 2003.
-  – P. 563–574.
-\bibitem{Denny05}
-  \textit{Denny M., Franklin M.J.}
-  Predicate result range caching for continuous queries
-  // Proc. of the 2005 ACM SIGMOD Intern. Conf. on Management of Data (SIGMOD '05).
-  – N.-Y., 2005.
-  – P. 646–657.
+\bibitem{chang}
+\textit{Chang-Sup Park, Myoung Ho Kim b, Yoon-Joon Lee}
+Usability-based caching of query results in OLAP systems.
+// The Journal of Systems and Software
+2003.
+Vol. 68, - P. 103--119
 
-%logic
+\bibitem{baralis}
+\textit{Baralis, E., Paraboschi, S., Teniente, E.}
+Materialized view selection in a multidimensional database.
+// Proc. of the 23rd International Conference on Very Large Data Bases,
+Athens, Greece.
+1997.
+- P. 318--329
+
+\bibitem{kalnis}
+\textit{Kalnis, P., Papadias, D.}
+Proxy-server architectures for OLAP.
+// Proceedings of the 2001 ACM SIGMOD International Conference on Management of Data.
+- Santa Barbara, CA, 2001.
+- P. 367--378.
+
 \bibitem{Afrati06}
@@
 // Theoretical Computer Science.
-  – 2006.
- – Vol. 368, N 1–2. – P. 88–123.
+- 2006.
+- Vol. 368.
+- P. 88--123.
 
-%static approach
-\bibitem{baralis}
-  \textit{Baralis, E., Paraboschi, S., Teniente, E.} Materialized view selection in a multidimensional database.
-  // Proc. of the 23rd International Conference on Very Large Data Bases,
-  Athens, Greece.
-  - 1997
-  - P. 318–329
-\bibitem{gupta}
-  \textit{Gupta, H.}
-  Selection of views to materialize in a data warehouse.
-  // Proceedings of the International Conference on Database Theory, Delphi.
-  - Greece, 1997.
-  - P. 98–112.
-\bibitem{gupta-mumick}
-  \textit{Gupta, H., Mumick, I.S.}
-  Selection of views to materialize under a maintenance cost constraint.
-  // Proceedings of the International Conference on Database Theory.
-  - Israel, 1999.
-  - P. 453–470.
+\bibitem{Keller96}
+\textit{Keller M., Basu J.}
+A predicate-based caching scheme for client-server database architectures
+// VLDB Journal.
+- 1996. N 5.
+- P. 35--47.
 
-%dynamic
-\bibitem{scheuermann}
-  \textit{Scheuermann, P., Shim, J., Vingralek R.}
-  WATCHMAN: A data warehouse intelligent cache manager.
-  // Proceedings of the 22nd International Conference on Very Large Data Bases.
-  - Bombay, India, 1996.
-  - P. 51–62.
 \bibitem{shim}
@@
 Dynamic caching of query results for decision support systems.
-  // Proceedings of the 11th International Conference on Scientific and
-  Statistical Database Management.
+// Proceedings of the 11th International Conference on Scientific and Statistical Database Management.
 - Cleveland, OH, 1999.
-  - P. 254–263.
-\bibitem{kalnis}
-  \textit{Kalnis, P., Papadias, D.}
-  Proxy-server architectures for OLAP.
-  // Proceedings of the 2001 ACM SIGMOD International Conference on Management
-  of Data.
-  - Santa Barbara, CA, 2001.
-  - P. 367–378.
-\bibitem{chang}
-  \textit{Chang-Sup Park, Myoung Ho Kim b, Yoon-Joon Lee}
-  Usability-based caching of query results in OLAP systems.
-  // The Journal of Systems and Software
-  - 2003
-  - Vol. 68, - P. 103–119
-
-%domain
-\bibitem{Keller96}
-  \textit{Keller M., Basu J.}
-  A predicate-based caching scheme for client-server database architectures
-  // VLDB Journal. – 1996. N 5. – P. 35–47.
+- P. 254--263.
 

### 2014-11-11 english_version/IEEE_TKDE/paper_main.tex
Paper version 2.0. Examples updated. Introduction and Conclusion massively refactored.

--- 2014-05-04 english_version/IEEE_TKDE/paper_main.tex
@@
 
-\author{Sergey~Vladimirovich~Mosin,
-        Sergey~Vladimirovich~Zykin % <-this % stops a space
+\author{Sergey~V.~Mosin,
+        Sergey~V.~Zykin % <-this % stops a space
 \IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. V. Mosin, S. V. Zykin are
@@
 \markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-{Shell \MakeLowercase{\textit{et al.}}: Building table-like applications using
-cached data}
+{Shell \MakeLowercase{\textit{et al.}}: Truth space method for cached database
+queries}
 % The only time the second header will appear is for the odd numbered pages
@@
 %\boldmath
-We primarily address the usage of cached Database query results in our paper.
-This problem was already discussed and has some software implementations. In our
-work, we propose the usage of logical formulas' domain subsets, called Truth
-spaces, while querying a Database.
-This approach is supposed to help an analyst, who works with dynamically built
-data representations logically described as multidimensional tables. We consider
-queries in a form of relational algebra expressions, similar to "universal
-relational query". Logical formulas in the queries are
-presented in Disjunctive Normal Form. They consist of atomic expressions
-consisting of SQL predicates. In the paper we consider single and multiple table
-cases to build new data representations.
+We propose a new method of client-side data caching for relational databases
+with a central server and distant clients. Data is loaded into client cache
+based on queries executed on the central database at the server. These
+queries have a special form - "universal relational query". The majority of
+search queries to database can be expressed in such form. Besides, this form
+allows us to analyze query correctness by checking lossless join property. A
+subsequent query may be executed in client's local cache if we can determine
+that the query result is entirely contained in the cache. For this we compare
+truth spaces of the logical restrictions in new user's query and results of
+the queries execution in cache. This method may be used to define lacking data
+in cache and execute the query on server for this data only. Problem of data
+actualizations in cache is not discussed in this paper. However, it can be
+solved by cataloging of queries on server and their serving by triggers in
+background mode.
 \end{abstract}
@@
 \begin{keywords}
-Database, OLAP, cache, Data Warehouse, predicate
+Relational databases, Caching, Truth space
 \end{keywords}}
@@
 % and "HIS" in caps to complete the first word.
-\IEEEPARstart{A}{lot} of current Big Data analysis solutions use Online
-Analytical Processing technologies, formed in early 90s \cite{codd}. Many
-fundamental \cite{lecht,lehner,mazon} and applied \cite{vassi, peder,
-progressive, giorg} results were obtained in this field lately.
-
-The purpose of this paper is to study the problem of building Dynamic
-Multidimensional Data Representation from the Relational Database using
-cached data representations on a user computer, equipped with a Graphical
-Processing Unit (GPU). We don't require any transformation of the initial
-Database schema, even to the hierarchical model.
-
-This problem is connected with a field of query optimization, because it aims to
-decrease the data transfer from a Database server. Cached data is actively
-used in Database Management Systems (DBMS), but mostly, it is only repeating use
-of data, written in cache, without any prior data analysis to define any partial
-or combined use. The DBMS can only avoid requesting blocks of data from external
-devices while serving the query, if they are present in the cache. So, the
-block numbers are analyzed, not their information.
-
-This paper is based on the results, obtained in \cite{zyk_pol}. We removed
-condition that limited Intermediate Data Representation (IDR) attributes and
-also made some generalizations on multiple IDRs case.
+\IEEEPARstart{I}n this paper we discuss the use of data views that are
+cached on the client's computer. We assume client-server environment with
+server based Relational Database (RDB). The cache stores query results in order
+to provide maximum usage of saved data in subsequent queries.
+
+The purpose of this paper is to study the problem of building and using data
+views on a client's computer. This problem is similar to query optimization
+because it aims to decrease the data transfer from a Database server. Cached
+data is actively used in Database Management Systems (DBMS), but mostly it is
+only repeating use of data written in cache, without any prior data analysis
+aimed to define any partial or combined use. The DBMS can only avoid requesting
+blocks of data from external devices, while serving the query, if they are
+present in the cache. So, only block numbers are analyzed, not their information.
+
+This paper is based on the results, obtained in \cite{zyk_pol}. We have removed
+a constraint that limits the choice of attributes for Intermediate Data View
+(IDV) and also made some generalizations on multiple IDVs case.
+
+The second section provides an overview of publications on the subject of this
+paper. We have considered only those closest to the problem at hand. In the end
+of this section we present an example to explain our approach for solving the
+problem.
+
+In the third section we provide a formalization of the problem. We also present
+the approach for removing the uncertainties in database queries and present
+auxiliary properties and definitions that are used later.
+
+The main results are presented in Section 4. We research the possibility of
+using cached results of previous queries when performing a new one. The results
+are presented in the form of theorems.
 
 \section{Existing solutions overview and comparison}
-In a field of fundamental researches, huge amounts of publications aim at optimal
-query plan generation based on a formal set of rules that don't use SQL
-operator predicate domains (logical optimization). In other cases, those domains
-are taken into account for statistical evaluations to improve physical access to
-the data. Stream data query executing problems, described in \cite{Olston03,
-Denny05}, are related to our problem closer, but different aims lead us to
-different results. The most similar to our work is paper \cite{Afrati06}. They
-analyze conjunctive queries on data domains with predicates in the form of
-arithmetical comparisons, and present query computation algorithms using IDRs. In
-our paper, the special case of universal relational query is considered. It is a
-query on the Database relations, not particular domains. Although we had similar
-aims, the results obtained differ because of stated factors. In particular,
-there is no need to create any algorithms of data selection in our work, as they
-are replaced by Relational Algebra.
-
-There are some works, which examine the query optimization problem in the case of
-multidimensional warehouse representation with the use of dynamically formed
-representations. Two basic methods are used: statical \cite{baralis,
-gupta, gupta-mumick} and dynamical \cite{scheuermann, shim, kalnis, chang}.
-The former is based on a fixed amount of queries, whereas in the latter, queries
-for reservation are chosen based on appearance statistic and computational cost
-of query execution. Data Warehouses or Relational Databases in
-the hierarchical form are used as a data source. Special queries for data manipulation are
-considered. In our paper, we use split dimension generation for the data
-required for current analysis. Hence, we are interested in standard SQL queries
-to the Database and IDRs. The resulting representation can then be processed by
-special multidimensional data queries.
-
-In the \cite{Keller96} paper, the authors analyze the cache content,
-and predicates in their work correspond to IDRs in our approach. Cached data
-usage problem is resolved in terms of predicate deduction. In our paper, we address this problem by means of
-computation of cached query results domains that correspond to the IDRs. It allows
-us to define IR records that can be used to form a new representation and new SQL
-queries that will let us load missing data from Database server. The following
-example demonstrates suggested approach.
-
-{\bf Example 1.} The task is to receive data representation as in the table
-\ref{def_T_1}. Dimension attributes are in bold, measure attributes are italic,
-the rest are attribute values.
-
-\begin{table*}[h!]
-\renewcommand{\arraystretch}{1.3}
-\caption{\label{def_T_1} Curriculum.}
-\centering
-\begin{tabular}{|c|c||c|c|c|c|}
-\cline{2-6}
-\multicolumn{1}{c|}{} & \multicolumn{1}{|r||}{\bf Semester} & \multicolumn{2}{|c|}{2} & \multicolumn{2}{|c|}{3}  \\
-\cline{2-6}
-\multicolumn{1}{c|}{} & {\bf Discipline} & Physics & Chemistry & Physics & Chemistry \\
-\hline \cline{2-6}
-{\bf Department} & {\bf Class} & {\it Hours} & {\it Hours} & {\it Hours} & {\it Hours} \\
-\hline
- & C-321 & & 36 & 18 & 9 \\
-\cline{2-6}
-D. of Chemistry & C-322 & & 48 & 18 & 9 \\
-\cline{2-6}
- & C-331 & & 24 & 18 & 9 \\
-\hline
- & P-321 & 48 & & 24 & \\
-\cline{2-6}
-D. of Physics & P-322 & 48 & & 24 & \\
-\cline{2-6}
- & P-331 & 64 & & 24 & \\
-\hline
-\end{tabular}
-\end{table*}
-Assume the precomputed relations on the user PC:  $P_1$,
-$P_2$, $P_3$ with the following schemes: $P_1$=(Department number, Department),
-without any data constraints; $P_2$=(Discipline number, Discipline), without
-data constraints; $P_3$=(Department number, Discipline number, Semester, Class,
-Hours), with the following constraint: $F_3 =(\mbox{Semester} \leq 2) \wedge
-(\mbox{Department} = \mbox{D. of Chemistry})$. Assume the user has made the
-following logical constraint on the data to be represented in the table:
-$F^{\ast} = (\mbox{Semester} \leq 3)$. Comparing these formulas' domains, it's
-easy to see that $P_1$ and $P_2$ have enough data to form table \ref{def_T_1},
-whereas $P_3$ needs to be expanded with data. After defining $F^{\ast}$ and
-$F_3$ domains we can form a new formula: $F=((\mbox{Semester} = 3) \wedge
-(\mbox{Deparmtent=D. of Chemistry})) \vee ((\mbox{Semester} \leq 3) \wedge
-(\mbox{Deparmtent} \neq \mbox{D. of Chemistry}))$. Using $P_3$ schema and $F$
-formula, we can now make an SQL command to be executed on the server. As a
-result, we transfer less data over the network.
+One of the main research problems of queries optimization to database is the
+construction of an optimal query plan. Queries are transformed without content
+analysis of the database content and cache. In other cases, this information is
+taken into account for calculation statistical estimations to improve physical
+access to data. These problems are the subject of many studies, but they are
+outside of our approach.
+
+A lot of publications are devoted to the problem of cache content management.
+For example, the heuristic algorithms update the cache with regards to user data
+\cite{chang}, storage of important user queries in a cache
+\cite{baralis}, storage of the cached data on multiple servers \cite{kalnis}.
+This paper solves the problem of the best use of a cache for execution of the
+user's queries. Cache content management and the best use of the cache are two
+complementary problem.
+
+The most similar to our work is paper \cite{Afrati06}. The authors analyze
+conjunctive queries on data domains with predicates in the form of arithmetical
+comparisons, and present query computation algorithms using IDVs. In our paper,
+the special case of universal relational query is considered. It is a query on
+the Database relations, not particular domains. Although we had similar aims,
+the results obtained differ because of stated factors. In particular, there is
+no need to create any algorithms of data selection in our work, as they are
+replaced by Relational Algebra.
+
+In papers \cite{Keller96, shim} similar problem is discussed. In our paper we
+make correspondence between cache contents and predicates. Cached data usage
+problem is resolved in terms of truth spaces. We compute the truth spaces of
+query results in the cache. It allows us to define records in IDV that can be
+used to form a new view and new SQL queries that will let us load missing data
+from Database server. The following example demonstrates suggested approach.
+
+{\bf Example 1.} Let's assume the following database schema fragment, which
+represents the University Study plan:\\
+$R_1 =\mbox{\it Students}\ (\mbox{\bf Stud\_ID}, \mbox{Stud\_Name},
+\mbox{Group})$\\
+$R_2 =\mbox{\it Schedule}\ (\mbox{\bf Group}, \mbox{\bf Room\_ID},
+\mbox{Course})$\\
+$R_3 =\mbox{\it Progress}\ (\mbox{\bf Stud\_ID}, \mbox{\bf Course},
+\mbox{Score})$,\\
+relation names are italic, Primary Key attributes are in bold. Assume that on 
+he user's computer the following queries are cached:\\
+Query 1: List of students studying physics, whose ID is bigger than 210:
+$$P_1 = \pi_{X_1}(\sigma_{F_1} (R_1 \Join R_2)),$$
+where $\pi_{X_1}$ -- projection operation over the set of attributes
+$X_1 = \{\text{Stud\_ID}, \text{Stud\_Name}, \text{Group}, \text{Course}\}$,
+$\sigma$ -- selecting operation,
+$F_1$ -- logical formula: $(\text{Stud\_ID} > 210\ \&\ \text{Course} =
+\text{Physics})$, $\Join$ --natural join.
+
+Query 2: Examination sheets of the group M10:
+$$P_2 = \pi_{X_2}(\sigma_{F_2} (R_1 \Join R_3)),$$
+where
+$X_2 = \{\text{Stud\_ID}, \text{Group}, \text{Course}, \text{Score}\}$, $F_2$
+-- logical condition: $(\text{Group}=\text{M10})$.
+
+Let us assume that the user has requested information formalized with the
+following query:
+$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R_1 \Join R_2 \Join R_3 )),$$
+where $X^{\ast}= \{\text{Stud\_ID}, \text{Group}, \text{Score}\}$, $F^{\ast}$
+-- logical condition: $(\text{Stud\_ID} > 300\ \&\ \text{Group} =
+\text{M10}\ \&\ \text{Course} = \text{Physics})$.
+
+Using the calculation and comparison of the truth domains $P_1$, $P_2$ and
+$P^{\ast}$, we obtain that the query can be executed in the cache:
+$P^{\ast} =\pi_{X^{\ast}}(\sigma_{F_3} (P_1 \Join P_2 ))$, where $F_3$
+-- logical condition: $(\text{Stud\_ID} > 300)$. Requesting the server in this
+case is not required.
 
@@
 
-To simplify domain computations we will consider logical formulas in Disjunctive Normal Form (DNF).
-In general case formula $F$ has the following form:
+To simplify domain computations we will consider logical formulas in
+Disjunctive Normal Form (DNF). In general case formula $F$ has the following
+form:
 
@@
 \begin{equation}
-K_i = T_1 \& T_2 \dots \& T_n, i = 1, \dots, m ,
+K_i = T_1 \&\ T_2 \dots \&\ T_n, i = 1, \dots, m ,
 \label{def_F_2}
 \end{equation}
-here $T_j, j = 1, \dots, n$ - predicates, where expanded attribute names are specified. $R_i.A_j$ means attribute $A_j$ in relation $R_i$.
+here $T_j, j = 1, \dots, n$ - predicates, where expanded attribute names are
+specified. $R_i.A_j$ means attribute $A_j$ in relation $R_i$.
 Those predicates can be:
 \begin{itemize}
-  \item	comparison operation $ \n{Expr}_1 \theta \n{Expr}_2$, $\theta$ ?
-  comparison operator $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$
-  ? type conformant expressions, defined in a space of expanded attribute names and constants;
-  \item operation $\n{Expr}_1 \n{[NOT]} \n{BETWEEN} \n{Expr}_2 \n{AND}
-  \n{Expr}_3$ (symbols inside square brackets $[*]$ aren't mandatory for the predicate);
-  \item operation $\n{Expr} \n{[NOT]} \n{IN} S$, here $S$ ? values list or subquery, having a set of attribute $R_i.A_j$ values as a result;
-  \item operation $\n{Str}_1 \n{[NOT]} \n{LIKE} \n{Str}_2$, here $\n{Str}_i$ ?
-  strings;
-  \item operation $\n{Expr} \theta \n{ALL/ANY} S$.
+\item comparison operation $ \n{Expr}_1\ \theta\ \n{Expr}_2\ $, $\theta$
+-- comparison operator $(\theta \in \{=, \neq, >, <, \leq, \geq\})$,
+$\n{Expr}_i$ -- type conformant expressions, defined in a space of expanded
+attribute names and constants;
+\item operation $\n{Expr}_1\ \n{[NOT]}\ \n{BETWEEN}\ \n{Expr}_2\ \n{AND}\
+\n{Expr}_3$
+(symbols inside square brackets $[*]$ are optional);
+\item operation $\n{Expr}\ \n{[NOT]}\ \n{IN}\ S$, here $S$ - values list or
+subquery, having a set of attribute $R_i.A_j$ values as a result;
+\item operation $\n{Str}_1\ \n{[NOT]}\ \n{LIKE}\ \n{Str}_2$, here $\n{Str}_i$
+-- strings;
+\item operation $\n{Expr}\ \theta\ \n{ALL/ANY}\ S$.
 \end{itemize}
@@
 \begin{note}
-We assume logical formulas have no trivial conditions on attributes, for example
-, $Expr_1 = Expr_1$ and those reduced to such form. In general, we assume that
-$R_i.A_l$ domain is not fully contained in $T_j (\dots , R_i . A_l ,\dots )$
-predicate truth space. Such conditions can be removed from the formula without a
-change in the truth space (we will define it later).
+We assume logical formulas have no trivial conditions on attributes, for
+example, $Expr_1 = Expr_1$ and those reduced to such form. In general, we assume
+that $R_i.A_l$ domain is not fully contained in
+$T_j (\dots , R_i . A_l ,\dots )$ predicate truth space. Such conditions can be
+removed from the formula without a change in the truth space (we will define it
+later).
 \label{trivial}
@@
 \begin{equation}
-\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\}
+\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\} .
 \label{def_F_3}
@@
 
-Listed variants of operations don't use every single SQL capability. For
-example, we don't use the $\n{EXISTS}$ predicate, because it has no expanded attribute
+Listed variants of operations don't use every single SQL capability. For example,
+we don't use the $\n{EXISTS}$ predicate, because it has no expanded attribute
 names in it. The $\n{NULL}$ predicate is used for another purpose in our paper.
@@
 During logical formula domain calculation, if we have some attribute having 
-$NULL$ value on a row $t$, we then get $\n{UNKNOWN}$ value for the whole
+$NULL$ value on a tuple $t$, we then get $\n{UNKNOWN}$ value for the whole
 formula, because SQL-query results follow the Three-valued logic. It leads us to
@@
 attribute in $F^{\ast}$ is supplied with a property ``Use of undefined value''
-with two mutually exclusive values: ``Yes'' or ``No''. The reasoning behind
-this property is the following: if it is assigned ``Yes'', then we leave rows
-with $\n{NULL}$ value for further consideration. Otherwise, having ``No'' in
-that property guarantees us removing all such rows.
+with two mutually exclusive values: ``Yes'' or ``No''. The reasoning behind this
+property is the following: if it is assigned ``Yes'', then we leave tuples with
+$\n{NULL}$ value for further consideration. Otherwise, having ``No'' in that
+property guarantees us removing all such tuples.
 
 Let's write expression (\ref{def_F_1}) for $F$ in the following form: $ F (\dots
-, T_j , \dots )$, here $T_j$ - predicates of expression (\ref{def_F_2}).
-After the modification it will be the following: 
-$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, here
-$\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$ - conjunction of all $F$ attributes,for
-which $\n{NULL}$ is not allowed, and $T'_j = (T_j \vee_{i,j}(R_i .A_j =
-\n{NULL}))$, here  $\vee_{i,j}(R_i .A_j = \n{NULL})$ - disjunction of all $F$
-attributes, for which $\n{NULL}$ is allowed. Outer brackets for $T'_j$ predicate
-define operation priority. We can see that this logical formula can only have
-$\n{TRUE}$ and $\n{FALSE}$ values when considering it in Three-valued logic. We
-can also note that if rows don't have undefined values, the initial formula $F$
-will be equivalent to the transformed one. There can be some cases when the
-meaning of $P$ IR changes with such a transformation. For example, assume $F =
-R_1 .A_2 > 3 \vee R_3 .A_4 < 4$. Let row $t$ has $R_1 .A_2$ value set to
-$\n{NULL}$, with a property ``Use of undefined value'' set to ``Yes'', and $R_3
-.A_4$ value set to 5, so $R_3 .A_4 < 4$ predicate is $\n{FALSE}$. Then $F$
-transformation on $t$ row will be $\n{TRUE}$, which is, indeed, not so obvious.
+, T_j , \dots )$, here $T_j$ -- predicates of expression (\ref{def_F_2}). After
+the modification it will be the following: $F( \dots, T'_j , \dots )\wedge_{i,j}
+(R_i .A_j \neq \n{NULL})$, here $\wedge_{i,j}(R_i .A_j \neq \n{NULL})$
+-- conjunction of all $F$ attributes,for which $\n{NULL}$ is not allowed, and
+$T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, here $\vee_{i,j}(R_i .A_j =
+\n{NULL})$ -- disjunction of all $F$ attributes, for which $\n{NULL}$ is allowed.
+Outer brackets for $T'_j$ predicate define operation priority. We can see that
+this logical formula can only have $\n{TRUE}$ and $\n{FALSE}$ values when
+considering it in Three-valued logic. We can also note that if tuples don't have
+undefined values, the initial formula $F$ will be equivalent to the transformed
+one, so semantics of view $P$ almost undistorted. For the disclosure of the term
+"almost" we consider example. Assume $F = R_1.A_2 > 3 \vee R_3.A_4 < 4$. Let
+tuple $t$ for $R_1.A_2$ has value $\n{NULL}$, with a property ``Use of undefined
+value'' equal to ``Yes'', and $R_3.A_4$ value equal to 5, so value of
+$R_3.A_4 < 4$ is $\n{FALSE}$. Then transform of formula $F$ on tuple $t$ will be
+equal $\n{TRUE}$, which is not obvious.
 
@@
 
-We consider $\mathcal{A} =$ $\{(a_1, \dots, a_n)
-\mid a_i \in Dom(A_i), i=1,\dots,n\}$ set, here $Dom(A_i)$ - is a domain of
-$A_i$. Cartesian product $Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$
-is an $n$-dimensional space of all values for all Database attributes. Data
-constraints bound this space to some set of points that represents available
-Database states. In the Note \ref{trivial} we consider the whole space, not only
-the available states.
+We consider set $\mathcal{A} =$ $\{(a_1, \dots, a_n) \mid a_i \in Dom(A_i),
+i=1,\dots,n\}$, here $Dom(A_i)$ -- is a domain of $A_i$. Cartesian product
+$Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ is an $n$-dimensional
+space of all values for all Database attributes. Data constraints bound this
+space to some set of points that represents a set of available Database
+states.
+
+In our example with University Study plan, the $\mathcal{A}$ can be the
+following:
+
+\begin{multline*}
+\mathcal{A} = Dom(\text{Stud\_ID})\times Dom(\text{Stud\_Name}) \times\\
+\times Dom(\text{Group}) \times Dom(\text{Room\_ID})
+\times Dom(\text{Course}) \times\\
+\times Dom(\text{Score}) = \{205, 315, 461\}\times\{\text{Rachel Davis, Noam}\\
+\text{Angrist, Cameron McCord}\}
+\times\{\text{M10, M11}\} \times \{100, 101,\\
+102, 103\} \times \{\text{Physics, Chemistry}\} \times \{70, 80, 90\}
+\end{multline*}
+
+Obviously, without constraints this set represents more states than DB can have.
+For example, each student has his/her own ID, whereas there are 3 corresponding
+IDs for each student in $\mathcal{A}$: $\{205, 315, 461\}\times \{\text{Rachel
+Davis, Noam Angrist, Cameron McCord}\}$.
+
+\begin{note}
+Dimension of formula $F$ can be smaller than dimension of $\mathcal{A}$. In this
+case, we consider the equivalent form of the formula that has all other
+attributes taking any values in their domains.
+\end{note}
+
+Looking back to our Example 1, we see that the formula in Query 1 contains only
+two attributes: $F_1 = (\text{Stud\_ID} > 210\ \&\ \text{Course} =
+\text{Physics})$, in other words $\<F> = 2$. The equivalent representation that
+we will use is the following:
+
+\begin{multline*}
+F_1 = (\text{Stud\_ID} > 210\ \&\
+\text{Course} = \text{Physics}\ \&\\
+\text{Stud\_Name}\ \textit{IN}\ \{\text{Rachel Davis, Noam Angrist, Cameron}\\
+\text{McCord}\}\ \&\ \text{Group}\ \textit{IN}\ \{\text{M10, M11}\}\ \&\
+\text{Room\_ID}\ \textit{IN}\\
+\{100, 101, 102, 103\}\ \&\ \text{Score}\ \textit{IN}\
+\{70, 80, 90\})
+\end{multline*}
+
+The motivation behind this transformation is to allow us comparing truth spaces
+of formulas defined on different attributes.
 
@@
 
-\begin{note}
-$F$ formula dimension can be smaller than $\mathcal{A}$ dimension. In this case,
-attributes that aren't presented in formula can take any values in their
-domains.
-\end{note}
+So, the truth space of our formula $F_1$ will be the following subset of
+$\mathcal{A}$: $M (F) = $ \{(315, Rachel Davis, M10, 100, Physics, 70), (315,
+Noam Angrist, M10, 100, Physics, 70), \dots (461, Cameron McCord, M10, 100,
+Physics, 90)\}
+
+$M (F)$ for a given formula $F$, written in DNF, is nothing but a union of Truth
+spaces of distinct conjunctive clauses. Truth space of a single conjunctive
+clause is an intersection of Truth spaces of it's predicates.
 
@@
 The complexity of $\n{Expr}$, $S$ and $\n{Str}$ predicates is defined by the
-software's ability to compute Truth spaces of formulas, as it is shown in example
-1.
+software's ability to compute Truth spaces of formulas, as it is shown in
+example 1.
 \end{note}
-
-We now see how Truth space operations can be handled with respect to the
-corresponding formulas. $M (F)$ for a given formula $F$, written in DNF, is
-nothing but a union of Truth spaces of distinct conjunctive clauses. Truth space
-of a single conjunctive clause is an intersection of Truth spaces of it's
-predicates.
 
@@
 
+So, the projection of $F_1$ on the attribute set $X = \{\text{Stud\_ID}\}$ will be
+$F_1[\text{Stud\_ID}] = (\text{Stud\_ID} > 210\ \&\ \text{TRUE}) = (\text{Stud\_ID} >
+210)$
+
 \begin{statement}[Inclusion property]
@@
 Consider arbitrary point $a \in M(F)$, i.e. $F(a)=\n{TRUE}$. $F$ to $F[X]$
-transformation is made by replacing $T_j$ predicates that contain $A_j \in \<F> 
-$ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value. We get
+transformation is made by replacing $T_j$ predicates that contain $A_j \in
+\<F> $ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value. We get
 $F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}). Therefore,
@@
 
-This property of logical formulas is used while building new data representation
-from given IDRs.
-
-\section{Intermediate Data Representation properties research}
-
-We denote saved IDRs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v =
-\pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$,
-$s(v)$ ? amount of relations in Database that were used while building $P_v$
-IDR, $\pi_{X_v}$ - projection on $X_v$ attributes, $\sigma_{F_v}$ - selection
-with $F_v$ logical formula. We need to get the following representation out of
-$P$ IDRs: $$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join
-R^{\ast}_2\Join\dots \Join R^{\ast}_l )$$
-
-Lets study the problem of building data representation $P^{\ast}$ using existing
-$P_v$ IDRs.
+This property of logical formulas is used while building new data view from given
+IDVs.
+
+\section{Intermediate Data View properties research}
+
+We denote saved IDVs as $P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$, here $P_v$ is an
+IDV, $P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join
+R^v_{s(v)} ))$, $s(v)$ -- amount of relations in Database that were used while
+building $P_v$,
+$\pi_{X_v}$ -- projection on $X_v$ attributes, $\sigma_{F_v}$ -- selection
+with $F_v$ logical formula. We need to get the following view out of $P$ IDVs:
+$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}}
+(R^{\ast}_1 \Join R^{\ast}_2\Join\dots \Join R^{\ast}_l )$$
+Let’s study the problem of building data view $P^{\ast}$ using existing $P_v$ IDVs.
 
@@
 
-The conditions given in the aforementioned theorem guarantee the data to build $P^{\ast}$ to be
-contained in $P_v$ IDR. However, there can be excess rows that make $F^{\ast}$
+The conditions given in the aforementioned theorem guarantee the data to build
+$P^{\ast}$ to be contained in $P_v$ IDV. However, there can be excess tuples that
+make $F^{\ast}$
 take $\n{TRUE}$ value. They appear because there are some relations in
@@
 formulas we can query DBMS to get a minimal required data set to define excess
-rows.
+tuples.
 
@@
 
-It is possible to use several IRs to build a resulting representation. First,
+It is possible to use several IDVs to build a resulting representation. First,
 let's consider the following simple property of natural join.
@@
 It is so indeed, because joining additional relations to $\Re_1$ can only
-remove some rows that were presented there. After the projection, we will get
-the same relation $\Re_1$ if no rows were deleted or some subset otherwise. The
-order of joins doesn't matter thanks to commutativity property of this
-operation.
+remove some tuples that were presented there. After the projection, we will get
+the same relation $\Re_1$ if no tuples were deleted or some subset otherwise. The
+order of joins doesn't matter because of commutativity of this operation.
 \end{proof}
@@
 
-Just like in the previous case, there are some additional conditions that allow us
-to get precise data presentation from IRs.
+Just like in the previous case, there are some additional conditions that allow
+us to get precise data presentation from IDVs.
 
@@
 
-The conditions listed above aren't exceptional in any case, because the set of
-dimensions that form a hypercube changes rarely, so we can create new dimensions
-(data representations) from IRs with much less DB interaction.
-
+The conditions listed above aren't exceptional in any case. Many applications
+require repeat of data entry with slight modifications. For example,
+applications that
+work with multidimensional data. Hypercube dimensions do not change often, so
+that we can create new hypercubes from IDVs without frequent interaction with
+the database.
 % An example of a floating figure using the graphicx package.
@@
 \section{Conclusion}
-The prime result of our paper is the justification of IR usage conditions for
-new hypercube creation. Those IRs can be stored on an analyst's PC and
-drastically shorten the time of getting the required data.
-
-The common problem of systems that don't use any DBMS is the data refreshing. In
-the MOLAP approach, the technical support staff is responsible for periodical
-refreshing of data. The same method can be used to refresh $P_v$
-representations. The access to the DMBS's DB modification log and refreshing
-only modified data decreases the refresh time period. However, it should be
-noted again that those dimensions don't change often.
+The results obtained will be used in the technology of dynamic building of
+multidimensional data. In addition, they are intended to be used with graphic
+processors (GPU) to perform the intermediate filtering operations and compound
+of different data representations. Researches in this field have shown that the
+operations performed on GPU are much faster than on CPU for small amounts of
+data. Performance decreases with increasing amounts of data. The reason is
+obvious: as long as the data is put in fast memory of GPU the operations with
+them are performed fast. When increasing the amount of data, it is necessary to
+move the data to different type of memory, which reduces productivity. However,
+GPU
+manufacturers are constantly expanding fast memory. At the same time, dimensions
+of multidimensional data require relatively small memory, so GPUs are suitable
+in solving this problem.
+
+Multidimensional data building systems that use redundant data often face the
+problem of updating it. It is often solved by the periodic updating of the
+hypercube contents.
+A similar method can be used to update the views $P_v$. To reduce the update
+time it might be a good practice to use the change log on the database server
+and update only those views that have original data changed. However, it should
+be noted again that those dimensions don't change often.
 
@@
 
-The authors would like to thank the Russian Foundation for Basic Research for the
-grant \#12-07-00066-a.
+The authors would like to thank the Russian Foundation for Basic Research for
+the grant \#12-07-00066-a.
 
@@
 is a professor, doctor of science (informatics) with a focus on theory of
-databases designing and head of the laboratory at the Sobolev Institute of
+Databases designing and head of the laboratory at the Sobolev Institute of
 Mathematics of the Siberian Branch of the Russian Academy of Sciences. He has

### 2014-11-15 english_version/IEEE_TKDE/paper_main.tex
IEEE version slightly modified

@@
 % can use linebreaks \\ within to get better formatting as desired
-\title{Building table-like applications using cached data}
+\title{Truth space method for caching database queries}
 %
@@
 \markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-{Shell \MakeLowercase{\textit{et al.}}: Truth space method for cached database
+{Shell \MakeLowercase{\textit{et al.}}: Truth space method for caching database
 queries}
@@
 % references section
-
 % can use a bibliography generated by BibTeX as a .bbl file
@@
 \begin{IEEEbiographynophoto}{Sergey Vladimirovich Mosin}
-is a PhD student in an Omsk Department of Sobolev Institute of Mathematics. His
+is a PhD student in the Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences. His
 research interests include Databases, Data analysis, Algorithm Theory and

### 2014-11-15 english_version/TODS/paper_main.tex
added version for Transactions on Database Systems

--- 2014-11-15 english_version/IEEE_TKDE/paper_main.tex
@@
-%% bare_jrnl_compsoc.tex
-%% V1.3
-%% 2007/01/11
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.7 or later) with an IEEE Computer
-%% Society journal paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/tex-archive/macros/latex/contrib/IEEEtran/
-%% and
-%% http://www.ieee.org/
+% This document is created with a help of
+% v2-acmsmall-sample.tex sample file, v1.3 - March 2012
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. IEEE's font choices can trigger bugs that do  ***
-% *** not appear when using other class files.                            ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-% Note that the a4paper option is mainly intended so that authors in
-% countries using A4 can easily print to A4 and see how their papers will
-% look in print - the typesetting of the document will not typically be
-% affected with changes in paper size (but the bottom and side margins will).
-% Use the testflow package mentioned above to verify correct handling of
-% both paper sizes by the user's LaTeX system.
-%
-% Also note that the "draftcls" or "draftclsnofoot", not "draft", option
-% should be used if it is desired that the figures are to be displayed in
-% draft mode.
-%
-% The Computer Society usually requires 10pt for submissions.
-%
-\documentclass[10pt,journal,cspaper,compsoc]{IEEEtran}
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[12pt,journal,compsoc]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/oberdiek/
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-\ifCLASSOPTIONcompsoc
-  % IEEE Computer Society needs nocompress option
-  % requires cite.sty v4.0 or later (November 2003)
-  % \usepackage[nocompress]{cite}
-\else
-  % normal IEEE
-  % \usepackage{cite}
-\fi
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 4.0 (2003-05-27) and later if using hyperref.sty. cite.sty does
-% not currently provide for hyperlinked citations.
-% The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/cite/
-% The documentation is contained in the cite.sty file itself.
-%
-% Note that some packages require special options to format as the Computer
-% Society requires. In particular, Computer Society  papers do not use
-% compressed citation ranges as is done in typical IEEE papers
-% (e.g., [1]-[4]). Instead, they list every citation separately in order
-% (e.g., [1], [2], [3], [4]). To get the latter we need to load the cite
-% package with the nocompress option which is supported by cite.sty v4.0
-% and later. Note also the use of a CLASSOPTION conditional provided by
-% IEEEtran.cls V1.7 and later.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
-\ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
-\else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
-\fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at: 
-% http://www.ctan.org/tex-archive/macros/latex/required/graphics/
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found as epslatex.ps or
-% epslatex.pdf at: http://www.ctan.org/tex-archive/info/
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage[cmex10]{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics. If using
-% it, be sure to load this package with the cmex10 option to ensure that
-% only type 1 fonts will utilized at all point sizes. Without this option,
-% it is possible that some math symbols, particularly those within
-% footnotes, will be rendered in bitmap form which will result in a
-% document that can not be IEEE Xplore compliant!
-%
-% Also, note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/amslatex/math/
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithms/
-% There is also a support site at:
-% http://algorithms.berlios.de/index.html
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/algorithmicx/
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/required/tools/
-
-%\usepackage{mdwmath}
-%\usepackage{mdwtab}
-% Also highly recommended is Mark Wooding's extremely powerful MDW tools,
-% especially mdwmath.sty and mdwtab.sty which are used to format equations
-% and tables, respectively. The MDWtools set is already installed on most
-% LaTeX systems. The lastest version and documentation is available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/mdwtools/
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-%\usepackage{eqparbox}
-% Also of notable interest is Scott Pakin's eqparbox package for creating
-% (automatically sized) equal width boxes - aka "natural width parboxes".
-% Available at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/eqparbox/
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%\usepackage[tight,normalsize,sf,SF]{subfigure}
-%\else
-%\usepackage[tight,footnotesize]{subfigure}
-%\fi
-% subfigure.sty was written by Steven Douglas Cochran. This package makes it
-% easy to put subfigures in your figures. e.g., "Figure 1a and 1b". For IEEE
-% work, it is a good idea to load it with the tight package option to reduce
-% the amount of white space around the subfigures. Computer Society papers
-% use a larger font and \sffamily font for their captions, hence the
-% additional options needed under compsoc mode. subfigure.sty is already
-% installed on most LaTeX systems. The latest version and documentation can
-% be obtained at:
-% http://www.ctan.org/tex-archive/obsolete/macros/latex/contrib/subfigure/
-% subfigure.sty has been superceeded by subfig.sty.
-
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false]{caption}
-%  \usepackage[font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false]{caption}
-%  \usepackage[font=footnotesize]{subfig}
-%\fi
-% subfig.sty, also written by Steven Douglas Cochran, is the modern
-% replacement for subfigure.sty. However, subfig.sty requires and
-% automatically loads Axel Sommerfeldt's caption.sty which will override
-% IEEEtran.cls handling of captions and this will result in nonIEEE style
-% figure/table captions. To prevent this problem, be sure and preload
-% caption.sty with its "caption=false" package option. This is will preserve
-% IEEEtran.cls handing of captions. Version 1.3 (2005/06/28) and later 
-% (recommended due to many improvements over 1.2) of subfig.sty supports
-% the caption=false option directly:
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-%
-% The latest version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/subfig/
-% The latest version and documentation of caption.sty can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/caption/
-
-% *** FLOAT PACKAGES ***
-%
-\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure. The latest version and documentation can be found at:
-% http://www.ctan.org/tex-archive/macros/latex/base/
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/sttools/
-% Documentation is contained in the stfloats.sty comments as well as in the
-% presfull.pdf file. Do not use the stfloats baselinefloat ability as IEEE
-% does not allow \baselineskip to stretch. Authors submitting work to the
-% IEEE should note that IEEE rarely uses double column equations and
-% that authors should try to avoid such use. Do not be tempted to use the
-% cuted.sty or midfloat.sty packages (also by Sigitas Tolusis) as IEEE does
-% not format its papers in such ways.
-
-%\ifCLASSOPTIONcaptionsoff
-%  \usepackage[nomarkers]{endfloat}
-% \let\MYoriglatexcaption\caption
-% \renewcommand{\caption}[2][\relax]{\MYoriglatexcaption[#2]{#2}}
-%\fi
-% endfloat.sty was written by James Darrell McCauley and Jeff Goldberg.
-% This package may be useful when used in conjunction with IEEEtran.cls'
-% captionsoff option. Some IEEE journals/societies require that submissions
-% have lists of figures/tables at the end of the paper and that
-% figures/tables without any captions are placed on a page by themselves at
-% the end of the document. If needed, the draftcls IEEEtran class option or
-% \CLASSINPUTbaselinestretch interface can be used to increase the line
-% spacing as well. Be sure and use the nomarkers option of endfloat to
-% prevent endfloat from "marking" where the figures would have been placed
-% in the text. The two hack lines of code above are a slight modification of
-% that suggested by in the endfloat docs (section 8.3.1) to ensure that
-% the full captions always appear in the list of figures/tables - even if
-% the user used the short optional argument of \caption[]{}.
-% IEEE papers do not typically make use of \caption[]'s optional argument,
-% so this should not be an issue. A similar trick can be used to disable
-% captions of packages such as subfig.sty that lack options to turn off
-% the subcaptions:
-% For subfig.sty:
-% \let\MYorigsubfloat\subfloat
-% \renewcommand{\subfloat}[2][\relax]{\MYorigsubfloat[]{#2}}
-% For subfigure.sty:
-% \let\MYorigsubfigure\subfigure
-% \renewcommand{\subfigure}[2][\relax]{\MYorigsubfigure[]{#2}}
-% However, the above trick will not work if both optional arguments of
-% the \subfloat/subfig command are used. Furthermore, there needs to be a
-% description of each subfigure *somewhere* and endfloat does not add
-% subfigure captions to its list of figures. Thus, the best approach is to
-% avoid the use of subfigure captions (many IEEE journals avoid them anyway)
-% and instead reference/explain all the subfigures within the main caption.
-% The latest version of endfloat.sty and its documentation can obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/endfloat/
-%
-% The IEEEtran \ifCLASSOPTIONcaptionsoff conditional can also be used
-% later in the document, say, to conditionally put the References on a 
-% page by themselves.
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version can be obtained at:
-% http://www.ctan.org/tex-archive/macros/latex/contrib/misc/
-% Read the url.sty source comments for usage information. Basically,
-% \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
-\hyphenation{op-tical net-works semi-conduc-tor}
+\documentclass[prodmode,acmtods]{acmsmall} % Aptara syntax
 
@@
 \usepackage{amssymb,amsmath}
-\usepackage[utf8]{inputenc}
-%\usepackage[russian]{babel}
-\usepackage{amsthm}
-\usepackage{amsfonts}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-\usepackage{xcolor}
-\usepackage{textcomp}
-\newtheorem{theorem}{Theorem}
-\newtheorem{mydef}{Definition}
+%\usepackage{amsfonts}
+%\usepackage{algorithm}
+%\usepackage{xcolor}
+%\usepackage{textcomp}
+\newtheorem{note}{Note}
 \newtheorem{statement}{Statement}
-\newtheorem*{consequence}{Consequence}
-\newtheorem{note}{Note}
 \def \eval #1#2{\left.#1\right\vert_{#2}}
@@
 \def \n #1{\mathit{#1}}
+
+% Package to generate and customize Algorithm as per ACM style
+\usepackage[ruled]{algorithm2e}
+\renewcommand{\algorithmcfname}{ALGORITHM}
+\SetAlFnt{\small}
+\SetAlCapFnt{\small}
+\SetAlCapNameFnt{\small}
+\SetAlCapHSkip{0pt}
+\IncMargin{-\parindent}
+
+% Metadata Information
+\acmVolume{9}
+\acmNumber{4}
+\acmArticle{39}
+\acmYear{2010}
+\acmMonth{3}
+
+% Document starts
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
+
+% Page heads
+\markboth{S.~V. Mosin, S.~V. Zykin}{Truth space method for caching database
+queries}
+
+% Title portion
 \title{Truth space method for caching database queries}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-%
-%\IEEEcompsocitemizethanks is a special \thanks that produces the bulleted
-% lists the Computer Society journals use for "first footnote" author
-% affiliations. Use \IEEEcompsocthanksitem which works much like \item
-% for each affiliation group. When not in compsoc mode,
-% \IEEEcompsocitemizethanks becomes like \thanks and
-% \IEEEcompsocthanksitem becomes a line break with idention. This
-% facilitates dual compilation, although admittedly the differences in the
-% desired content of \author between the different types of papers makes a
-% one-size-fits-all approach a daunting prospect. For instance, compsoc 
-% journal papers have the author affiliations above the "Manuscript
-% received ..."  text while in non-compsoc journals this is reversed. Sigh.
+\author{SERGEY V. MOSIN
+\affil{Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences}
+SERGEY V. ZYKIN
+\affil{Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences}
+}
 
-\author{Sergey~V.~Mosin,
-        Sergey~V.~Zykin % <-this % stops a space
-\IEEEcompsocitemizethanks{\IEEEcompsocthanksitem S. V. Mosin, S. V. Zykin are
-with the the Siberian Branch of the Russian Academy of Sciences, Omsk, Russia.
-\protect\\
-% note need leading \protect in front of \\ to get a newline within \thanks as
-% \\ is fragile and will error, could use \hfil\break instead.
-E-mail: svmosin@gmail.com, szykin@mail.ru}% <-this % stops a space
-\thanks{}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-{Shell \MakeLowercase{\textit{et al.}}: Truth space method for caching database
-queries}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% The publisher's ID mark at the bottom of the page is less important with
-% Computer Society journal papers as those publications place the marks
-% outside of the main text columns and, therefore, unlike regular IEEE
-% journals, the available text space is not reduced by their presence.
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
-% or like this to get the Computer Society new two part style.
-%\IEEEpubid{\makebox[\columnwidth]{\hfill 0000--0000/00/\$00.00~\copyright~2007 IEEE}%
-%\hspace{\columnsep}\makebox[\columnwidth]{Published by the IEEE Computer Society\hfill}}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark (Computer Society jorunal
-% papers don't need this extra clearance.)
-
-% for Computer Society papers, we must declare the abstract and index terms
-% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
-% command as these need to go into the title area created by \maketitle.
-\IEEEcompsoctitleabstractindextext{%
 \begin{abstract}
-%\boldmath
 We propose a new method of client-side data caching for relational databases
@@
 \end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the journal you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals frown on math
-% in the abstract anyway. In particular, the Computer Society does
-% not want either math or citations to appear in the abstract.
 
-% Note that keywords are not normally used for peer review papers.
-\begin{keywords}
-Relational databases, Caching, Truth space
-\end{keywords}}
+\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
 
-% make the title area
+\terms{Caching, Optimization}
+
+\keywords{Relational database, Truth space, Intermediate Data View, Logical
+formula}
+
+\acmformat{Sergey V. Mosin and Sergey V. Zykin, 2014. Truth space method for
+caching database queries}
+
+\begin{bottomstuff}
+This work is supported by the Russian Foundation for Basic Research under
+grant \#12-07-00066-a.
+
+Author's addresses: S.V. Mosin, Laboratory of information transformation and
+representations methods, Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences; S.V. Zykin, Laboratory of information
+transformation and representations methods, Sobolev Institute of Mathematics of
+the Siberian Branch of the Russian Academy of Sciences.
+\end{bottomstuff}
+
 \maketitle
 
-% To allow for easy dual compilation without having to reenter the
-% abstract/keywords data, the \IEEEcompsoctitleabstractindextext text will
-% not be used in maketitle, but will appear (i.e., to be "transported")
-% here as \IEEEdisplaynotcompsoctitleabstractindextext when compsoc mode
-% is not selected <OR> if conference mode is selected - because compsoc
-% conference papers position the abstract like regular (non-compsoc)
-% papers do!
-\IEEEdisplaynotcompsoctitleabstractindextext
-% \IEEEdisplaynotcompsoctitleabstractindextext has no effect when using
-% compsoc under a non-conference mode.
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-\IEEEpeerreviewmaketitle
-
 \section{Introduction}
-% Computer Society journal papers do something a tad strange with the very
-% first section heading (almost always called "Introduction"). They place it
-% ABOVE the main text! IEEEtran.cls currently does not do this for you.
-% However, You can achieve this effect by making LaTeX jump through some
-% hoops via something like:
-%
-%\ifCLASSOPTIONcompsoc
-%  \noindent\raisebox{2\baselineskip}[0pt][0pt]%
-%  {\parbox{\columnwidth}{\section{Introduction}\label{sec:introduction}%
-%  \global\everypar=\everypar}}%
-%  \vspace{-1\baselineskip}\vspace{-\parskip}\par
-%\else
-%  \section{Introduction}\label{sec:introduction}\par
-%\fi
-%
-% Admittedly, this is a hack and may well be fragile, but seems to do the
-% trick for me. Note the need to keep any \label that may be used right
-% after \section in the above as the hack puts \section within a raised box.
-
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps (small caps for compsoc).
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-\IEEEPARstart{I}n this paper we discuss the use of data views that are
+In this paper we discuss the use of data views that are
 cached on the client's computer. We assume client-server environment with
@@
 (IDV) and also made some generalizations on multiple IDVs case.
-
+\cite{codd}
 The second section provides an overview of publications on the subject of this
@@
 
+% Head 1
 \section{Existing solutions overview and comparison}
@@
 relation names are italic, Primary Key attributes are in bold. Assume that on 
-he user's computer the following queries are cached:\\
+he user's computer the following queries are cached:
+
 Query 1: List of students studying physics, whose ID is bigger than 210:
@@
 
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
-
 \section{Logical constraint definition and properties}
-
 To simplify domain computations we will consider logical formulas in
@@
 
-\begin{mydef}
+\begin{definition}
 Space of attributes contained in a formula show the dimension of the formula and
@@
 \end{equation}
-\end{mydef}
+\end{definition}
 
@@
 \begin{multline*}
-\mathcal{A} = Dom(\text{Stud\_ID})\times Dom(\text{Stud\_Name}) \times\\
-\times Dom(\text{Group}) \times Dom(\text{Room\_ID})
-\times Dom(\text{Course}) \times\\
-\times Dom(\text{Score}) = \{205, 315, 461\}\times\{\text{Rachel Davis, Noam}\\
-\text{Angrist, Cameron McCord}\}
-\times\{\text{M10, M11}\} \times \{100, 101,\\
-102, 103\} \times \{\text{Physics, Chemistry}\} \times \{70, 80, 90\}
+\mathcal{A} = Dom(\text{Stud\_ID})\times Dom(\text{Stud\_Name}) \times
+Dom(\text{Group}) \times Dom(\text{Room\_ID}) \times\\
+\times Dom(\text{Course}) \times Dom(\text{Score}) = \{205, 315, 461\} \times
+\{\text{Rachel Davis, Noam Angrist},\\
+\text{Cameron McCord}\} \times\{\text{M10, M11}\} \times \{100, 101, 102, 103\}
+\times \{\text{Physics, Chemistry}\} \times\\ \times \{70, 80, 90\}
 \end{multline*}
@@
 \begin{multline*}
-F_1 = (\text{Stud\_ID} > 210\ \&\
-\text{Course} = \text{Physics}\ \&\\
-\text{Stud\_Name}\ \textit{IN}\ \{\text{Rachel Davis, Noam Angrist, Cameron}\\
-\text{McCord}\}\ \&\ \text{Group}\ \textit{IN}\ \{\text{M10, M11}\}\ \&\
-\text{Room\_ID}\ \textit{IN}\\
-\{100, 101, 102, 103\}\ \&\ \text{Score}\ \textit{IN}\
-\{70, 80, 90\})
+F_1 = (\text{Stud\_ID} > 210\ \&\ \text{Course} = \text{Physics}\ \&\
+\text{Stud\_Name}\ \textit{IN}\ \{\text{Rachel Davis, Noam}\\
+\text{Angrist, Cameron McCord}\}\ \&\ \text{Group}\ \textit{IN}\ \{\text{M10, M11}\}\ \&\
+\text{Room\_ID}\ \textit{IN} \{100, 101, 102,\\
+103\}\ \&\ \text{Score}\ \textit{IN}\ \{70, 80, 90\})
 \end{multline*}
@@
 
-\begin{mydef}
+\begin{definition}
 Truth space of logical formula $F$, defined by (\ref{def_F_1}),
@@
 F(a) = \n {TRUE}\}$.
-\end{mydef}
+\end{definition}
 
@@
 
-\begin{mydef}
+\begin{definition}
 Given a logical formula $F$, defined by (\ref{def_F_1}), (\ref{def_F_2}),
@@
 \label{projection}
-\end{mydef}
+\end{definition}
 
@@
 $\pi_{X_v}$ -- projection on $X_v$ attributes, $\sigma_{F_v}$ -- selection
-with $F_v$ logical formula. We need to get the following view out of $P$ IDVs:
+with $F_v$ logical formula. Eventually we need to get the following view:
 $$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}}
-(R^{\ast}_1 \Join R^{\ast}_2\Join\dots \Join R^{\ast}_l )$$
+(R^{\ast}_1 \Join R^{\ast}_2\Join\dots \Join R^{\ast}_l ))$$
 Let’s study the problem of building data view $P^{\ast}$ using existing $P_v$ IDVs.
@@
 the database.
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE CS typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-% However, the Computer Society has been known to put floats at the bottom.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE CS papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals use top floats exclusively.
-% However, Computer Society journals sometimes do use bottom floats - bear
-% this in mind when choosing appropriate optional arguments for the
-% figure/table environments.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
-% stfloats package.
 
@@
 processors (GPU) to perform the intermediate filtering operations and compound
-of different data representations. Researches in this field have shown that the
-operations performed on GPU are much faster than on CPU for small amounts of
+of different data representations. Researches in this field show that the
+operations performed using GPU are much faster than using CPU for small amounts of
 data. Performance decreases with increasing amounts of data. The reason is
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
+% Bibliography
+\bibliographystyle{ACM-Reference-Format-Journals}
+\bibliography{references}
+                             % Sample .bib file with references that match those in
+                             % the 'Specifications Document (V1.5)' as well containing
+                             % 'legacy' bibs and bibs with 'alternate codings'.
+                             % Gerry Murray - March 2012
 
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
+% History dates
+\received{February 2007}{March 2009}{June 2009}
 
-% use section* for acknowledgement
-\ifCLASSOPTIONcompsoc
-  % The Computer Society usually uses the plural form
-  \section*{Acknowledgments}
-\else
-  % regular IEEE prefers the singular form
-  \section*{Acknowledgment}
-\fi
+% Electronic Appendix
+\elecappendix
 
-The authors would like to thank the Russian Foundation for Basic Research for
-the grant \#12-07-00066-a.
+\medskip
 
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
-
-%\begin{thebibliography}{2}
-%\input{bibliography}
-%\bibitem{IEEEhowto:kopka}
-%%This is an example of a book reference
-%H. Kopka and P.W. Daly, \emph{A Guide to {\LaTeX}}, third ed. Harlow, U.K.:
-%Addison-Wesley, 1999.
-
-%\bibitem{IEEEhowto:coming}
-%%This is an example of a Transactions article reference
-%D.S. Coming and O.G. Staadt, "Velocity-Aligned Discrete Oriented Polytopes for
-%Dynamic Collision Detection," IEEE Trans. Visualization and Computer Graphics,
-%vol. 14,  no. 1,  pp. 1-12,  Jan/Feb  2008, doi:10.1109/TVCG.2007.70405.
-
-%This is an example of a article from a conference proceeding
-%H.
-Goto, Y.
-Hasegawa, and M.
-Tanaka, "Efficient Scheduling Focusing on the Duality of MPL Representation," Proc.
-IEEE Symp.
-Computational Intelligence in Scheduling (SCIS '07), pp.
-57-64, Apr.
-2007, doi:10.1109/SCIS.2007.367670.
-
-%This is an example of a PrePrint reference
-%J.M.P.
-Martinez, R.B.
-Llavori, M.J.A.
-Cabo, and T.B.
-Pedersen, "Integrating Data Warehouses with Web Data: A Survey," IEEE Trans.
-Knowledge and Data Eng., preprint, 21 Dec.
-2007, doi:10.1109/TKDE.2007.190746.
-
-%Again, see the IEEEtrans_HOWTO.pdf for several more bibliographical examples. Also, more style examples
-%can be seen at http://www.computer.org/author/style/transref.htm
-%\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Michael Shell}
-%Biography text here.
-%\end{IEEEbiography}
-
-\newpage
-
-% if you will not have a photo at all:
-\begin{IEEEbiographynophoto}{Sergey Vladimirovich Mosin}
-is a PhD student in the Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences. His
-research interests include Databases, Data analysis, Algorithm Theory and
-Internet technologies. Mosin received a Specialist degree in Applied Mathematics
-and Computer Science from Omsk F. M. Dostoevsky State University. Contact him at
-svmosin@gmail.com
-\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-\begin{IEEEbiographynophoto}{Sergey Vladimirovich Zykin}
-is a professor, doctor of science (informatics) with a focus on theory of
-Databases designing and head of the laboratory at the Sobolev Institute of
-Mathematics of the Siberian Branch of the Russian Academy of Sciences. He has
-published some 70 scientific papers as conference-, journal-, and book
-contributions. Contact him at szykin@mail.ru
-\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}
+% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM

### 2014-11-16 english_version/TODS/blind_review/blind_main.tex
Version for blind review was added.

--- 2014-11-15 english_version/TODS/paper_main.tex
@@
 
-\author{Mosin Sergey, Zykin Sergey}
 \usepackage{amssymb,amsmath}
@@
 
-% Page heads
-\markboth{S.~V. Mosin, S.~V. Zykin}{Truth space method for caching database
-queries}
-
 % Title portion
 \title{Truth space method for caching database queries}
-\author{SERGEY V. MOSIN
-\affil{Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences}
-SERGEY V. ZYKIN
-\affil{Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences}
-}
 
@@
 
-\category{C.2.2}{Computer-Communication Networks}{Network Protocols}
+\category{H.2.3, H.2.4}{Database Management}{Languages, Systems}
 
-\terms{Caching, Optimization}
+\terms{Languages, Theory, Performance}
 
@@
 formula}
-
-\acmformat{Sergey V. Mosin and Sergey V. Zykin, 2014. Truth space method for
-caching database queries}
-
-\begin{bottomstuff}
-This work is supported by the Russian Foundation for Basic Research under
-grant \#12-07-00066-a.
-
-Author's addresses: S.V. Mosin, Laboratory of information transformation and
-representations methods, Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences; S.V. Zykin, Laboratory of information
-transformation and representations methods, Sobolev Institute of Mathematics of
-the Siberian Branch of the Russian Academy of Sciences.
-\end{bottomstuff}
 
@@
 
-In papers \cite{Keller96, shim} similar problem is discussed. In our paper we
+In papers \cite{Keller96}, \cite{shim} similar problem is discussed. In our
+paper we
 make correspondence between cache contents and predicates. Cached data usage
@@
 
-% Electronic Appendix
-\elecappendix
-
-\medskip
-
 \end{document}
-% End of v2-acmsmall-sample.tex (March 2012) - Gerry Murray, ACM

### 2014-12-06 english_version/Elsevier/paper_main.tex
Elsevier version added; gitignore updated.

--- 2014-11-16 english_version/TODS/blind_review/blind_main.tex
@@
-% This document is created with a help of
-% v2-acmsmall-sample.tex sample file, v1.3 - March 2012
+%% 
+%% Copyright 2007, 2008, 2009 Elsevier Ltd
+%% 
+%% This file is part of the 'Elsarticle Bundle'.
+%% ---------------------------------------------
+%% 
+%% It may be distributed under the conditions of the LaTeX Project Public
+%% License, either version 1.2 of this license or (at your option) any
+%% later version.  The latest version of this license is in
+%%    http://www.latex-project.org/lppl.txt
+%% and version 1.2 or later is part of all distributions of LaTeX
+%% version 1999/12/01 or later.
+%% 
+%% The list of all files belonging to the 'Elsarticle Bundle' is
+%% given in the file `manifest.txt'.
+%%
 
-\documentclass[prodmode,acmtods]{acmsmall} % Aptara syntax
+%% Template article for Elsevier's document class `elsarticle'
+%% with numbered style bibliographic references
+%% SP 2008/03/01
 
+\documentclass[preprint,12pt]{elsarticle}
+
+%% Use the option review to obtain double line spacing
+%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
+
+%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
+%% for a journal layout:
+%% \documentclass[final,1p,times]{elsarticle}
+%% \documentclass[final,1p,times,twocolumn]{elsarticle}
+%% \documentclass[final,3p,times]{elsarticle}
+%% \documentclass[final,3p,times,twocolumn]{elsarticle}
+%% \documentclass[final,5p,times]{elsarticle}
+%% \documentclass[final,5p,times,twocolumn]{elsarticle}
+
+%% For including figures, graphicx.sty has been loaded in
+%% elsarticle.cls. If you prefer to use the old commands
+%% please give \usepackage{epsfig}
+
+%% The amssymb package provides various useful mathematical symbols
+\author{Mosin Sergey, Zykin Sergey}
 \usepackage{amssymb,amsmath}
+\usepackage{amsthm}
 %\usepackage{amsfonts}
@@
 \newtheorem{note}{Note}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
 \newtheorem{statement}{Statement}
@@
 \def \n #1{\mathit{#1}}
+%% The amsthm package provides extended theorem environments
+%% \usepackage{amsthm}
 
-% Package to generate and customize Algorithm as per ACM style
-\usepackage[ruled]{algorithm2e}
-\renewcommand{\algorithmcfname}{ALGORITHM}
-\SetAlFnt{\small}
-\SetAlCapFnt{\small}
-\SetAlCapNameFnt{\small}
-\SetAlCapHSkip{0pt}
-\IncMargin{-\parindent}
+%% The lineno packages adds line numbers. Start line numbering with
+%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
+%% for the whole article with \linenumbers.
+%% \usepackage{lineno}
 
-% Metadata Information
-\acmVolume{9}
-\acmNumber{4}
-\acmArticle{39}
-\acmYear{2010}
-\acmMonth{3}
+\journal{Theoretical Computer Science}
 
-% Document starts
 \begin{document}
 
-% Title portion
+\begin{frontmatter}
+
+%% Title, authors and addresses
+
+%% use the tnoteref command within \title for footnotes;
+%% use the tnotetext command for theassociated footnote;
+%% use the fnref command within \author or \address for footnotes;
+%% use the fntext command for theassociated footnote;
+%% use the corref command within \author for corresponding author footnotes;
+%% use the cortext command for theassociated footnote;
+%% use the ead command for the email address,
+%% and the form \ead[url] for the home page:
+%% \title{Title\tnoteref{label1}}
+%% \tnotetext[label1]{}
+%% \author{Name\corref{cor1}\fnref{label2}}
+%% \ead{email address}
+%% \ead[url]{home page}
+%% \fntext[label2]{}
+%% \cortext[cor1]{}
+%% \address{Address\fnref{label3}}
+%% \fntext[label3]{}
+
 \title{Truth space method for caching database queries}
+
+%% use optional labels to link authors explicitly to addresses:
+%% \author[label1,label2]{}
+%% \address[label1]{}
+%% \address[label2]{}
+
+\author{S.~V. Mosin, S.~V. Zykin}
+
+\address{Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences}
 
@@
 
-\category{H.2.3, H.2.4}{Database Management}{Languages, Systems}
+\begin{keyword}
+%% keywords here, in the form: keyword  keyword
 
-\terms{Languages, Theory, Performance}
+%% PACS codes here, in the form: \PACS code \sep code
 
-\keywords{Relational database, Truth space, Intermediate Data View, Logical
-formula}
+%% MSC codes here, in the form: \MSC code \sep code
+%% or \MSC[2008] code \sep code (2000 is the default)
+Relational databases \sep Caching \sep Truth space
+\end{keyword}
 
-\maketitle
+\end{frontmatter}
 
+%% \linenumbers
+
+%% main text
 \section{Introduction}
@@
 
-In papers \cite{Keller96}, \cite{shim} similar problem is discussed. In our
-paper we
+In papers \cite{Keller96}, \cite{shim} similar problem is discussed. In our paper we
 make correspondence between cache contents and predicates. Cached data usage
@@
 where $\pi_{X_1}$ -- projection operation over the set of attributes
-$X_1 = \{\text{Stud\_ID}, \text{Stud\_Name}, \text{Group}, \text{Course}\}$,
+$X_1 = \{\text{Stud\_ID},\\
+\text{Stud\_Name}, \text{Group}, \text{Course}\}$,
 $\sigma$ -- selecting operation,
@@
 $X_2 = \{\text{Stud\_ID}, \text{Group}, \text{Course}, \text{Score}\}$, $F_2$
--- logical condition: $(\text{Group}=\text{M10})$.
+-- logical condition:\\
+$(\text{Group}=\text{M10})$.
 
@@
 \mathcal{A} = Dom(\text{Stud\_ID})\times Dom(\text{Stud\_Name}) \times
-Dom(\text{Group}) \times Dom(\text{Room\_ID}) \times\\
-\times Dom(\text{Course}) \times Dom(\text{Score}) = \{205, 315, 461\} \times
-\{\text{Rachel Davis, Noam Angrist},\\
-\text{Cameron McCord}\} \times\{\text{M10, M11}\} \times \{100, 101, 102, 103\}
-\times \{\text{Physics, Chemistry}\} \times\\ \times \{70, 80, 90\}
+Dom(\text{Group}) \times\\
+Dom(\text{Room\_ID}) \times Dom(\text{Course}) \times Dom(\text{Score}) =
+\{205, 315, 461\} \times \{\text{Rachel}\\
+\text{Davis, Noam Angrist}, \text{Cameron McCord}\} \times\{\text{M10, M11}\}
+\times \{100, 101, 102,\\
+103\} \times \{\text{Physics, Chemistry}\} \times \{70, 80, 90\}
 \end{multline*}
@@
 IDs for each student in $\mathcal{A}$: $\{205, 315, 461\}\times \{\text{Rachel
-Davis, Noam Angrist, Cameron McCord}\}$.
+Davis},\\
+\text{Noam Angrist, Cameron McCord}\}$.
 
@@
 F_1 = (\text{Stud\_ID} > 210\ \&\ \text{Course} = \text{Physics}\ \&\
-\text{Stud\_Name}\ \textit{IN}\ \{\text{Rachel Davis, Noam}\\
-\text{Angrist, Cameron McCord}\}\ \&\ \text{Group}\ \textit{IN}\ \{\text{M10, M11}\}\ \&\
-\text{Room\_ID}\ \textit{IN} \{100, 101, 102,\\
-103\}\ \&\ \text{Score}\ \textit{IN}\ \{70, 80, 90\})
+\text{Stud\_Name}\ \textit{IN}\ \{\text{Rachel Davis},\\
+\text{Noam Angrist, Cameron McCord}\}\ \&\ \text{Group}\ \textit{IN}\ \{\text{M10, M11}\}\ \&\
+\text{Room\_ID}\ \textit{IN}\\ \{100, 101, 102, 103\}\ \&\ \text{Score}\ \textit{IN}\ \{70, 80, 90\})
 \end{multline*}
@@
 
-% Bibliography
-\bibliographystyle{ACM-Reference-Format-Journals}
+\bibliographystyle{elsarticle-num}
 \bibliography{references}
-                             % Sample .bib file with references that match those in
-                             % the 'Specifications Document (V1.5)' as well containing
-                             % 'legacy' bibs and bibs with 'alternate codings'.
-                             % Gerry Murray - March 2012
-
-% History dates
-\received{February 2007}{March 2009}{June 2009}
 
 \end{document}
+\endinput

### 2015-02-22 english_version/Elsevier/paper_main.tex
Bibliography file in TKDE version was removed. Acknowledgements were added to TODS version.
Conclusion modified: GPU references were removed. Work summarization was added.

@@
 \section{Conclusion}
-The results obtained will be used in the technology of dynamic building of
-multidimensional data. In addition, they are intended to be used with graphic
-processors (GPU) to perform the intermediate filtering operations and compound
-of different data representations. Researches in this field show that the
-operations performed using GPU are much faster than using CPU for small amounts of
-data. Performance decreases with increasing amounts of data. The reason is
-obvious: as long as the data is put in fast memory of GPU the operations with
-them are performed fast. When increasing the amount of data, it is necessary to
-move the data to different type of memory, which reduces productivity. However,
-GPU
-manufacturers are constantly expanding fast memory. At the same time, dimensions
-of multidimensional data require relatively small memory, so GPUs are suitable
-in solving this problem.
+The approach suggested in this paper is a theoretical base for a system
+interacting with Relational Database. Such system will be capable of defining
+cached data availability when executing consecutive queries. This solution is
+fresh and wasn't used before. The conditions stated in the theorems don't
+require hitting Database to be checked. Results obtained allow us to
+analitycally define data missing in cache and query only this data.
 
+The proposed solution will be used in dynamic building of multidimensional data.
 Multidimensional data building systems that use redundant data often face the
-problem of updating it. It is often solved by the periodic updating of the
+problem of updating it. It is sometimes solved by the periodic updating of the
 hypercube contents.
@@
 time it might be a good practice to use the change log on the database server
-and update only those views that have original data changed. However, it should
-be noted again that those dimensions don't change often.
+and update only those views that have original data changed. However, those
+dimensions don't change often.
+
+\section{Acknowledgements}
+This work is supported by the Russian Foundation for Basic Research under
+grant \#12-07-00066-a.
 

### 2015-03-01 english_version/MAIS/paper_main.tex
MAIS version was added. Codd OLAP citation was removed from all versions.

--- 2015-02-22 english_version/Elsevier/paper_main.tex
@@
-%% 
-%% Copyright 2007, 2008, 2009 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
+\documentclass{maiseng}
 
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
+\author{Mosin Sergey, Zykin Sergey}
 
-\documentclass[preprint,12pt]{elsarticle}
-
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
-\author{Mosin Sergey, Zykin Sergey}
+\usepackage[cp1251]{inputenc}
+\usepackage[english,russian]{babel}
+\usepackage[T2A]{fontenc}
 \usepackage{amssymb,amsmath}
 \usepackage{amsthm}
-%\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage{xcolor}
-%\usepackage{textcomp}
+
+\selectlanguage{english}
+
+%%
+\date{April 15, }\ydate{2013}
+\firstpage{175}
+
+\renewcommand{\refname}{References}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Macros used in the paper
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \newtheorem{note}{Note}
@@
 \def \n #1{\mathit{#1}}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
 
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
+\addto\captionsrussian{\renewcommand*{\proofname}{Proof}}
 
-\journal{Theoretical Computer Science}
+%%%%%%%%%%%%%%%%%% end my macros %%%%%%%%%%%%%%%%%%%
+
+%%
+\title{Truth space method for caching database queries}
+\author{Mosin~S.~V., Zykin~S.~V.\footnote{Supported by the Russian Foundation for Basic Research under
+grant \#12-07-00066-a.}}
+\authorcr{Mosin~S.~V., Zykin~S.~V.}
+\keywords{Relational databases, Caching, Truth space}
+\rutitle{Кэширование запросов к реляционной базе данных с использованием областей истинности}
+\ruauthor{Мосин С.В., Зыкин С.В.}
+
+\rukeywords{реляционная база данных, кэш, область истинности}
+
+\UDC{519.987} \shorttitle{Truth space method for caching database queries}
+\affiliation{\small Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences, \\Pevtsova str., 13, Omsk, 644043, Russia}
+\ruaffiliation{\small Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук \\644043, Россия, г.
+Омск,
+ул. Певцова, 13,}
+%%%%
+\email{svmosin@gmail.com, szykin@mail.ru}
+\authorinform{
+\textbf{Мосин Сергей Владимирович},\\Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук, аспирант\\
+\textbf{Зыкин Сергей Владимирович},\\Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук, д-р техн.
+наук, профессор
+}
+
+\ruabstract{
+В данной статье предлагается новый метод кэширования запросов к реляционной базе данных
+для систем с центральным сервером и распределенными клиентами. Данные загружаются в
+клиентский кэш, основываясь на запросах, выполненных на сервере БД.
+Эти запросы имеют специальный вид, называемый "универсальный реляционный запрос".
+Большинство запросов может быть записано в этом виде.
+Кроме того, эта форма записи позволяет анализировать корректность запроса, проверяя свойство соединения без потери информации (СБПИ).
+Последовательные запросы могут исполняться на клиенте, используя кэш, если удастся определить, что результаты искомого запроса полностью содержатся в кэше.
+Для осуществления такой проверки анализируются области истинности логических ограничений искомого запроса и запросов, результаты которых уже содержатся в кэше.
+Этот метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.
+Проблема актуализации данных не затрагивается в этой статье.
+Однако она может быть решена путем учета запросов на сервере и обновлении данных при помощи триггеров.}
 
 \begin{document}
-
-\begin{frontmatter}
-
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-\title{Truth space method for caching database queries}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \address[label1]{}
-%% \address[label2]{}
-
-\author{S.~V. Mosin, S.~V. Zykin}
-
-\address{Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences}
+\maketitle
 
@@
 with a central server and distant clients. Data is loaded into client cache
-based on queries executed on the central database at the server. These
-queries have a special form - "universal relational query". The majority of
+based on queries executed on the server. These
+queries have a special form called "universal relational query". The majority of
 search queries to database can be expressed in such form. Besides, this form
@@
 
-\begin{keyword}
-%% keywords here, in the form: keyword  keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-Relational databases \sep Caching \sep Truth space
-\end{keyword}
-
-\end{frontmatter}
-
-%% \linenumbers
-
-%% main text
-\section{Introduction}
+\section *{Introduction}
 In this paper we discuss the use of data views that are
@@
 (IDV) and also made some generalizations on multiple IDVs case.
-\cite{codd}
 The second section provides an overview of publications on the subject of this
@@
 
-% Head 1
 \section{Existing solutions overview and comparison}
@@
 
-\section{Acknowledgements}
-This work is supported by the Russian Foundation for Basic Research under
-grant \#12-07-00066-a.
-
-\bibliographystyle{elsarticle-num}
+\renewcommand{\refname}{References}
 \bibliography{references}

### 2015-03-01 english_version/MAIS/theorem_1.tex
MAIS version was added. Codd OLAP citation was removed from all versions.

--- 2014-04-23 english_version/IEEE_TKDE/theorem_1.tex
@@
 \begin{theorem}
-$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$
-, if:
+$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$, if:
 \\a) $X^{\ast} \subseteq X_{v}$
-\\b) 
-$ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\}
-\subseteq
-\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
+\\b) $ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} \subseteq \{R^{\ast}_{1}, \ldots,
+R^{\ast}_{l}\} $
 \\c) $M (F^{\ast}) \subseteq M (F_{v}) $.
@@
 there is the tuple $t' \in R^{\ast}_1 \Join R^{\ast}_2\Join\ldots
-\Join R^{\ast}_l$ and $t^{\ast} = t'[X^{\ast}]$. There may be several such tuples. We will choose the one that satisfies $F^{\ast}(t')=TRUE$.
-This tuple exists indeed, because otherwise $t^{\ast}$ would not be in $P^{\ast}$. Thus, there are tuples
-$t^{\ast}_i \in R^{\ast}_i$:
+\Join R^{\ast}_l$ and $t^{\ast} = t'[X^{\ast}]$. There may be several such tuples.
+We will choose the one that satisfies $F^{\ast}(t')=TRUE$. This tuple exists
+indeed, because otherwise $t^{\ast}$ would not be in $P^{\ast}$. Thus, there
+are tuples $t^{\ast}_i \in R^{\ast}_i$:
 
@@
 \ref{proj_property_of_inclusion} imply the truth of the projection $F^{\ast}
-[X^{\ast}]$ on a tuple $t'$, and consequently on $t''$, as the formula is defined at common attributes of these tuples.
-Furthermore, according to condition 3, we have $F_v (t') = \n{TRUE} \Rightarrow F_v (t'') = \n{TRUE}$.
-It means $t^{\ast} = t''[X^{\ast}] \in \pi_{X^{\ast}} (P_{v})$. The theorem is
-proved.
+[X_v]$ on a tuple $t'$, and consequently on $t''$, as the formula is defined at common attributes of these tuples.
+Furthermore, according to condition c), we have $F_v (t') = \n{TRUE} \Rightarrow F_v (t'') = \n{TRUE}$.
+It means $t^{\ast} = t''[X^{\ast}] \in \pi_{X^{\ast}} (P_{v})$.
+The theorem has been proven.
 \end{proof}

### 2015-03-01 english_version/MAIS/paper_main.tex
MAIS version modified

@@
 %%
-\date{April 15, }\ydate{2013}
-\firstpage{175}
+\date{March 1, }\ydate{2015}
+%\firstpage{175}
 
@@
 \authorinform{
-\textbf{Мосин Сергей Владимирович},\\Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
-С.Л.~Соболева Сибирского отделения Российской академии наук, аспирант\\
-\textbf{Зыкин Сергей Владимирович},\\Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
-С.Л.~Соболева Сибирского отделения Российской академии наук, д-р техн.
+\textbf{Мосин Сергей Владимирович},\\Федеральное государственне бюджетное учреждение науки Институт математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал), аспирант\\
+\textbf{Зыкин Сергей Владимирович},\\Федеральное государственне бюджетное учреждение науки Институт математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал), д-р техн.
 наук, профессор
@@
 \renewcommand{\refname}{References}
-\bibliography{references}
+\begin{thebibliography}{1}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{zyk_pol}
+S.~Zykin, A.~Poluyanov, Multidimensional data building using intermediate
+  representations., Administration problems~(5) (2013) 54--59.
+
+\bibitem{chang}
+C.-S. Park, M.-H. Kim, Y.-J. Lee, Usability-based caching of query results in
+  olap systems., Journal of Systems and Software 68~(2) (2003) 103--119.
+
+\bibitem{baralis}
+E.~Baralis, S.~Paraboschi, E.~Teniente,
+  \href{http://dl.acm.org/citation.cfm?id=645923.671019}{Materialized views
+  selection in a multidimensional database}, in: Proceedings of the 23rd
+  International Conference on Very Large Data Bases, VLDB '97, Morgan Kaufmann
+  Publishers Inc., San Francisco, CA, USA, 1997, pp. 156--165.
+\newline\urlprefix\url{http://dl.acm.org/citation.cfm?id=645923.671019}
+
+\bibitem{kalnis}
+P.~Kalnis, D.~Papadias, Proxy-server architectures for olap., in: S.~Mehrotra,
+  T.~K. Sellis (Eds.), SIGMOD Conference, ACM, 2001, pp. 367--378.
+
+\bibitem{Afrati06}
+F.~N. Afrati, C.~Li, P.~Mitra, Rewriting queries using views in the presence of
+  arithmetic comparisons., Theor. Comput. Sci. 368~(1-2) (2006) 88--123.
+
+\bibitem{Keller96}
+A.~M. Keller, J.~Basu, A predicate-based caching scheme for client-server
+  database architectures, VLDB J. 5~(1) (1996) 35--47.
+
+\bibitem{shim}
+J.~Shim, P.~Scheuermann, R.~Vingralek, Dynamic caching of query results for
+  decision support systems., in: SSDBM, 1999, pp. 254--263.
+
+\end{thebibliography}
 

### 2015-03-08 english_version/IEEE_conf_sib/paper_main.tex
IEEE siberian conference version was added.

--- 2015-02-22 english_version/Elsevier/paper_main.tex
@@
 
-\documentclass[preprint,12pt]{elsarticle}
+\documentclass{article}
 
@@
 \author{Mosin Sergey, Zykin Sergey}
+\title{Truth space method for caching database queries}
+
 \usepackage{amssymb,amsmath}
 \usepackage{amsthm}
-%\usepackage{amsfonts}
-%\usepackage{algorithm}
-%\usepackage{xcolor}
-%\usepackage{textcomp}
+
 \newtheorem{note}{Note}
@@
 
-\journal{Theoretical Computer Science}
-
 \begin{document}
 
-\begin{frontmatter}
+\maketitle
 
-%% Title, authors and addresses
-
-%% use the tnoteref command within \title for footnotes;
-%% use the tnotetext command for theassociated footnote;
-%% use the fnref command within \author or \address for footnotes;
-%% use the fntext command for theassociated footnote;
-%% use the corref command within \author for corresponding author footnotes;
-%% use the cortext command for theassociated footnote;
-%% use the ead command for the email address,
-%% and the form \ead[url] for the home page:
-%% \title{Title\tnoteref{label1}}
-%% \tnotetext[label1]{}
-%% \author{Name\corref{cor1}\fnref{label2}}
-%% \ead{email address}
-%% \ead[url]{home page}
-%% \fntext[label2]{}
-%% \cortext[cor1]{}
-%% \address{Address\fnref{label3}}
-%% \fntext[label3]{}
-
-\title{Truth space method for caching database queries}
-
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{}
-%% \address[label1]{}
-%% \address[label2]{}
-
-\author{S.~V. Mosin, S.~V. Zykin}
-
-\address{Sobolev Institute of Mathematics of the Siberian Branch
-of the Russian Academy of Sciences}
+Adress: Sobolev Institute of Mathematics of the Siberian Branch
+of the Russian Academy of Sciences
 
@@
 background mode.
+
+Keywords: Relational databases, Caching, Truth space
 \end{abstract}
-
-\begin{keyword}
-%% keywords here, in the form: keyword  keyword
-
-%% PACS codes here, in the form: \PACS code \sep code
-
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-Relational databases \sep Caching \sep Truth space
-\end{keyword}
-
-\end{frontmatter}
 
@@
 \end{statement}
-\begin{proof}
-If $X = \<F> $, then $F = F[X]$ and $M(F) = M(F[X])$. Assume $X \subset \<F> $.
-Consider arbitrary point $a \in M(F)$, i.e. $F(a)=\n{TRUE}$. $F$ to $F[X]$
-transformation is made by replacing $T_j$ predicates that contain $A_j \in
-\<F> $ attribute, such that $A_j \notin X$ with $\n{TRUE}$ value. We get
-$F([X])=\n{TRUE}$ according to (\ref{def_F_1}) and (\ref{def_F_2}). Therefore,
-$a \in M(F[X])$.
-\end{proof}
 
@@
 \end{statement}
-\begin{proof}
-It is so indeed, because joining additional relations to $\Re_1$ can only
-remove some tuples that were presented there. After the projection, we will get
-the same relation $\Re_1$ if no tuples were deleted or some subset otherwise. The
-order of joins doesn't matter because of commutativity of this operation.
-\end{proof}
 
@@
 
-\bibliographystyle{elsarticle-num}
-\bibliography{references}
+\begin{thebibliography}{1}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{zyk_pol}
+S.~Zykin, A.~Poluyanov, Multidimensional data building using intermediate
+  representations., Administration problems~(5) (2013) 54--59.
+
+\bibitem{chang}
+C.-S. Park, M.-H. Kim, Y.-J. Lee, Usability-based caching of query results in
+  olap systems., Journal of Systems and Software 68~(2) (2003) 103--119.
+
+\bibitem{baralis}
+E.~Baralis, S.~Paraboschi, E.~Teniente,
+  \href{http://dl.acm.org/citation.cfm?id=645923.671019}{Materialized views
+  selection in a multidimensional database}, in: Proceedings of the 23rd
+  International Conference on Very Large Data Bases, VLDB '97, Morgan Kaufmann
+  Publishers Inc., San Francisco, CA, USA, 1997, pp. 156--165.
+\newline\urlprefix\url{http://dl.acm.org/citation.cfm?id=645923.671019}
+
+\bibitem{kalnis}
+P.~Kalnis, D.~Papadias, Proxy-server architectures for olap., in: S.~Mehrotra,
+  T.~K. Sellis (Eds.), SIGMOD Conference, ACM, 2001, pp. 367--378.
+
+\bibitem{Afrati06}
+F.~N. Afrati, C.~Li, P.~Mitra, Rewriting queries using views in the presence of
+  arithmetic comparisons., Theor. Comput. Sci. 368~(1-2) (2006) 88--123.
+
+\bibitem{Keller96}
+A.~M. Keller, J.~Basu, A predicate-based caching scheme for client-server
+  database architectures, VLDB J. 5~(1) (1996) 35--47.
+
+\bibitem{shim}
+J.~Shim, P.~Scheuermann, R.~Vingralek, Dynamic caching of query results for
+  decision support systems., in: SSDBM, 1999, pp. 254--263.
+
+\end{thebibliography}
 

### 2015-03-09 english_version/MAIS/paper_main.tex
MAIS version was corrected. References file was removed.

@@
 \textbf{Мосин Сергей Владимирович},\\Федеральное государственне бюджетное учреждение науки Институт математики им.
-С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал), аспирант\\
+С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал), аспирант, ORCID 0000-0001-5459-5853\\
 \textbf{Зыкин Сергей Владимирович},\\Федеральное государственне бюджетное учреждение науки Институт математики им.
 С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал), д-р техн.
-наук, профессор
+наук, профессор, ORCID 0000-0002-0576-2149
 }
@@
 \ruabstract{
-В данной статье предлагается новый метод кэширования запросов к реляционной базе данных
-для систем с центральным сервером и распределенными клиентами. Данные загружаются в
-клиентский кэш, основываясь на запросах, выполненных на сервере БД.
-Эти запросы имеют специальный вид, называемый "универсальный реляционный запрос".
-Большинство запросов может быть записано в этом виде.
+В данной статье предлагается новый метод кэширования запросов к реляционной базе данных для систем с центральным сервером и распределенными клиентами.
+Данные загружаются в клиентский кэш, основываясь на запросах, выполненных на сервере БД.
+Каждому запросу ставится в соответствие таблица – результат выполнения запроса.
+Эти запросы имеют специальный вид, называемый "универсальный реляционный запрос", основанный на трех базисных операциях реляционной алгебры: селекции, проекции естественном соединении (natural join).
+Следует отметить, что такая форма запроса наиболее близка к естественному языку и большинство запросов может быть записано в этом виде.
 Кроме того, эта форма записи позволяет анализировать корректность запроса, проверяя свойство соединения без потери информации (СБПИ).
@@
 Для осуществления такой проверки анализируются области истинности логических ограничений искомого запроса и запросов, результаты которых уже содержатся в кэше.
-Этот метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.
+Требуемые операции могут быть проведены аналитически, без необходимости дополнительных запросов к базе банных.
+Предложенный метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.
+Для этого также используются аналитические вычисления, что является принципиальным отличием данной статьи от существующих технологий.
+Для этой цели в статье представлено четыре теоремы.
+В первой и третьей теореме получены условия, позволяющие определить наличие необходимых данных, а во второй и четвертой получены условия вычисления данных только с использованием кэша.
 Проблема актуализации данных не затрагивается в этой статье.
@@
 with a central server and distant clients. Data is loaded into client cache
-based on queries executed on the server. These
-queries have a special form called "universal relational query". The majority of
-search queries to database can be expressed in such form. Besides, this form
+based on queries executed on the server. Every query has corresponding DB table - the result of the query execution. These
+queries have a special form called "universal relational query" based on three fundamental Relational Algebra operations: selection, projection and natural join.
+We have to mention that such a form is the closest one to the natural language and the majority of database
+search queries can be expressed this way. Besides, this form
 allows us to analyze query correctness by checking lossless join property. A
@@
 truth spaces of the logical restrictions in new user's query and results of
-the queries execution in cache. This method may be used to define lacking data
-in cache and execute the query on server for this data only. Problem of data
-actualizations in cache is not discussed in this paper. However, it can be
+the queries execution in cache. Such comparison can be performed analytically , without need in additional Database queries. This method may be used to define lacking data
+in cache and execute the query on server for this data only.
+Analyical approach is also used for this, what distinguishes our paper from existing technologies.
+We propose four theorems for testing the required conditions.
+The first and the third theorems' conditions allow us to define the existence of required data in cache.
+The second and the fourth theorems state conditions to execute queries with cache only.
+Problem of cache data
+actualizations is not discussed in this paper. However, it can be
 solved by cataloging of queries on server and their serving by triggers in
@@
 
+One of the directions of further researches is combining this technology with efficient data manipulations approaches.
+Those can be performed with a help of Graphical Processors.
+GPUs gain serious attention for big data processing \cite{owens, bakkum, govin, yang}.
+Such processor architectures have big potential in multitasking and can reach enormous speeds compared to CPUs, especialy if the operations are performed similarly on different parts of data.
+
 \section{Existing solutions overview and comparison}
@@
 
-In papers \cite{Keller96}, \cite{shim} similar problem is discussed. In our paper we
+Another set of works describes processing scheme for event-driven continuous queries \cite{wata, yates}.
+In the proposed approach, query result caching is introduced to achieve a flexible way to share common operators among queries activated by unpredictable events.
+When a query is activated, an intermediate result generated for the query is stored into the cache area if it is expected to be reused by other queries.
+When other queries including the same operator are activated, they reuse the cached result if the cache includes reusable data.
+
+The work \cite{andrade} proposes methodology for functionally decomposing complex queries in terms of primitives so that multiple reuse sites are exposed to the query optimizer, to increase the amount of reuse.
+
+In \cite{mershad, abbani} queries are separated into subqueries that can yield results from cached nodes. Then query result can be combined from several subqueries.
+
+Using cloud storages one can afford a lot of computational resources for a short time to execute complicated queries effectively on large data with a help of virtual machine clusters \cite{doker, beran, meij, aranda}.
+A set of reliable heuristic algorithms is used.
+Finally the authors carry out a series of experiments that show that their optimizations speed up the federated query evaluation process.
+
+In papers \cite{Keller96}, \cite{shim} similar to our problem is discussed. In this paper we
 make correspondence between cache contents and predicates. Cached data usage
@@
 \renewcommand{\refname}{References}
-\begin{thebibliography}{1}
+\begin{thebibliography}{10}
 \expandafter\ifx\csname url\endcsname\relax
@@
   representations., Administration problems~(5) (2013) 54--59.
+
+\bibitem{owens}
+J.~D. Owens, D.~Luebke, N.~Govindaraju, M.~Harris, J.~Kr??ger, A.~Lefohn, T.~J.
+  Purcell,
+  \href{http://www.blackwell-synergy.com/doi/pdf/10.1111/j.1467-8659.2007.01012.x}{A
+  survey of general-purpose computation on graphics hardware}, Computer
+  Graphics Forum 26~(1) (2007) 80--113.
+\newline\urlprefix\url{http://www.blackwell-synergy.com/doi/pdf/10.1111/j.1467-8659.2007.01012.x}
+
+\bibitem{bakkum}
+P.~Bakkum, K.~Skadron,
+  \href{http://dblp.uni-trier.de/db/conf/asplos/gpgpu2010.html}{Accelerating
+  sql database operations on a gpu with cuda.}, in: D.~R. Kaeli, M.~Leeser
+  (Eds.), GPGPU, Vol. 425 of ACM International Conference Proceeding Series,
+  ACM, 2010, pp. 94--103.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/conf/asplos/gpgpu2010.html}
+
+\bibitem{govin}
+N.~K. Govindaraju, B.~Lloyd, W.~W. 0010, M.~C. Lin, D.~Manocha,
+  \href{http://dblp.uni-trier.de/db/conf/sigmod/sigmod2004.html}{Fast
+  computation of database operations using graphics processors.}, in:
+  G.~Weikum, A.~C. K?¶nig, S.~De??loch (Eds.), SIGMOD Conference, ACM, 2004,
+  pp. 215--226.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/conf/sigmod/sigmod2004.html}
+
+\bibitem{yang}
+B.~He, K.~Yang, R.~Fang, M.~Lu, N.~K. Govindaraju, Q.~Luo, P.~V. Sander,
+  \href{http://dblp.uni-trier.de/db/conf/sigmod/sigmod2008.html}{Relational
+  joins on graphics processors.}, in: J.~T.-L. Wang (Ed.), SIGMOD Conference,
+  ACM, 2008, pp. 511--524.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/conf/sigmod/sigmod2008.html}
 
@@
 
+\bibitem{wata}
+Y.~Watanabe, H.~Kitagawa,
+  \href{http://dblp.uni-trier.de/db/journals/is/is35.html\#WatanabeK10}{Query
+  result caching for multiple event-driven continuous queries.}, Inf. Syst.
+  35~(1) (2010) 94--110.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/is/is35.html\#WatanabeK10}
+
+\bibitem{yates}
+D.~J. Yates, E.~M. Nahum, J.~F. Kurose, P.~Shenoy,
+  \href{http://dblp.uni-trier.de/db/conf/percom/percom2008.html\#YatesNKS08}{Data
+  quality and query cost in pervasive sensing systems.}, in: PerCom, IEEE
+  Computer Society, 2008, pp. 195--205.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/conf/percom/percom2008.html\#YatesNKS08}
+
+\bibitem{andrade}
+H.~Andrade, T.~M. Kur?§, A.~Sussman, J.~H. Saltz,
+  \href{http://dblp.uni-trier.de/db/journals/pc/pc33.html\#AndradeKSS07}{Active
+  semantic caching to optimize multidimensional data analysis in parallel and
+  distributed environments.}, Parallel Computing 33~(7-8) (2007) 497--520.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/pc/pc33.html\#AndradeKSS07}
+
+\bibitem{mershad}
+K.~W. Mershad, H.~Artail,
+  \href{http://dblp.uni-trier.de/db/journals/jpdc/jpdc71.html\#MershadA11}{Codisc:
+  Collaborative and distributed semantic caching for maximizing cache
+  effectiveness in wireless networks.}, J. Parallel Distrib. Comput. 71~(3)
+  (2011) 495--511.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/jpdc/jpdc71.html\#MershadA11}
+
+\bibitem{abbani}
+H.~A. Noor~Abbani, Protecting data flow anonymity in mobile ad hoc networks
+  that employ cooperative caching, Ad Hoc Networks 26 (2015) 69--87.
+
+\bibitem{doker}
+A.~C. Tansel~Dokeroglu, Murat Ali~Bayir, Robust heuristic algorithms for
+  exploiting the common tasks of relational cloud database queries, Applied
+  Soft Computing.
+
+\bibitem{beran}
+P.~P. Beran, W.~Mach, E.~Schikuta, R.~Vigne,
+  \href{http://dblp.uni-trier.de/db/journals/procedia/procedia4.html\#BeranMSV11}{A
+  multi-staged blackboard query optimization framework for world-spanning
+  distributed database resources.}, in: M.~Sato, S.~Matsuoka, P.~M.~A. Sloot,
+  G.~D. van Albada, J.~Dongarra (Eds.), ICCS, Vol.~4 of Procedia Computer
+  Science, Elsevier, 2011, pp. 156--165.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/procedia/procedia4.html\#BeranMSV11}
+
+\bibitem{meij}
+E.~Meij, M.~Bron, L.~Hollink, B.~Huurnink, M.~de~Rijke,
+  \href{http://dblp.uni-trier.de/db/journals/ws/ws9.html\#MeijBHHR11}{Mapping
+  queries to the linking open data cloud: A case study using dbpedia.}, J. Web
+  Sem. 9~(4) (2011) 418--433.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/ws/ws9.html\#MeijBHHR11}
+
+\bibitem{aranda}
+C.~B. Aranda, M.~Arenas, ?.~Corcho, A.~Polleres,
+  \href{http://dblp.uni-trier.de/db/journals/ws/ws18.html\#ArandaACP13}{Federating
+  queries in sparql 1.1: Syntax, semantics and evaluation.}, J. Web Sem. 18~(1)
+  (2013) 1--17.
+\newline\urlprefix\url{http://dblp.uni-trier.de/db/journals/ws/ws18.html\#ArandaACP13}
+
 \bibitem{Keller96}

### 2015-04-11 english_version/IEEE_conf_sib/paper_main.tex
Paper version for IEEE conference was modified.

@@
-%% 
-%% Copyright 2007, 2008, 2009 Elsevier Ltd
-%% 
-%% This file is part of the 'Elsarticle Bundle'.
-%% ---------------------------------------------
-%% 
-%% It may be distributed under the conditions of the LaTeX Project Public
-%% License, either version 1.2 of this license or (at your option) any
-%% later version.  The latest version of this license is in
-%%    http://www.latex-project.org/lppl.txt
-%% and version 1.2 or later is part of all distributions of LaTeX
-%% version 1999/12/01 or later.
-%% 
-%% The list of all files belonging to the 'Elsarticle Bundle' is
-%% given in the file `manifest.txt'.
-%%
-
-%% Template article for Elsevier's document class `elsarticle'
-%% with numbered style bibliographic references
-%% SP 2008/03/01
-
 \documentclass{article}
 
-%% Use the option review to obtain double line spacing
-%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}
-
-%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
-%% for a journal layout:
-%% \documentclass[final,1p,times]{elsarticle}
-%% \documentclass[final,1p,times,twocolumn]{elsarticle}
-%% \documentclass[final,3p,times]{elsarticle}
-%% \documentclass[final,3p,times,twocolumn]{elsarticle}
-%% \documentclass[final,5p,times]{elsarticle}
-%% \documentclass[final,5p,times,twocolumn]{elsarticle}
-
-%% For including figures, graphicx.sty has been loaded in
-%% elsarticle.cls. If you prefer to use the old commands
-%% please give \usepackage{epsfig}
-
-%% The amssymb package provides various useful mathematical symbols
 \author{Mosin Sergey, Zykin Sergey}
-\title{Truth space method for caching database queries}
+\title{Using logical formulas for caching uniform RDB queries}
 
@@
 \def \n #1{\mathit{#1}}
-%% The amsthm package provides extended theorem environments
-%% \usepackage{amsthm}
-
-%% The lineno packages adds line numbers. Start line numbering with
-%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
-%% for the whole article with \linenumbers.
-%% \usepackage{lineno}
 
@@
 %% main text
-\section{Introduction}
 In this paper we discuss the use of data views that are
@@
 
-% Head 1
+One of the directions of further researches is combining this technology with efficient data manipulations approaches.
+Those can be performed with a help of Graphical Processors.
+GPUs gain serious attention for big data processing \cite{owens, bakkum, govin, yang}.
+Such processor architectures have big potential in multitasking and can reach enormous speeds compared to CPUs, especialy if the operations are performed similarly on different parts of data.
+
 \section{Existing solutions overview and comparison}
@@
 
-\begin{thebibliography}{1}
+\begin{thebibliography}{10}
 \expandafter\ifx\csname url\endcsname\relax

### 2015-04-11 report/report_main.tex
Basic report was created.

\documentclass[xcolor=dvipsnames]{beamer}

\usepackage{amssymb,amsmath}
%\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[russian]{babel}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{algorithmic, algorithm}
\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{ {images/} }

%\usepackage{cmap}
%\usepackage{lmodern}
\newtheorem*{mynote}{Замечание}
\newtheorem{mydef}{Определение}
\newtheorem{mytheorem}{Теорема}
\newtheorem{statement}{Утверждение}
\newtheorem*{consequence}{Следствие}

\def \eval #1#2{\left.#1\right\vert_{#2}}
\def \<#1> {\langle #1 \rangle}
\def \n #1{\mathit{#1}}
\def \bigcupn {\bigcup\limits_{v=1}^{n}}
%Information to be included in the title page:
\title{Использование логических формул для кэширования универсальных запросов к Реляционной Базе Данных.}
\author{Мосин С.В.}
\institute{ИМ СО РАН (ОмФ)}
\date{2015}

\usetheme{Boadilla}
%\usecolortheme{seahorse}
\usefonttheme{serif}
%[onlymath]
%\usefonttheme{structuresmallcapsserif}
%\usepackage{mathptmx}
%\usepackage{mathptmx}
%\usepackage[scaled=0.9]{helvet}
%\usepackage{courier}

\setbeameroption{show notes}

\setbeamertemplate{note page}[plain]
\setbeamertemplate{theorems}[numbered]
\setbeamertemplate{footline}{%
    \raisebox{5pt}{\makebox[\paperwidth]{\hfill\makebox[20pt]{\color{gray}
          \scriptsize\insertframenumber}}}\hspace*{5pt}}

%\setbeamercolor*{frametitle}{bg=white, fg=brown}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Содержание}
\tableofcontents
\end{frame}

\section{Клиент-серверная архитектура}

\begin{frame}
\frametitle{\insertsection}
\begin{center}
  \makebox[\textwidth]{\includegraphics[width=0.6\paperwidth]{client-server}}
\end{center}
\end{frame}
\note{В данной работе рассматривается новый метод кэширования запросов к реляционной базе данных для систем с центральным сервером и распределенными клиентами. Данные загружаются в клиентский кэш, основываясь на запросах, выполненных на сервере БД.}

\begin{frame}
\frametitle{Обзор существующих работ}
\begin{itemize}
 \item<1->S.~Zykin, A.~Poluyanov, Multidimensional data building using intermediate
   representations., Administration problems~(5) (2013) 54--59.
 \item<2-> C.-S. Park, M.-H. Kim, Y.-J. Lee, Usability-based caching of query results in
   olap systems., Journal of Systems and Software 68~(2) (2003) 103--119.
 \item<2-> E.~Baralis, S.~Paraboschi, E.~Teniente,
   \href{http://dl.acm.org/citation.cfm?id=645923.671019}{Materialized views
   selection in a multidimensional database}, in: Proceedings of the 23rd
   International Conference on Very Large Data Bases, VLDB '97, Morgan Kaufmann
   Publishers Inc., San Francisco, CA, USA, 1997, pp. 156--165.
 \item<2->  P.~Kalnis, D.~Papadias, Proxy-server architectures for olap., in: S.~Mehrotra,
    T.~K. Sellis (Eds.), SIGMOD Conference, ACM, 2001, pp. 367--378.
\end{itemize}
\note<1>{Данная работа основывается на результатах, полученных в
статье Сергея Владимировича Зыкина и Андрея Полуянова. Устранено условие, ограничивающее атрибуты
промежуточного представления, а также сделано обобщение теорем на случай нескольких
промежуточных представлений.}
\note<2>{Множество публикаций посвящено проблеме управления содержимым кэша.

В первой работе эвристические алгоритмы обновляют кэш, основываясь на данных пользователя
и динамически подстраиваясь под них.

В следующей статье предлагается статический выбор наиболее важных пользовательских запросов
с последующим применением статистических алгоритмов для дополнения кэша

В последней статье исследуется хранение кэшированных данных на различных удаленных серверах.

Наша статья решает проблему наилучшего использования имеющегося кэша для исполнения
пользовательских запросов. Управление кэшем и наилучшее использование кэша - это две
взаимодополняющие задачи.}
\end{frame}
 %--------------------------------------------------------------------------------------- %

\section{Обзор существующих работ}

\begin{frame}
\frametitle{\insertsection}
\begin{itemize} 
 \item<1-> F.~N. Afrati, C.~Li, P.~Mitra, Rewriting queries using views in the presence of
    arithmetic comparisons., Theor. Comput. Sci. 368~(1-2) (2006) 88--123.
 \item<2-> A.~M. Keller, J.~Basu, A predicate-based caching scheme for client-server
    database architectures, VLDB J. 5~(1) (1996) 35--47.
 \item<3-> J.~Shim, P.~Scheuermann, R.~Vingralek, Dynamic caching of query results for
    decision support systems., in: SSDBM, 1999, pp. 254--263.
\end{itemize}
\note<1>{Наиболее близка к рассматриваемой проблеме работа Фото Афрати. В ней рассматриваются
конъюнктивные запросы над доменами данных с предикатами в виде арифметических сравнений и
представлены алгоритмы вычисления запросов с использованием существующих представлений.}
\note<2>{В следующей работе решается задача анализа содержимого кеша, при этом промежуточным
представлениям в кэше ставятся в соответствие предикаты. Проблема использования кэша
решается за счет выводимости предикатов. В данной работе эту проблему предлагается решать за счет
вычисления областей истинности формул соответствующих пользовательских запросов, результаты которых
находятся в кеше.}
\note<3>{Завершает список работ со схожей тематикой статья, в которой предпринята попытка решить
как задачу использования, так и управления кэшем. Кэш используется как при полном совпадении
запросов, так и в определенных условиях в случае если результаты кэша объемлют требуемый запрос.}
\end{frame}
 %---------------------------------------------------------------------------------------%

\section{Пример использования кэша}

\begin{frame}
\frametitle{\insertsection}
Рассмотрим фрагмент схемы БД, представляющий учебный план в универитете:
\vspace{\baselineskip}

$R_1 =\mbox{\it Студенты}\ (\mbox{\bf № студента}, \mbox{ФИО студента},
\mbox{Группа})$\\
$R_2 =\mbox{\it Расписание}\ (\mbox{\bf Группа}, \mbox{\bf № аудитории},
\mbox{Дисциплина})$\\
$R_3 =\mbox{\it Успеваемость}\ (\mbox{\bf № студента}, \mbox{\bf Дисциплина},
\mbox{Оценка})$
\note{На слайде фрагмент БД, где представлены 3 отношения. Имена отношений выделены курсивом, их
первичные ключи - жирным шрифтом.
Номер студента уникально определяет его имя и группу, у данной группы в конкретной аудитории
проходят занятия по данной конкретной дисциплине, а номер студента и дисциплина соответствуют одной
единственной оценке.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\underline{Запрос 1}: Список студентов, изучающих физику, чей номер больше 210:
$$P_1 = \pi_{X_1}(\sigma_{F_1} (R_1 \Join R_2)),$$
где $\pi_{X_1}$ -- проекция по множеству атрибутов $X_1$
$X_1 = \{\text{№ студента},
\text{ФИО студента}, \text{Группа}, \text{Дисциплина}\}$,
$\sigma$ -- операция селекции,
$F_1$ -- логическая формула: $F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} =
\text{Физика})$, $\Join$ --операция естественного соедения.
\note{В результате данного запроса будут получены те кортежи БД, которые удовлетворяют формуле
$F_1$, причем с сервера будут запрошены только атрибуты $\text{№ студента},
\text{ФИО студента}, \text{Группа}, \text{Дисциплина}$}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\underline{Запрос 2}: Ведомости группы M10:
$$P_2 = \pi_{X_2}(\sigma_{F_2} (R_1 \Join R_3)),$$
где
$X_2 = \{\text{№ студента}, \text{Группа}, \text{Дисциплина}, \text{Оценка}\}$, $F_2$
-- логическая формула:\\
$F_2 = (\text{Группа}=\text{M10})$.
\note{Аналогично осуществляется Запрос 2, представляющий набор ведомостей по всем предметам для
студентов группы M10.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R_1 \Join R_2 \Join R_3 )),$$
где $X^{\ast}= \{\text{№ студента}, \text{Группа}, \text{Оценка}\}$, $F^{\ast}$
-- логическая формула: $F^{\ast} = (\textcolor<2>{blue}{\text{№ студента} > 300}\ \&\ 
\textcolor<3>{PineGreen}{\text{Группа} = \text{M10}}\ \&\ \textcolor<2>{brown}{\text{Дисциплина} = \text{Физика}})$.

\only<2>{$F_1 = (\textcolor{blue}{\text{№ студента} > 210}\ \&\ \textcolor{brown}{\text{Дисциплина} =
\text{Физика}})$}
\only<3>{$F_2 = (\textcolor{PineGreen}{\text{Группа}=\text{M10}})$.}

\vspace{\baselineskip}

Запрос может быть выполнен с использованием кэша:
$P^{\ast} =\pi_{X^{\ast}}(\sigma_{F_3} (P_1 \Join P_2 ))$, где $F_3$
-- логическая формула: $F_3 = (\text{Stud\_ID} > 300)$. 
\note<1>{Предположим теперь, что, пользователь хочет получить информацию, формализованную следующим
запросом (запрос $P^{\ast}$ на слайде)}
\note<2-3>{Сравнивая области истинности логических формул, используемых в $P_1$, $P_2$ и $P^{\ast}$,
получаем, что запрос может быть выполнен с помощью кэша.

Запрос к серверу в этом случае не требуется.}
\end{frame}
 %--------------------------------------------------------------------------------------- %

\section{Логические ограничения}

\begin{frame}
\frametitle{\insertsection}
Общий вид логической формулы:
\begin{equation}
F = K_1 \vee K_2 \vee \dots \vee K_m ,
\label{def_F_1}
\end{equation}
\begin{equation}
K_i = T_1 \&\ T_2 \dots \&\ T_n, i = 1, \dots, m ,
\label{def_F_2}
\end{equation}
здесь $T_j, j = 1, \dots, n$ - предикаты на множестве атрибутов БД

\begin{mynote}
Будем рассматривать расширенные имена атрибутов БД, где атрибут
$A_j$ в отношении $R_i$ носит название $R_i.A_j$.
\end{mynote}
\note{
Чтобы упростить вычисления областей истинности, будем рассматривать логические формулы в
Дизъюнктивной Нормальной Форме (ДНФ)

Замечание: Таким образом, можно избежать коллизий имен одинаковых атрибутов из разных отношений}

\begin{mydef}
Множество атрибутов, содержащихся в формуле, определяет её размерность и обозначается $\<F> $.

\begin{equation}
\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\} .
\label{def_F_3}
\end{equation}
\end{mydef}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{itemize}
  \item<1-> операция сравнения $ \n{Expr}_1\ \theta\ \n{Expr}_2\ $, $\theta$ – операция
  сравнения $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$ –
  согласованные по типам допустимые выражения, определенные на множестве
  расширенных имен атрибутов и констант;
  \item<2-> операция $\n{Expr}_1\ \n{[NOT]}\ \n{BETWEEN}\ \n{Expr}_2\ \n{AND}\
  \n{Expr}_3$ (содержимое в прямоугольных скобках $[*]$ для предиката не
  является обязательным при написании);
  \item<3-> операция $\n{Expr}\ \n{[NOT]}\ \n{IN}\ S$, где $S$ – список значений либо
  подзапрос, результатом которого является столбец атрибута $A_j$ в отношении
  $R_i$;
  \item<4-> операция $\n{Str}_1\ \n{[NOT]}\ \n{LIKE}\ \n{Str}_2$, где $\n{Str}_i$ –
  строки;
  \item<5-> операция $\n{Expr}\ \theta\ \n{ALL/ANY}\ S$.
\end{itemize}
\note<1>{Обычное сравнение двух операторов, как пример, Группа = M10.}
\note<2>{Оператор (не) содержится в указанных границах.}
\note<3>{Оператор (не) содержится в списке значений. Отличие от предыдущей операции в том, что
здесь указывается полный перечень возможных значений атрибута, тогда как в $\text{BETWEEN}$ лишь
крайние значения.}
\note<4>{Оператор (не) подходит под указанный паттерн. Это специальное строковое сравнение с
использованием подобия регулярных выражений.}
\note<5>{Операция сравнения $\theta$ выполнена со всеми (ALL), хотя бы с одним из значений из списка (подзапроса) $S$

Перечисленные варианты операций используют не все возможности языка SQL. Например, предикат
$\n{EXISTS}$ не используется, поскольку в нем явно не специфицированы расширенные имена атрибутов,
предикат $\n{NULL}$ используется в данной работе для другой цели.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
$\forall R_i.A_j \in \<F> $:\\
<<Использование неопределенного значения>>: <<Да>> или <<Нет>>.\\
<<Нет>>: $\forall t$ - кортеж БД $t[R_i.A_j] = NULL \rightarrow t$ исключается из рассмотрения

\note{При вычислении логического выражения может быть получено значение UNKNOWN, если на текущем
кортеже t атрибут принимает значение NULL, поскольку результаты вычисления логических выражений в
SQL-запросах соответствуют трехзначной логике. Это приводит к неоднозначной интерпретации результата
не только обычными пользователями, но и опытными программистами.

Для решения этой проблемы предлагается ограничение: каждому атрибуту, входящему в $F^{\ast}$, явно
присваивается признак <<Использование неопределенного значения>> с двумя взаимоисключающими
значениями <<Да>> или <<Нет>>. Семантика этого признака такова, что если ему присвоено значение
<<Да>>, то появление значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$ не служит
основанием удаления этого кортежа из дальнейшего рассмотрения. В противном случае значение признака
<<Нет>> гарантирует, что появление значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$
приведет к удалению этого кортежа из дальнейшего рассмотрения.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
Запишем формулу (\ref{def_F_1}) для $F$ в следующем виде:

\vspace{\baselineskip}

$ F ( \dots, T_j , \dots )$,
где $T_j$ - предикаты формулы (\ref{def_F_2}).

\vspace{\baselineskip}

После преобразования:

\vspace{\baselineskip}

$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, где\\
\begin{itemize}
\item $\wedge_{i,j}(R_i .A_j \neq \n{NULL})$ - конъюнкция по всем атрибутам формулы $F$, для которых
не допустимо значение $\n{NULL}$.
\item Операция $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, где  $\vee_{i,j}(R_i .A_j =
\n{NULL})$ - дизъюнкция по всем атрибутам предиката $T_j$, для которых допустимо значение
$\n{NULL}$.
\end{itemize}
\note{После проведения предложенного преобразования формула примет следующий вид.

Внешние скобки для предиката $T'_j$ определяют приоритет выполнения операций.

Несложно убедиться, что в рамках трехзначной логики преобразованная формула принимает только
значения $\n{TRUE}$ и $\n{FALSE}$. Кроме того, несложно убедиться, что в рамках двузначной логики,
когда в кортежах отсутствуют неопределенные значения, исходная формула $F$ будет эквивалентна
преобразованной формуле, поэтому семантика представления $P$ практически не искажается.

Здесь и далее будем считать все формулы F преобразованными}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
$\mathcal{A} =$ $\{(a_1, \dots, a_n) \mid a_i \in Dom(A_i), i=1,\dots,n\}$, где $Dom(A_i)$ -
множество всех допустимых значений атрибута $A_i$.

\vspace{\baselineskip}

$\mathcal{A} = Dom(A_1)\times Dom(A_2)\times \dots \times Dom(A_n)$ - $n$-мерное пространство
значений всех атрибутов базы данных. Текущее состояние базы данных соответствует подмножеству этого
пространства.

\begin{block}{Пример}
\textcolor<2>{blue}{$\mathcal{A} = Dom(\text{№ студента})\times Dom(\text{ФИО студента}) \times
Dom(\text{Группа}) \times Dom(\text{№ аудитории}) \times Dom(\text{Дисциплина}) \times
Dom(\text{Оценка}) = $}
\textcolor<3->{blue}{$ \{205,315,461\} \times \{\text{Иван Иванов, Петр Петров, Василий Сидоров}\}
\times \{\text{M10, M11}\} \times \{100, 101, 102, 103\} \times \{\text{Физика, Химия}\} \times
\{3, 4, 5\} $}
\end{block}
\note<1>{Введем в рассмотрение множество $\mathcal{A}$.

Данное множество суть декартово произведение областей определения всех атрибутов, определенных в БД.

Допустимые состояния БД образуют лишь некоторую область в этом пространстве, соответствующую
ограничениям целостности на данные.}
\note<2>{Синим выделено декартово произведение всех атрибутов в БД из рассмотренного примера.}
\note<3>{Теперь выделение показывет возможные области определения атрибутов. Все их комбинации и
образуют множество $\mathcal{A}$.

Очевидно, что данное множество содержит больше данных, чем может быть состояний в БД. Например, в данном множестве содержатся точки, сопоставляющие
каждому из студентов каждый из номеров, тогда как у каждого студента такой номер уникален по ограничению
первичного ключа.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{mynote}
Размерность формулы $F$ может быть меньше
размерности $\mathcal{A}$, при этом атрибуты, не входящие в формулу, могут
принимать любые значения.
\end{mynote}

\begin{block}{Пример}
$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}) \rightarrow \<F> = 2$.

\vspace{\baselineskip}

Эквивалентное представление:
\begin{multline*}
F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}\ \&\ \text{ФИО студента}\\
\textit{IN}\
\{\text{Иван Иванов, Петр Петров, Василий Сидоров}\}\ \&\ \text{Группа}\ \textit{IN}\\
\{\text{M10, M11}\} \&\ \text{№ аудитории}\ \textit{IN}\ \{100, 101, 102, 103\}\ \&\ \text{Оценка}\ \textit{IN}\\
\{70, 80, 90\})
\end{multline*}
\end{block}
\note{Если размерность формулы меньше размерности множества $\mathcal{A}$, то нами будет
рассматриваться эквивалентная формула $F$, у кототрой все отсутствующие атрибуты принимают любые
значения из своих доменов.

Рассмотрим преобразование одной из формул примера.

\textbf{\{пример\}}

Данное условие было введено для того, чтобы позволить сравнивать формулы различных размерностей}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{mydef}
Областью истинности формулы $F$, опредененной (\ref{def_F_1}),
(\ref{def_F_2}), (\ref{def_F_3}), называется множество $M (F) = \{a \in \mathcal{A} \mid
F(a) = \n {TRUE}\}$.
\end{mydef}

\begin{block}{Пример}
$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика})$

\vspace{\baselineskip}

$M (F_1) = $ \{(315, Иван Иванов, M10, 100, Физика, 4), (315, Петр Петров, M10, 100, Физика, 4),
\dots (461, Василий Сидоров, M10, 100, Физика, 5)\}
\end{block}
\note{На данном слайде приведено понятие, на котором во многом основывается данная работа. Это
определение области истинности логической формулы.

На практике область истинности определяет все кортежи БД, которые удовлетворяют данной формуле, а
следовательно будут возвращены с сервера в ответ на пользовательский запрос. Однако обращаю внимание,
что в определении не используется конкретная реализация БД, а это значит, что области истинности
можно сравнивать аналитически, не осуществляя каких-либо запросов к серверу БД.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}

\begin{mydef}
Проекцией логической формулы $F$, опредененной (\ref{def_F_1}), (\ref{def_F_2}),
(\ref{def_F_3}) на множество атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в
которой все предикаты $R_i^F.A_i^F \notin X$ заменены тривиальным предикатом $\n{TRUE}$.
\label{projection}
\end{mydef}

\begin{block}{Пример}
$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика})$

\vspace{\baselineskip}

$F_1[\text{№ студента}] = (\text{№ студента} > 210\ \&\ \text{TRUE}) = (\text{№ студента} >
210)$
\end{block}

\begin{statement}[Свойство включения]
$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
\label{proj_property_of_inclusion}
\end{statement}
\note{Здесь приведено понятие проекции логической формулы, являющееся вспомогательным и необходимым
для формулировки дальнейших результатов.

\textbf{\{пример\}}

Проекция формулы обладает свойством включения, которое гласит, что проекция области истинности
формулы всегда содержит область истинности исходной формулы. Это значит, что можно проецировать
формулу, не боясь, что кортежи, которые удовлетворяли формуле F, больше не будут удовлетворять
спроецированной формуле.}
\end{frame}
 %---------------------------------------------------------------------------------------%

\section{Промежуточные представления данных}

\begin{frame}
\frametitle{\insertsection}
$P$=$\{ P_1$, $P_2$, $\dots$, $P_m \}$ - сохраненные промежуточные представления данных

\vspace{\baselineskip}

$$P_v = \pi_{X_v}(\sigma_{F_v} (R^v_1 \Join R^v_2 \Join \dots \Join R^v_{s(v)} ))$$\\
$s(v)$ – количество отношений БД, использованных при формировании $P_v$,\\
$\pi_{X_v}$ - операция проекции по множеству атрибутов $X_v$,\\
$\sigma_{F_v}$ - операция селекции с логическим ограничением на кортежи $F_v$.

\vspace{\baselineskip}

Целевое представление данных:
$$P^{\ast} = \pi_{X^{\ast}}(\sigma_{F^{\ast}} (R^{\ast}_1 \Join R^{\ast}_2\Join \dots \Join
R^{\ast}_l )$$

\note{Рассмотрим формализацию задачи. На входе имеем множество $P$ сохраненных в кеше запросов
пользователя.

Его элементы $P_v$ представлены в виде так называемых «универсальных запросов».
$(R^v_1, R^v_2, \dots, R^v_l )$ - некоторые отношения БД.
По сути это композиция операций естественного соединения, вырезки по строкам и по столбцам.
Большинство запросов на чтение к БД имеют в своей основе такой универсальный запрос.

Целевое представление данных, т.е. запрос пользователя, имеет такую же форму.
$(R^{\ast}_1, R^{\ast}_2, \dots, R^{\ast}_l )$ - некоторые вообще говоря другие отношения БД.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{mytheorem}
$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$, если:
\\а) $X^{\ast} \subseteq X_{v}$
\\б) $ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} \subseteq \{R^{\ast}_{1}, \ldots,
R^{\ast}_{l}\} $
\\в) $M (F^{\ast}) \subseteq M (F_{v}) $.
\label{th_base}
\end{mytheorem}

\note{Предложенные в теореме условия гарантируют, что данные, необходимые для формирования представления $P^{\ast}$, содержатся в промежуточном представлении $P_v$. Однако в нем могут быть лишние кортежи, которые дают значение TRUE при подстановке в формулу $F^{\ast}$. Дело в том, что эти кортежи будут удалены при выполнении операции естественного соединения с отношениями, которых не хватает в множестве отношений $P_v$ для совпадения с множеством отношений $P^{\ast}$}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{mytheorem}
$P^{\ast} = \pi_{X^{\ast}} (\sigma_{F^{\ast}} (P_{v}))$, если:
\\а) $X^{\ast} \subseteq X_{v}$
\\б) $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R^{\ast}_{1}, \ldots,
R^{\ast}_{l}\}$
\\в) $M (F^{\ast}) \subseteq M (F_{v}) $
\\г) $ \<F^{\ast}> \subseteq X_{v} $.
\label{th_base_eq}
\end{mytheorem}

\note{Следующая теорема соответствует частному случаю, где проблема лишних кортежей не возникает.
Здесь добавляется условие на атрибуты формулы F, чтобы сделать возможным вырезку по строкам с
использованием самой формулы F, а не её проекции.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
% Убрал утвеждение, т.к. оно используется только в доказательстве и не несет информации, относящейся
% к теме данной статьи
%\begin{statement}
%Пусть $\Re_1 = R_1 \Join \dots \Join R_k$ - результат естественного
%соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \Join \dots
%\Join R_k \Join R_{k+1} \Join \dots \Join R_{n}$. 
%Тогда $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
%\label{join_property}
%\end{statement}
\begin{mytheorem}
$P^{\ast} \subseteq \pi_{X^{\ast}} ( \sigma_{F^{\ast}[X]} (P_1 \Join \dots \Join
P_n))$, where $X = \bigcupn X_{v}$ if:
\\a) $X^{\ast} \subseteq X$
\\b)
$ \bigcupn \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R'_{1}, \ldots, R'_{s'}\}
\subseteq
\{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
\\c) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $.
\label{th_mult}
\end{mytheorem}
\note{Третья теорема является обобщением первой на случай нескольких промежуточных представлений.
Все условия претерпели изменения в соответствии с этим. Проекция ведется по атрибутам всех промежуточных представлений. Требуется включение всех отношений, участвующих в формировании промежуточных представлений во можество отошений искомого представления. Область истинности формулы $F^{\ast}$ должна содержаться во всех областях истинности имеющихся формул.}
\end{frame}

\begin{frame}
\frametitle{\insertsection}
\begin{mytheorem}
$P^{\ast} =  \pi_{X^{\ast}} ( \sigma_{F^{\ast}} (P_1 \Join \dots \Join
P_n))$, where $X = \bigcupn X_{v}$ if:
\\a) $X^{\ast} \subseteq X$, $X_v \supseteq \<\Join_{i=1}^{s(v)} R^v_i> \cap (\bigcup\limits_{\substack{w=1\\ w \neq v}}^{n} \<\Join_{i=1}^{s(w)} R^w_i> ), v = 1,\dots,n$
\\b)
$ \bigcupn \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R'_{1}, \ldots, R'_{s'}\}
= \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
\\c) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $
\\d) $ \<F^{\ast}> \subseteq X $.
\label{th_mult_eq}
\end{mytheorem}

\note{Снова частный случай, гарантирующий полное совпадение данных пользовательского запроса с кэшем.

Первое условие сильно осложнилось для удовлетворения свойству операции проекции. Фактически здесь требуется, чтобы каждое множество атрибутов промежуточных представлений содержало в себе все атрибуты, по которым пересекаются отношения из данного промежуточного представления и всех остальных. Если этого не потребовать, то могут появиться лишние данные при выполнении операций естественного соединения.

Остальные условия суть комбинация условий теорем 2 и 3.

Данная теорема является в некотором смысле ключевой, так как именно она позволяет на практике воспользоваться содержимым кэша. Она применяется после проверки условий теорем 1 и 3.}
\end{frame}
\note{
\textbf{Заключение к секции \insertsection}.\\
Итак, были рассмотрены четыре теоремы. В первой и третьей теореме получены условия, позволяющие определить наличие необходимых данных, а во второй и четвертой получены условия вычисления данных только с использованием кэша.}
 %---------------------------------------------------------------------------------------%

\section{Выводы}

\begin{frame}
\frametitle{\insertsection}
\begin{center}
\textbf{Что было сделано?}
\end{center}
\begin{itemize}
\item<2-> Подготовлена теоретическая база для разработки системы, взаимодействующей с БД
\end{itemize}

\onslide<3->
{
  \begin{center}
  \textbf{В чем отличие от существующих подходов?}
  \end{center}
}
\begin{itemize}
\item<4-> Использование семантики данных
\item<5-> \textbf<7->{Аналитическая} проверка возможности использования кэша
\item<6-> \textbf<7->{Аналитическое} определение недостающих данных
\onslide<7->\underline{нет необходимости делать запросы на этом этапе}
\end{itemize}
\note<4>{
Зарезервированные данные активно используется в системах управления базами данных (СУБД). Но в большинстве случаев это касается повторного использования данных, записанных в кеш, без предварительного анализа содержимого на предмет возможности частичного или комбинированного использования. Работа СУБД ограничивается тем, что при выполнении очередного запроса блоки данных не запрашиваются с внешних устройств, если они есть в кеше, т. е. анализируются номера блоков, а не их содержимое. }
\note<5>{Для осуществления такой проверки анализируются области истинности логических ограничений искомого
запроса и запросов, результаты которых уже содержатся в кэше. Требуемые операции могут быть
проведены аналитически, без необходимости дополнительных запросов к базе банных.}
\note<6>{Предложенный метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные. Для этого также используются аналитические вычисления, что является принципиальным отличием данной статьи от существующих технологий.}

\onslide<8->
{
  \begin{center}
  \textbf{Что дальше?}
  \end{center}
}
\begin{itemize}
\item<9-> Актуализация данных
\item<10-> Динамическое формирование многомерных данных
\end{itemize}
\note<9>{Проблема актуализации данных не затрагивается в этой статье. Однако она может быть решена
путем учета запросов на сервере и обновлении данных при помощи триггеров.}
\note<10>{Предложенная технология будет использована при динамическом построении многомерных данных.
Промежуточные представления имеют ту же структуру данных, что и таблицы соединений, используемые для
построения гиперкубов. Сохраненные представления данных могут храниться на компьютере пользователя-аналитика и существенно сократить время на формирование данных, необходимых для принятия решений.}
\end{frame}
\end{document}

### 2015-04-24 report/konf_pmfi/report_main.tex
Report for Omsk Technical University was added. Additional encoding fixes for references and old report were introduced.

--- 2015-04-11 report/report_main.tex
@@
 \usepackage{amssymb,amsmath}
-%\usepackage[T2A,T1]{fontenc}
 \usepackage[utf8]{inputenc}
@@
 
-%\usepackage{cmap}
-%\usepackage{lmodern}
 \newtheorem*{mynote}{Замечание}
@@
 \usefonttheme{serif}
-%[onlymath]
-%\usefonttheme{structuresmallcapsserif}
-%\usepackage{mathptmx}
-%\usepackage{mathptmx}
-%\usepackage[scaled=0.9]{helvet}
-%\usepackage{courier}
 
-\setbeameroption{show notes}
+\setbeameroption{hide notes}
 
@@
 
-\begin{frame}
-\frametitle{Содержание}
-\tableofcontents
-\end{frame}
+%\begin{frame}
+%\frametitle{Содержание}
+%\tableofcontents
+%\end{frame}
 
@@
 Данные загружаются в клиентский кэш, основываясь на запросах, выполненных на сервере БД.}
-
-\begin{frame}
-\frametitle{Обзор существующих работ}
-\begin{itemize}
- \item<1->S.~Zykin, A.~Poluyanov, Multidimensional data building using intermediate
-   representations., Administration problems~(5) (2013) 54--59.
- \item<2-> C.-S. Park, M.-H. Kim, Y.-J. Lee, Usability-based caching of query results in
-   olap systems., Journal of Systems and Software 68~(2) (2003) 103--119.
- \item<2-> E.~Baralis, S.~Paraboschi, E.~Teniente,
-   \href{http://dl.acm.org/citation.cfm?id=645923.671019}{Materialized views
-   selection in a multidimensional database}, in: Proceedings of the 23rd
-   International Conference on Very Large Data Bases, VLDB '97, Morgan Kaufmann
-   Publishers Inc., San Francisco, CA, USA, 1997, pp. 156--165.
- \item<2->  P.~Kalnis, D.~Papadias, Proxy-server architectures for olap., in: S.~Mehrotra,
-    T.~K. Sellis (Eds.), SIGMOD Conference, ACM, 2001, pp. 367--378.
-\end{itemize}
-\note<1>{Данная работа основывается на результатах, полученных в
-статье Сергея Владимировича Зыкина и Андрея Полуянова. Устранено условие, ограничивающее атрибуты
-промежуточного представления, а также сделано обобщение теорем на случай нескольких
-промежуточных представлений.}
-\note<2>{Множество публикаций посвящено проблеме управления содержимым кэша.
-
-В первой работе эвристические алгоритмы обновляют кэш, основываясь на данных пользователя
-и динамически подстраиваясь под них.
-
-В следующей статье предлагается статический выбор наиболее важных пользовательских запросов
-с последующим применением статистических алгоритмов для дополнения кэша
-
-В последней статье исследуется хранение кэшированных данных на различных удаленных серверах.
-
-Наша статья решает проблему наилучшего использования имеющегося кэша для исполнения
-пользовательских запросов. Управление кэшем и наилучшее использование кэша - это две
-взаимодополняющие задачи.}
-\end{frame}
- %--------------------------------------------------------------------------------------- %
-
-\section{Обзор существующих работ}
-
-\begin{frame}
-\frametitle{\insertsection}
-\begin{itemize} 
- \item<1-> F.~N. Afrati, C.~Li, P.~Mitra, Rewriting queries using views in the presence of
-    arithmetic comparisons., Theor. Comput. Sci. 368~(1-2) (2006) 88--123.
- \item<2-> A.~M. Keller, J.~Basu, A predicate-based caching scheme for client-server
-    database architectures, VLDB J. 5~(1) (1996) 35--47.
- \item<3-> J.~Shim, P.~Scheuermann, R.~Vingralek, Dynamic caching of query results for
-    decision support systems., in: SSDBM, 1999, pp. 254--263.
-\end{itemize}
-\note<1>{Наиболее близка к рассматриваемой проблеме работа Фото Афрати. В ней рассматриваются
-конъюнктивные запросы над доменами данных с предикатами в виде арифметических сравнений и
-представлены алгоритмы вычисления запросов с использованием существующих представлений.}
-\note<2>{В следующей работе решается задача анализа содержимого кеша, при этом промежуточным
-представлениям в кэше ставятся в соответствие предикаты. Проблема использования кэша
-решается за счет выводимости предикатов. В данной работе эту проблему предлагается решать за счет
-вычисления областей истинности формул соответствующих пользовательских запросов, результаты которых
-находятся в кеше.}
-\note<3>{Завершает список работ со схожей тематикой статья, в которой предпринята попытка решить
-как задачу использования, так и управления кэшем. Кэш используется как при полном совпадении
-запросов, так и в определенных условиях в случае если результаты кэша объемлют требуемый запрос.}
-\end{frame}
  %---------------------------------------------------------------------------------------%
@@
 \frametitle{\insertsection}
-Рассмотрим фрагмент схемы БД, представляющий учебный план в универитете:
+Рассмотрим фрагмент схемы БД, представляющий учебный план в университете:
 \vspace{\baselineskip}
@@
 
-Запрос может быть выполнен с использованием кэша:
+\onslide<4>{Запрос может быть выполнен с использованием кэша:
 $P^{\ast} =\pi_{X^{\ast}}(\sigma_{F_3} (P_1 \Join P_2 ))$, где $F_3$
--- логическая формула: $F_3 = (\text{Stud\_ID} > 300)$. 
+-- логическая формула: $F_3 = (\text{Stud\_ID} > 300)$.}
 \note<1>{Предположим теперь, что, пользователь хочет получить информацию, формализованную следующим
-запросом (запрос $P^{\ast}$ на слайде)}
-\note<2-3>{Сравнивая области истинности логических формул, используемых в $P_1$, $P_2$ и $P^{\ast}$,
+запросом (запрос $P^{\ast}$ на слайде). Это список оценок по физике студентов группы М10, чей номер больше 300.}
+\note<4>{Сравнивая области истинности логических формул, используемых в $P_1$, $P_2$ и $P^{\ast}$,
 получаем, что запрос может быть выполнен с помощью кэша.
@@
 \frametitle{\insertsection}
+Возможные значения  $T_j$:
 \begin{itemize}
@@
 
-\begin{block}{Пример}
-\textcolor<2>{blue}{$\mathcal{A} = Dom(\text{№ студента})\times Dom(\text{ФИО студента}) \times
-Dom(\text{Группа}) \times Dom(\text{№ аудитории}) \times Dom(\text{Дисциплина}) \times
-Dom(\text{Оценка}) = $}
-\textcolor<3->{blue}{$ \{205,315,461\} \times \{\text{Иван Иванов, Петр Петров, Василий Сидоров}\}
-\times \{\text{M10, M11}\} \times \{100, 101, 102, 103\} \times \{\text{Физика, Химия}\} \times
-\{3, 4, 5\} $}
-\end{block}
-\note<1>{Введем в рассмотрение множество $\mathcal{A}$.
+\note{Введем в рассмотрение множество $\mathcal{A}$.
 
@@
 ограничениям целостности на данные.}
-\note<2>{Синим выделено декартово произведение всех атрибутов в БД из рассмотренного примера.}
-\note<3>{Теперь выделение показывет возможные области определения атрибутов. Все их комбинации и
-образуют множество $\mathcal{A}$.
-
-Очевидно, что данное множество содержит больше данных, чем может быть состояний в БД. Например, в данном множестве содержатся точки, сопоставляющие
-каждому из студентов каждый из номеров, тогда как у каждого студента такой номер уникален по ограничению
-первичного ключа.}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-\begin{mynote}
-Размерность формулы $F$ может быть меньше
-размерности $\mathcal{A}$, при этом атрибуты, не входящие в формулу, могут
-принимать любые значения.
-\end{mynote}
-
-\begin{block}{Пример}
-$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}) \rightarrow \<F> = 2$.
-
-\vspace{\baselineskip}
-
-Эквивалентное представление:
-\begin{multline*}
-F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}\ \&\ \text{ФИО студента}\\
-\textit{IN}\
-\{\text{Иван Иванов, Петр Петров, Василий Сидоров}\}\ \&\ \text{Группа}\ \textit{IN}\\
-\{\text{M10, M11}\} \&\ \text{№ аудитории}\ \textit{IN}\ \{100, 101, 102, 103\}\ \&\ \text{Оценка}\ \textit{IN}\\
-\{70, 80, 90\})
-\end{multline*}
-\end{block}
-\note{Если размерность формулы меньше размерности множества $\mathcal{A}$, то нами будет
-рассматриваться эквивалентная формула $F$, у кототрой все отсутствующие атрибуты принимают любые
-значения из своих доменов.
-
-Рассмотрим преобразование одной из формул примера.
-
-\textbf{\{пример\}}
-
-Данное условие было введено для того, чтобы позволить сравнивать формулы различных размерностей}
 \end{frame}
@@
 \begin{mydef}
-Областью истинности формулы $F$, опредененной (\ref{def_F_1}),
+Областью истинности формулы $F$, определенной (\ref{def_F_1}),
 (\ref{def_F_2}), (\ref{def_F_3}), называется множество $M (F) = \{a \in \mathcal{A} \mid
@@
 
-\begin{block}{Пример}
-$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика})$
+\onslide<2->
+\begin{mydef}
+Проекцией логической формулы $F$, определенной (\ref{def_F_1}), (\ref{def_F_2}),
+(\ref{def_F_3}) на множество атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в
+которой все предикаты $R_i^F.A_i^F \notin X$ заменены тривиальным предикатом $\n{TRUE}$.
+\label{projection}
+\end{mydef}
 
-\vspace{\baselineskip}
-
-$M (F_1) = $ \{(315, Иван Иванов, M10, 100, Физика, 4), (315, Петр Петров, M10, 100, Физика, 4),
-\dots (461, Василий Сидоров, M10, 100, Физика, 5)\}
-\end{block}
-\note{На данном слайде приведено понятие, на котором во многом основывается данная работа. Это
+\onslide<3->
+\begin{statement}[Свойство включения]
+$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
+\label{proj_property_of_inclusion}
+\end{statement}
+\note<1>{На данном слайде приведено понятие, на котором во многом основывается данная работа. Это
 определение области истинности логической формулы.
@@
 можно сравнивать аналитически, не осуществляя каких-либо запросов к серверу БД.}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-
-\begin{mydef}
-Проекцией логической формулы $F$, опредененной (\ref{def_F_1}), (\ref{def_F_2}),
-(\ref{def_F_3}) на множество атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в
-которой все предикаты $R_i^F.A_i^F \notin X$ заменены тривиальным предикатом $\n{TRUE}$.
-\label{projection}
-\end{mydef}
-
-\begin{block}{Пример}
-$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика})$
-
-\vspace{\baselineskip}
-
-$F_1[\text{№ студента}] = (\text{№ студента} > 210\ \&\ \text{TRUE}) = (\text{№ студента} >
-210)$
-\end{block}
-
-\begin{statement}[Свойство включения]
-$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
-\label{proj_property_of_inclusion}
-\end{statement}
-\note{Здесь приведено понятие проекции логической формулы, являющееся вспомогательным и необходимым
-для формулировки дальнейших результатов.
-
-\textbf{\{пример\}}
-
-Проекция формулы обладает свойством включения, которое гласит, что проекция области истинности
+\note<2>{Здесь приведено понятие проекции логической формулы, являющееся вспомогательным и необходимым
+для формулировки дальнейших результатов.}
+\note<3>{Проекция формулы обладает свойством включения, которое гласит, что проекция области истинности
 формулы всегда содержит область истинности исходной формулы. Это значит, что можно проецировать
@@
 % Убрал утвеждение, т.к. оно используется только в доказательстве и не несет информации, относящейся
-% к теме данной статьи
+% к теме данной работы
 %\begin{statement}
@@
 $P^{\ast} \subseteq \pi_{X^{\ast}} ( \sigma_{F^{\ast}[X]} (P_1 \Join \dots \Join
-P_n))$, where $X = \bigcupn X_{v}$ if:
-\\a) $X^{\ast} \subseteq X$
-\\b)
+P_n))$, где $X = \bigcupn X_{v}$ если:
+\\а) $X^{\ast} \subseteq X$
+\\б)
 $ \bigcupn \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R'_{1}, \ldots, R'_{s'}\}
@@
 \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
-\\c) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $.
+\\в) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $.
 \label{th_mult}
@@
 $P^{\ast} =  \pi_{X^{\ast}} ( \sigma_{F^{\ast}} (P_1 \Join \dots \Join
-P_n))$, where $X = \bigcupn X_{v}$ if:
-\\a) $X^{\ast} \subseteq X$, $X_v \supseteq \<\Join_{i=1}^{s(v)} R^v_i> \cap (\bigcup\limits_{\substack{w=1\\ w \neq v}}^{n} \<\Join_{i=1}^{s(w)} R^w_i> ), v = 1,\dots,n$
-\\b)
+P_n))$, где $X = \bigcupn X_{v}$ если:
+\\а) $X^{\ast} \subseteq X$, $X_v \supseteq \<\Join_{i=1}^{s(v)} R^v_i> \cap (\bigcup\limits_{\substack{w=1\\ w \neq v}}^{n} \<\Join_{i=1}^{s(w)} R^w_i> ), v = 1,\dots,n$
+\\б)
 $ \bigcupn \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R'_{1}, \ldots, R'_{s'}\}
 = \{R^{\ast}_{1}, \ldots, R^{\ast}_{l}\} $
-\\c) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $
-\\d) $ \<F^{\ast}> \subseteq X $.
+\\в) $M(F^{\ast}) \subseteq M(F_{v}), v = 1,\dots,n $
+\\г) $ \<F^{\ast}> \subseteq X $.
 \label{th_mult_eq}
@@
 \end{itemize}
+\note<2>{
+Данный материал может быть использован для разработки ПО, являющегося прослойкой над СУБД и
+определяющего возможности использования кэша.}
 
@@
 \note<6>{Предложенный метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.
-Для этого также используются аналитические вычисления, что является принципиальным отличием данной статьи от существующих технологий.}
+Для этого также используются аналитические вычисления, что является принципиальным отличием данной работы от существующих технологий.}
 
@@
 \end{itemize}
-\note<9>{Проблема актуализации данных не затрагивается в этой статье. Однако она может быть решена
+\note<9>{Проблема актуализации данных не затрагивается в этой работе. Однако она может быть решена
 путем учета запросов на сервере и обновлении данных при помощи триггеров.}

### 2015-04-25 report/laboratory/report_main.tex
lab report moved to another dir

--- 2015-04-11 report/report_main.tex
@@
 
-\setbeameroption{show notes}
+\setbeameroption{hide notes}
 
@@
 
-Наша статья решает проблему наилучшего использования имеющегося кэша для исполнения
+Наша работа решает проблему наилучшего использования имеющегося кэша для исполнения
 пользовательских запросов. Управление кэшем и наилучшее использование кэша - это две
@@
 \note<1>{Наиболее близка к рассматриваемой проблеме работа Фото Афрати. В ней рассматриваются
-конъюнктивные запросы над доменами данных с предикатами в виде арифметических сравнений и
+конъюнктивные запросы с предикатами в виде арифметических сравнений и
 представлены алгоритмы вычисления запросов с использованием существующих представлений.}
@@
 решается за счет выводимости предикатов. В данной работе эту проблему предлагается решать за счет
-вычисления областей истинности формул соответствующих пользовательских запросов, результаты которых
-находятся в кеше.}
+вычисления областей истинности формул пользовательских запросов.}
 \note<3>{Завершает список работ со схожей тематикой статья, в которой предпринята попытка решить
 как задачу использования, так и управления кэшем. Кэш используется как при полном совпадении
-запросов, так и в определенных условиях в случае если результаты кэша объемлют требуемый запрос.}
+запросов, так и в случае если результаты кэша объемлют требуемый запрос.}
 \end{frame}
@@
 \begin{block}{Пример}
-$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}) \rightarrow \<F> = 2$.
+$F_1 = (\text{№ студента} > 210\ \&\ \text{Дисциплина} = \text{Физика}) \rightarrow \dim(F) = 2$.
 
@@
 \begin{mydef}
-Проекцией логической формулы $F$, опредененной (\ref{def_F_1}), (\ref{def_F_2}),
+Проекцией логической формулы $F$, определенной (\ref{def_F_1}), (\ref{def_F_2}),
 (\ref{def_F_3}) на множество атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в

### 2015-05-25 english_version/MAIS/paper_main.tex
MAIS version was modified

@@
 \title{Truth space method for caching database queries}
-\author{Mosin~S.~V., Zykin~S.~V.\footnote{Supported by the Russian Foundation for Basic Research under
-grant \#12-07-00066-a.}}
+\author{Mosin~S.~V., Zykin~S.~V.}
 \authorcr{Mosin~S.~V., Zykin~S.~V.}
@@
 of the Russian Academy of Sciences, \\Pevtsova str., 13, Omsk, 644043, Russia}
-\ruaffiliation{\small Омский филиал Федерального государственного бюджетного учреждения науки Института математики им.
-С.Л.~Соболева Сибирского отделения Российской академии наук \\644043, Россия, г.
-Омск,
-ул. Певцова, 13,}
+\ruaffiliation{\small Федеральное государственное бюджетное учреждение науки Институт математики им.
+С.Л.~Соболева Сибирского отделения Российской академии наук (Омский филиал) \\644043, Россия, г.
+Омск, ул. Певцова, 13,}
 %%%%

### 2015-05-25 report/konf_pmfi/report_main.tex
Report version for PMFI OmGTU conference was added

@@
 \frame{\titlepage}
-
-%\begin{frame}
-%\frametitle{Содержание}
-%\tableofcontents
-%\end{frame}
 
@@
 \end{itemize}
-\note<1>{Обычное сравнение двух операторов, как пример, Группа = M10.}
-\note<2>{Оператор (не) содержится в указанных границах.}
-\note<3>{Оператор (не) содержится в списке значений. Отличие от предыдущей операции в том, что
+\note<5>{Обычное сравнение двух операторов, как пример, Группа = M10.
+Оператор (не) содержится в указанных границах.
+Оператор (не) содержится в списке значений. Отличие от предыдущей операции в том, что
 здесь указывается полный перечень возможных значений атрибута, тогда как в $\text{BETWEEN}$ лишь
-крайние значения.}
-\note<4>{Оператор (не) подходит под указанный паттерн. Это специальное строковое сравнение с
-использованием подобия регулярных выражений.}
-\note<5>{Операция сравнения $\theta$ выполнена со всеми (ALL), хотя бы с одним из значений из списка (подзапроса) $S$
+крайние значения.
+Оператор (не) подходит под указанный паттерн. Это специальное строковое сравнение с
+использованием подобия регулярных выражений.
+Операция сравнения $\theta$ выполнена со всеми (ALL), хотя бы с одним из значений из списка (подзапроса) $S$
 
@@
 предикат $\n{NULL}$ используется в данной работе для другой цели.}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-$\forall R_i.A_j \in \<F> $:\\
-<<Использование неопределенного значения>>: <<Да>> или <<Нет>>.\\
-<<Нет>>: $\forall t$ - кортеж БД $t[R_i.A_j] = NULL \rightarrow t$ исключается из рассмотрения
-
-\note{При вычислении логического выражения может быть получено значение UNKNOWN, если на текущем
-кортеже t атрибут принимает значение NULL, поскольку результаты вычисления логических выражений в
-SQL-запросах соответствуют трехзначной логике. Это приводит к неоднозначной интерпретации результата
-не только обычными пользователями, но и опытными программистами.
-
-Для решения этой проблемы предлагается ограничение: каждому атрибуту, входящему в $F^{\ast}$, явно
-присваивается признак <<Использование неопределенного значения>> с двумя взаимоисключающими
-значениями <<Да>> или <<Нет>>. Семантика этого признака такова, что если ему присвоено значение
-<<Да>>, то появление значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$ не служит
-основанием удаления этого кортежа из дальнейшего рассмотрения. В противном случае значение признака
-<<Нет>> гарантирует, что появление значения $\n{NULL}$ для указанного атрибута в текущем кортеже $t$
-приведет к удалению этого кортежа из дальнейшего рассмотрения.}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-Запишем формулу (\ref{def_F_1}) для $F$ в следующем виде:
-
-\vspace{\baselineskip}
-
-$ F ( \dots, T_j , \dots )$,
-где $T_j$ - предикаты формулы (\ref{def_F_2}).
-
-\vspace{\baselineskip}
-
-После преобразования:
-
-\vspace{\baselineskip}
-
-$F( \dots, T'_j , \dots )\wedge_{i,j}(R_i .A_j \neq  \n{NULL})$, где\\
-\begin{itemize}
-\item $\wedge_{i,j}(R_i .A_j \neq \n{NULL})$ - конъюнкция по всем атрибутам формулы $F$, для которых
-не допустимо значение $\n{NULL}$.
-\item Операция $T'_j = (T_j \vee_{i,j}(R_i .A_j = \n{NULL}))$, где  $\vee_{i,j}(R_i .A_j =
-\n{NULL})$ - дизъюнкция по всем атрибутам предиката $T_j$, для которых допустимо значение
-$\n{NULL}$.
-\end{itemize}
-\note{После проведения предложенного преобразования формула примет следующий вид.
-
-Внешние скобки для предиката $T'_j$ определяют приоритет выполнения операций.
-
-Несложно убедиться, что в рамках трехзначной логики преобразованная формула принимает только
-значения $\n{TRUE}$ и $\n{FALSE}$. Кроме того, несложно убедиться, что в рамках двузначной логики,
-когда в кортежах отсутствуют неопределенные значения, исходная формула $F$ будет эквивалентна
-преобразованной формуле, поэтому семантика представления $P$ практически не искажается.
-
-Здесь и далее будем считать все формулы F преобразованными}
 \end{frame}
@@
 \begin{mytheorem}
-$P^{\ast} \subseteq \pi_{X^{\ast}} (\sigma_{F^{\ast}[X_v]} (P_{v}))$, если:
-\\а) $X^{\ast} \subseteq X_{v}$
-\\б) $ \{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} \subseteq \{R^{\ast}_{1}, \ldots,
-R^{\ast}_{l}\} $
-\\в) $M (F^{\ast}) \subseteq M (F_{v}) $.
-\label{th_base}
-\end{mytheorem}
-
-\note{Предложенные в теореме условия гарантируют, что данные, необходимые для формирования представления $P^{\ast}$, содержатся в промежуточном представлении $P_v$.
-Однако в нем могут быть лишние кортежи, которые дают значение TRUE при подстановке в формулу $F^{\ast}$.
-Дело в том, что эти кортежи будут удалены при выполнении операции естественного соединения с отношениями, которых не хватает в множестве отношений $P_v$ для совпадения с множеством отношений $P^{\ast}$}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-\begin{mytheorem}
-$P^{\ast} = \pi_{X^{\ast}} (\sigma_{F^{\ast}} (P_{v}))$, если:
-\\а) $X^{\ast} \subseteq X_{v}$
-\\б) $\{R^{v}_{1}, \ldots, R^{v}_{s(v)}\} = \{R^{\ast}_{1}, \ldots,
-R^{\ast}_{l}\}$
-\\в) $M (F^{\ast}) \subseteq M (F_{v}) $
-\\г) $ \<F^{\ast}> \subseteq X_{v} $.
-\label{th_base_eq}
-\end{mytheorem}
-
-\note{Следующая теорема соответствует частному случаю, где проблема лишних кортежей не возникает.
-Здесь добавляется условие на атрибуты формулы F, чтобы сделать возможным вырезку по строкам с
-использованием самой формулы F, а не её проекции.}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-% Убрал утвеждение, т.к. оно используется только в доказательстве и не несет информации, относящейся
-% к теме данной работы
-%\begin{statement}
-%Пусть $\Re_1 = R_1 \Join \dots \Join R_k$ - результат естественного
-%соединения некоторых $k$ отношений. Пусть также $\Re_2 = R_1 \Join \dots
-%\Join R_k \Join R_{k+1} \Join \dots \Join R_{n}$. 
-%Тогда $\Re_2 [\langle \Join_{i=1}^{k} R_i \rangle] \subseteq \Re_1$
-%\label{join_property}
-%\end{statement}
-\begin{mytheorem}
 $P^{\ast} \subseteq \pi_{X^{\ast}} ( \sigma_{F^{\ast}[X]} (P_1 \Join \dots \Join
@@
 \end{mytheorem}
-\note{Третья теорема является обобщением первой на случай нескольких промежуточных представлений.
-Все условия претерпели изменения в соответствии с этим.
-Проекция ведется по атрибутам всех промежуточных представлений.
-Требуется включение всех отношений, участвующих в формировании промежуточных представлений во можество отошений искомого представления.
-Область истинности формулы $F^{\ast}$ должна содержаться во всех областях истинности имеющихся формул.}
+\note{Предложенные в теореме условия гарантируют, что данные, необходимые для формирования представления $P^{\ast}$, содержатся в промежуточном представлении $P_v$.
+Однако в нем могут быть лишние кортежи, которые дают значение TRUE при подстановке в формулу $F^{\ast}$.
+Дело в том, что эти кортежи будут удалены при выполнении операции естественного соединения с отношениями, которых не хватает в множестве отношений $P_v$ для совпадения с множеством отношений $P^{\ast}$}
 \end{frame}
@@
 
-\note{Снова частный случай, гарантирующий полное совпадение данных пользовательского запроса с кэшем.
+\note{Следующая теорема соответствует частному случаю, где проблема лишних кортежей не возникает.
+Здесь добавляется условие на атрибуты формулы F, чтобы сделать возможным вырезку по строкам с
+использованием самой формулы F, а не её проекции.
 
-Первое условие сильно осложнилось для удовлетворения свойству операции проекции.
-Фактически здесь требуется, чтобы каждое множество атрибутов промежуточных представлений содержало в себе все атрибуты, по которым пересекаются отношения из данного промежуточного представления и всех остальных.
-Если этого не потребовать, то могут появиться лишние данные при выполнении операций естественного соединения.
+Первое условие сильно осложнилось для удовлетворения свойству операции проекции. Если этого не
+потребовать, то могут появиться лишние данные при выполнении операций естественного соединения.
 
@@
 
-Данная теорема является в некотором смысле ключевой, так как именно она позволяет на практике воспользоваться содержимым кэша. Она применяется после проверки условий теорем 1 и 3.}
+Данная теорема является в некотором смысле ключевой, так как именно она позволяет на практике
+воспользоваться содержимым кэша. Она применяется после проверки условий теорем 1 и 3.}
 \end{frame}
-\note{
-\textbf{Заключение к секции \insertsection}.\\
-Итак, были рассмотрены четыре теоремы.
-В первой и третьей теореме получены условия, позволяющие определить наличие необходимых данных, а во второй и четвертой получены условия вычисления данных только с использованием кэша.}
  %---------------------------------------------------------------------------------------%
@@
 Данный материал может быть использован для разработки ПО, являющегося прослойкой над СУБД и
-определяющего возможности использования кэша.}
+управляющего использованием кэша, снижая тем самым количество передаваемых данных и общее время
+работы системы.}
 
@@
 }
-\note<5>{Для осуществления такой проверки анализируются области истинности логических ограничений искомого
-запроса и запросов, результаты которых уже содержатся в кэше. Требуемые операции могут быть
-проведены аналитически, без необходимости дополнительных запросов к базе банных.}
+\note<5>{Для осуществления проверки использования кэша анализируются области истинности логических
+ограничений искомого запроса и запросов, результаты которых уже содержатся в кэше. Требуемые
+операции могут быть проведены аналитически, без необходимости дополнительных запросов к базе банных.
+}
 \note<6>{Предложенный метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.

### 2015-05-25 report/ieee_sibconf/report_main.tex
IEEE Siberian conference report version was added.

--- 2015-05-25 report/konf_pmfi/report_main.tex
@@
 \end{frame}
-\note{В данной работе рассматривается новый метод кэширования запросов к реляционной базе данных для систем с центральным сервером и распределенными клиентами.
-Данные загружаются в клиентский кэш, основываясь на запросах, выполненных на сервере БД.}
+\note{В данной работе рассматривается новый метод кэширования запросов к реляционной базе данных для
+систем с центральным сервером и распределенными клиентами. Данные загружаются в клиентский кэш,
+основываясь на запросах, выполненных на сервере БД.}
  %---------------------------------------------------------------------------------------%
@@
 здесь $T_j, j = 1, \dots, n$ - предикаты на множестве атрибутов БД
-
-\begin{mynote}
-Будем рассматривать расширенные имена атрибутов БД, где атрибут
-$A_j$ в отношении $R_i$ носит название $R_i.A_j$.
-\end{mynote}
-\note{
-Чтобы упростить вычисления областей истинности, будем рассматривать логические формулы в
-Дизъюнктивной Нормальной Форме (ДНФ)
-
-Замечание: Таким образом, можно избежать коллизий имен одинаковых атрибутов из разных отношений}
-
-\begin{mydef}
-Множество атрибутов, содержащихся в формуле, определяет её размерность и обозначается $\<F> $.
-
-\begin{equation}
-\<F> = \{R_1^F.A_1^F, \dots, R_k^F.A_k^F\} .
-\label{def_F_3}
-\end{equation}
-\end{mydef}
-\end{frame}
-
-\begin{frame}
-\frametitle{\insertsection}
-Возможные значения  $T_j$:
-\begin{itemize}
-  \item<1-> операция сравнения $ \n{Expr}_1\ \theta\ \n{Expr}_2\ $, $\theta$ – операция
-  сравнения $(\theta \in \{=, \neq, >, <, \leq, \geq\})$, $\n{Expr}_i$ –
-  согласованные по типам допустимые выражения, определенные на множестве
-  расширенных имен атрибутов и констант;
-  \item<2-> операция $\n{Expr}_1\ \n{[NOT]}\ \n{BETWEEN}\ \n{Expr}_2\ \n{AND}\
-  \n{Expr}_3$ (содержимое в прямоугольных скобках $[*]$ для предиката не
-  является обязательным при написании);
-  \item<3-> операция $\n{Expr}\ \n{[NOT]}\ \n{IN}\ S$, где $S$ – список значений либо
-  подзапрос, результатом которого является столбец атрибута $A_j$ в отношении
-  $R_i$;
-  \item<4-> операция $\n{Str}_1\ \n{[NOT]}\ \n{LIKE}\ \n{Str}_2$, где $\n{Str}_i$ –
-  строки;
-  \item<5-> операция $\n{Expr}\ \theta\ \n{ALL/ANY}\ S$.
-\end{itemize}
-\note<5>{Обычное сравнение двух операторов, как пример, Группа = M10.
-Оператор (не) содержится в указанных границах.
-Оператор (не) содержится в списке значений. Отличие от предыдущей операции в том, что
-здесь указывается полный перечень возможных значений атрибута, тогда как в $\text{BETWEEN}$ лишь
-крайние значения.
-Оператор (не) подходит под указанный паттерн. Это специальное строковое сравнение с
-использованием подобия регулярных выражений.
-Операция сравнения $\theta$ выполнена со всеми (ALL), хотя бы с одним из значений из списка (подзапроса) $S$
-
-Перечисленные варианты операций используют не все возможности языка SQL. Например, предикат
-$\n{EXISTS}$ не используется, поскольку в нем явно не специфицированы расширенные имена атрибутов,
-предикат $\n{NULL}$ используется в данной работе для другой цели.}
 \end{frame}
@@
 Областью истинности формулы $F$, определенной (\ref{def_F_1}),
-(\ref{def_F_2}), (\ref{def_F_3}), называется множество $M (F) = \{a \in \mathcal{A} \mid
+(\ref{def_F_2}), называется множество $M (F) = \{a \in \mathcal{A} \mid
 F(a) = \n {TRUE}\}$.
@@
 \begin{mydef}
-Проекцией логической формулы $F$, определенной (\ref{def_F_1}), (\ref{def_F_2}),
-(\ref{def_F_3}) на множество атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в
+Проекцией логической формулы $F$, определенной (\ref{def_F_1}), (\ref{def_F_2}) на множество
+атрибутов $X$ называется логическая формула $F[X], \<F[X]> = X$, в
 которой все предикаты $R_i^F.A_i^F \notin X$ заменены тривиальным предикатом $\n{TRUE}$.
@@
 
-\onslide<3->
-\begin{statement}[Свойство включения]
-$\forall X \subseteq \<F> \quad M(F) \subseteq M(F[X])$
-\label{proj_property_of_inclusion}
-\end{statement}
 \note<1>{На данном слайде приведено понятие, на котором во многом основывается данная работа. Это
@@
 для формулировки дальнейших результатов.}
-\note<3>{Проекция формулы обладает свойством включения, которое гласит, что проекция области истинности
-формулы всегда содержит область истинности исходной формулы. Это значит, что можно проецировать
-формулу, не боясь, что кортежи, которые удовлетворяли формуле F, больше не будут удовлетворять
-спроецированной формуле.}
 \end{frame}
@@
 \item<4-> Использование семантики данных
-\item<5-> \textbf<7->{Аналитическая} проверка возможности использования кэша
-\item<6-> \textbf<7->{Аналитическое} определение недостающих данных
-\onslide<7->\underline{нет необходимости делать запросы на этом этапе}
+\item<5-> \textbf{Аналитическая} проверка возможности использования кэша и определения
+  недостающих данных
 \end{itemize}
 \note<4>{
-Зарезервированные данные активно используется в системах управления базами данных (СУБД).
-Но в большинстве случаев это касается повторного использования данных, записанных в кеш, без предварительного анализа содержимого на предмет возможности частичного или комбинированного использования.
-Работа СУБД ограничивается тем, что при выполнении очередного запроса блоки данных не запрашиваются с внешних устройств, если они есть в кеше, т.
-е.
-анализируются номера блоков, а не их содержимое.
+Зарезервированные данные активно используется в системах управления базами данных (СУБД). Но в
+большинстве случаев это касается повторного использования данных, записанных в кеш, без
+предварительного анализа содержимого на предмет возможности частичного или комбинированного
+использования.
 }
 \note<5>{Для осуществления проверки использования кэша анализируются области истинности логических
-ограничений искомого запроса и запросов, результаты которых уже содержатся в кэше. Требуемые
-операции могут быть проведены аналитически, без необходимости дополнительных запросов к базе банных.
+ограничений искомого запроса и запросов, результаты которых уже содержатся в кэше.
+
+Предложенный метод может быть использован для определения недостающих в кэше данных и
+последующего запроса только на эти данные.
+
+Для этого используются аналитические вычисления, что является принципиальным отличием данной работы от существующих технологий.
 }
-\note<6>{Предложенный метод может быть использован для определения недостающих в кэше данных и последующего запроса только на эти данные.
-Для этого также используются аналитические вычисления, что является принципиальным отличием данной работы от существующих технологий.}
 
-\onslide<8->
+\onslide<6->
 {
@@
 \begin{itemize}
-\item<9-> Актуализация данных
-\item<10-> Динамическое формирование многомерных данных
+\item<7-> Актуализация данных
+\item<8-> Динамическое формирование многомерных данных
 \end{itemize}
-\note<9>{Проблема актуализации данных не затрагивается в этой работе. Однако она может быть решена
+\note<7>{Проблема актуализации данных не затрагивается в этой работе. Однако она может быть решена
 путем учета запросов на сервере и обновлении данных при помощи триггеров.}
-\note<10>{Предложенная технология будет использована при динамическом построении многомерных данных.
-Промежуточные представления имеют ту же структуру данных, что и таблицы соединений, используемые для
-построения гиперкубов. Сохраненные представления данных могут храниться на компьютере пользователя-аналитика и существенно сократить время на формирование данных, необходимых для принятия решений.}
+\note<8>{Предложенная технология будет использована при динамическом построении многомерных данных,
+поскольку промежуточные представления имеют ту же структуру данных, что и таблицы соединений,
+используемые для построения гиперкубов.}
 \end{frame}
